### 2008-02-05 trunk/Papers/dl-gre-08/introduction.tex
added structure and some references

\section{Introduction} \label{sec:introduction}

\cite{Dale1995}
\cite{dale91:_gener_refer_expres_invol_relat}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "dl-gre-08"
%%% End: