### 2008-02-13 trunk/Papers/dl-gre-08/introduction.tex
added some more references

\section{Introduction} \label{sec:introduction}

\cite{Dale1995}
\cite{dale91:_gener_refer_expres_invol_relat}

Dale and Haddock is really bad in practice
\cite{viethen06:_algor_for_gener_refer_expres}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "dl-gre-08"
%%% End: