### 2008-03-18 trunk/Papers/dl-gre-08/conclusion.tex
iterated 4.3 and 5; updated runtimes picture

\section{Conclusion} \label{sec:conclusion}

referring to sets (or is this trivial?)

where do we talk about mutually constraining descriptions like ``the
rabbit in the hat''?

will make implementation available

first time to actually show efficient runtimes for generation of
relational REs

Because they compute referring expressions for all individuals in the
domain together, our algorithms will be strongest in static settings,
such as the generation of descriptions for museum exhibits, in which
the individuals and their properties don't change much.  Nevertheless,
they remain fast enough for real-time use in dynamic settings.
Another advantage that recent algorithms for relational REs
\cite{Krahmer2003,kelleher06:_increm_gener_of_spatial_refer} have over
ours is that they can take weights and salience into account.  How
this could be incorporated into our algorithms is an interesting
question for future research.

relational descriptions in data are comparatively rare and
inconsistent; not sufficiently annotated in TUNA
\cite{deemter06:_build_seman_trans_corpus_for}; need more.

Things that we care about but that are never said explicitly so far:

\begin{itemize}
\item in a static scenario, we only have to compute REs once -- and in
  any application that uses many REs before the context changes (e.g.\
  museums), this is exactly what we want to do
\item it is difficult to control the search for concepts by linguistic
  constrtaints
\item our GRE algorithms can take orderings over the propositional and
  relational symbols into account
\end{itemize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "dl-gre-08"
%%% End: