### 2008-03-20 trunk/Papers/dl-gre-08/abstract.tex

\begin{abstract}
  In this paper, we propose to reinterpret the problem of generating
  referring expressions (GRE) as the problem of computing a formula in a
  description logic that is only satisfied by the referent.  This view
  offers a new unifying perspective under which existing GRE
  algorithms can be compared.  We also show that by applying existing
  algorithms for computing bisimulation classes in description logic,
  we can obtain extremely efficient algorithms for relational
  referring expressions without any danger of running into infinite
  regress.
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "dl-gre-08"
%%% End: