### 2008-03-20 trunk/Papers/dl-gre-08/bisim.tex
iterated over 2-3; paper is down to 8 pages + 3 refs

\section{Description Languages and Bisimulation} \label{sec:bisim}

The main point of this paper is to represent referring expressions
(REs) as formulas of description logic \cite{baad:desc03}.  In order
to make this point, we will now first define the two description
logics we will be working with -- \alc\ and \el.

\begin{definition}
Formulas $\varphi$ of $\alc$ are generated by the following grammar:
$$
\varphi,\varphi' ::= p \mid \neg \varphi \mid \varphi \sqcap \varphi'
\mid \exists R. \varphi
$$
where $p$ is in the set of propositional symbols \prop, $R$ is in the
set of relational symbols \rel.

$\el$ is the negation-free fragment of $\alc$.

Formulas of both $\alc$ and $\el$ are interpreted in ordinary
relational first-order models $\gM = (\Delta^\gM,|\cdot|^\gM)$ where
$\Delta^\gM$ is some non-empy set, and $|\cdot|^\gM$ is an
interpretation function such that:
$$
\begin{array}{ccl}
|p|^\gM & \subseteq & \Delta^\gM  \mbox{ for $p \in \prop$}\\
|R|^\gM & \subseteq & \Delta^\gM \times \Delta^\gM  \mbox{ for $R \in \rel$}\\
|\neg \varphi|^\gM & = & \Delta^\gM \backslash |\varphi|^\gM\\
|\varphi \sqcap \varphi'|^\gM & = & |\varphi|^\gM \cap |\varphi'|^\gM\\
|\exists R.\varphi|^\gM & = & \{i \mid \mbox{for some } i', (i,i') \in |R|^\gM\\
& & \mbox{ and } i' \in |\varphi|^\gM \}.\\
\end{array}
$$
Given a model $\gM$ and an individual $i$ in the domain of $\gM$, let
$\propm(i) = \{p \in \prop \mid i \in |p|^\gM\}$.
\end{definition}

Every formula of a description logic denotes a set of individuals in
the domain; thus we can use such formulas to describe sets.  For
instance, in the model in Fig.~\ref{fig:dale-haddock}b, the formula
$\mathsf{flower}$ denotes the set $\{f_1,f_2\}$; the formula
$\mathsf{flower} \sqcap \exists \mathsf{in}.\mathsf{hat}$ denotes
$\{f_2\}$; and the formula $\mathsf{flower} \sqcap \neg
\exists.\mathsf{hat}$ denotes $\{f_1\}$.

Different description logics differ in the inventory of logical
connectives they permit.  The difference between \alc\ and \el\ is
that \alc\ permits negation and \el\ doesn't.  There are many other
description logics in the literature; some that we will get back to in
Section~\ref{sec:related} are $\mathcal{CL}$ (\el\ without existential
quantification, i.e.\ only conjunctions of atoms); $\mathcal{PL}$
(\alc\ without existential quantification, i.e.\ propositional logic);
and $\mathcal{ELU}_{(\neg)}$ (\el\ plus disjunction and atomic
negation).

A key notion in studying description logics is \emph{bisimulation}.
We will discuss in detail the case for $\alc$; further details can be
found in~\cite{blac:moda01,kurt:expr99}.

\begin{definition}
Given a model $\gM = (\Delta^\gM,|\cdot|^\gM)$, two elements $i$ and
$i'$ of $\Delta^\gM$ are \emph{$\alc$-bisimilar} if there is a relation $Z
\subseteq \Delta^\gM \times \Delta^\gM$ such that 
\begin{itemize}
\item[i)] $(i,i') \in Z$.\\[-1.5em]
\item[ii)] if $(e_1, e_2) \in Z$ then $\propm(e_1) = \propm(e_2)$.\\[-1.5em]
\item[iii)] if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_1'$ such that
$(e_1,e_1') \in |R|^\gM$ then there is $e_2'$ such that $(e_2,e_2') \in |R|^\gM$ and
$(e_1',e_2') \in Z$.\\[-1.5em]
\item[iv)] if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_2'$ such that
$(e_2,e_2') \in |R|^\gM$ then there is $e_1'$ such that $(e_1,e_1') \in |R|^\gM$ and
$(e_1',e_2') \in Z$.
\end{itemize}
A maximal subset of $\Delta^\gM$ all of whose elements are pairwise
bisimilar is called an \emph{\alc-bisimulation class} of $\gM$.
\end{definition}

The crucial property that makes bisimulation relevant for GRE is the
following theorem \cite{blac:moda01}.

\begin{theorem}\label{bisim}
  Given a finite model $\gM$, $i$ and $i'$ are $\alc$-bisimilar in
  $\gM$ if and only if there is no $\alc$-formula $\varphi$ such that
  $i \in |\varphi|^\gM$ and $i' \not \in |\varphi|^\gM$.
\end{theorem}

In other words, $\alc$-bisimilar individuals are indistinguishable in
the $\alc$ language.  In the example in Fig.~\ref{fig:dale-haddock}a,
the two individuals $t_1$ and $t_2$ are \alc-bisimilar; and indeed,
they satisfy exactly the same \alc-formulas.  Although we can't go
into details here, it is possible to define a suitable notion of
$\el$-bisimulation for which the analogous theorem holds
\cite{kurt:expr99}.  In general, \el-bisimulation is a weaker notion
than \alc-bisimulation; for instance, in Fig.~\ref{fig:dale-haddock},
$f_1$ and $f_2$ are \el-bisimilar \todo{oops! here's the asymmetry,
  there is a formula for $f_2$ but not for $f_1$} (no \el-formula will
distinguish them), but we have already seen that they satisfy
different \alc-formulas.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "dl-gre-08"
%%% End: