### 2008-03-21 trunk/Papers/dl-gre-08/bisim.tex
removed simulations

\section{Description logics and similarity}
\label{sec:bisim}

In this paper, we will represent referring expressions (REs) as
formulas of description logic~\cite{baad:desc03}.  In order to make
this point, we will now first define the two description logics we
will be working with: \alc\ and \el.

\begin{definition}
  Formulas (or concepts) $\varphi$ of $\alc$ are generated by the
  following grammar:
$$
\varphi,\varphi' ::= p \mid \neg \varphi \mid \varphi \sqcap \varphi'
\mid \exists R. \varphi
$$
where $p$ is in the set of propositional symbols \prop, $R$ is in the
set of relational symbols \rel. $\el$ is the negation-free fragment of $\alc$.

Formulas of both $\alc$ and $\el$ are interpreted in ordinary
relational first-order models $\gM = (\Delta^\gM,\interp{\cdot}^\gM)$ where
$\Delta^\gM$ is a non-empy set and $\interp{\cdot}^\gM$ is an
interpretation function such that:
$$
\begin{array}{ccl}
\interp{p}^\gM & \subseteq & \Delta^\gM  \mbox{ for $p \in \prop$}\\
\interp{R}^\gM & \subseteq & \Delta^\gM \times \Delta^\gM  \mbox{ for $R \in \rel$}\\
\interp{\neg \varphi}^\gM & = & \Delta^\gM \backslash \interp{\varphi}^\gM\\
\interp{\varphi \sqcap \varphi'}^\gM & = & \interp{\varphi}^\gM \cap \interp{\varphi'}^\gM\\
\interp{\exists R.\varphi}^\gM & = & \{i \mid \mbox{for some } i', (i,i') \in \interp{R}^\gM\\
& & \mbox{ and } i' \in \interp{\varphi}^\gM \}.\\
\end{array}
$$

Given a model $\gM$ and an individual $i$ in $\gM$, let
$\propm(i) = \{p \in \prop \mid i \in \interp{p}^\gM\}$.
\end{definition}

Every formula of a description logic denotes a set of individuals in
the domain; thus we can use such formulas to describe sets.  For
instance, in the model in Fig.~\ref{fig:dale-haddock}b, the formula
$\mathsf{flower}$ denotes the set $\{f_1,f_2\}$; the formula
$\mathsf{flower} \sqcap \exists \mathsf{in}.\mathsf{hat}$ denotes
$\{f_2\}$; and the formula $\mathsf{flower} \sqcap \neg
\exists \mathsf{in}.\mathsf{hat}$ denotes $\{f_1\}$.

Different description logics differ in the inventory of logical
connectives they permit.  The difference between \alc\ and \el\ is
that \alc\ permits negation and \el\ doesn't.  There are many other
description logics in the literature; some that we will get back to in
Section~\ref{sec:related} are $\mathcal{CL}$ (\el\ without existential
quantification, i.e., only conjunctions of atoms); $\mathcal{PL}$
(\alc\ without existential quantification, i.e., propositional logic);
and $\mathcal{ELU}_{(\neg)}$ (\el\ plus disjunction and atomic
negation).

A key notion in description logics that we will use here is
\emph{similarity} \todo{citation?}.  For any description logic $\gL$,
we will say that an individual $i$ is \emph{\gL-similar} to $i'$ in a
given model $\gM$ if for any formula $\varphi \in \gL$ such that $i
\in \interp{\varphi}$, we also have $i' \in \interp{\varphi}$.
Intuitively, this means that there is no $\gL$-formula that would
distinguish $i$ from $i'$.  We say that the \emph{\gL-similarity set}
of some individual $i$ is the set of all individuals to which $i$ is
\gL-similar.

Notice that similarity is not necessarily a symmetrical relation: For
instance, $f_1$ is \el-similar to $f_2$ in
Fig.~\ref{fig:dale-haddock}b, but $f_2$ is not \el-similar to $f_1$
(it satisfies the formula $\exists \mathsf{in}.\mathsf{hat}$ and $f_1$
doesn't).  However, \alc-similarity is a symmetrical relation because
\alc\ contains negation; and indeed, $f_1$ is not \alc-similar to
$f_2$ either because it satisfied $\neg \exists
\mathsf{in}.\mathsf{hat}$.  Because \alc\ is more expressive than \el,
it is generally possible for some individual $a$ to be \el-similar but
not \alc-similar to some individual $b$, but not vice versa.

\todo{need to cite \cite{blac:moda01,kurt:expr99} in the appropriate
  places}

\ignore{

Notice
that, given that the language includes negation, the notion of
$\alc$-similarity is symmetric: if $i$ is \alc-similar to $i'$ then
$i'$ is \alc-similar to $i$.  This is not the case for all description
logic.  In particular, the notion of $\el$-similarity is not
symmetric.  Given a model $\gM$ and an individual $i$ in $\gM$, we
will write \simm$_\alc(i)$, the similarity set, to all the elements of
$\gM$ which are \alc-similar to $i$.x

A key notion in studying description logics is \emph{simulation}.
We will discuss in detail the case for $\alc$; further details can be
found in~\cite{blac:moda01,kurt:expr99}.

\begin{definition}
Given a model $\gM = (\Delta^\gM,\interp{\cdot}^\gM)$, and two individuals $i$ and $i'$ in $\gM$, an \emph{$\alc$-simulation} linking $i$ to $i'$  is a relation $Z
\subseteq \Delta^\gM \times \Delta^\gM$ such that 
\begin{itemize}
\item[i)] $(i,i') \in Z$.\\[-1.5em]
\item[ii)] if $(e_1, e_2) \in Z$ then $\propm(e_1) = \propm(e_2)$.\\[-1.5em]
\item[iii)] if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_1'$ such that
$(e_1,e_1') \in \interp{R}^\gM$ then there is $e_2'$ such that $(e_2,e_2') \in \interp{R}^\gM$ and
$(e_1',e_2') \in Z$.\\[-1.5em]
\item[iv)] if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_2'$ such that
$(e_2,e_2') \in \interp{R}^\gM$ then there is $e_1'$ such that $(e_1,e_1') \in \interp{R}^\gM$ and
$(e_1',e_2') \in Z$.
\end{itemize}
%A maximal subset of $\Delta^\gM$ all of whose elements are pairwise
%bisimilar is called an \emph{\alc-bisimulation class} of $\gM$.
\end{definition}

The crucial property that makes simulation relevant for GRE is the
following theorem \cite{blac:moda01}.

\begin{theorem}\label{bisim}
  Given a finite model $\gM$ and two individuals $i$ and $i'$ in $\gM$, there is an $\alc$-simulation linking $i$ to $i'$ if and only if $i$ is $\alc$-similar to $i'$.
\end{theorem}

Notice that, because for $\alc$ the notion of similarity si symmetric, $\alc$-similar individuals are indistinguishable in the $\alc$ language.  In the example in Fig.~\ref{fig:dale-haddock}a,
the two individuals $t_1$ and $t_2$ are \alc-similar; and indeed,
they satisfy exactly the same \alc-formulas.

Although we can't go
into details here, it is possible to define a suitable notion of
$\el$-simulation for which the analogous theorem holds
\cite{kurt:expr99}.  In general, \el-simulation is a weaker notion
than \alc-simulation; for instance, in Fig.~\ref{fig:dale-haddock},
$f_1$ is \el-similar to $f_2$, but not vice-versa.

}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "dl-gre-08"
%%% End: