### 2008-03-26 trunk/Papers/rabbit-from-hat/howto.tex
structure

\section{Generating succinct expressions}
\label{sec:howto}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rabbit-from-hat"
%%% End: