### 2008-05-16 trunk/Papers/dl-gre-08/gre.tex
fits on 8 pages; finished as far as I am concerned

\section{Generating referring expressions} \label{sec:gre}

Now we apply description logic to GRE.  The core claim of this paper
is that it is natural and useful to view the GRE problem as the
problem of computing a formula of some description logic $\gL$ whose
extension is a given target set $A$ of individuals.

\medskip
\noindent
{\small
\begin{center}
\begin{tabular}{ll} \hline
\multicolumn{2}{l}{
\textsc{$\gL$-GRE Problem}}\\ \hline
\ \ Input: & A model $\gM$ and a target set $A \subseteq \Delta$.\\
\ \ Output: & A formula $\varphi \in \gL$ such that
$\interp{\varphi} = A$\\ 
& \hspace*{0.5cm} (if such a formula exists).\\ \hline
\end{tabular}
\end{center}}
\medskip

In the examples above, it is because $\mathsf{flower} \sqcap \exists
\mathsf{in}. \mathsf{hat}$ denotes exactly $\{f_2\}$ that we can say
``the flower in the hat'' to refer to $f_2$.  This perspective
provides a general framework into which many existing GRE approaches
fit: Traditional attribute selection \cite{Dale1995} corresponds to
building DL formulas that are conjunctions of atoms; relational REs as
in \newcite{dale91:_gener_refer_expres_invol_relat} are formulas of
\el; and so on.  We will further pursue the idea of organizing GRE
approaches with respect to the variant of DL they use in
Section~\ref{sec:related}.

For the rest of this paper, we assume that we are generating a
singular RE, i.e.\ the target set $A$ will be singleton.  In this
case, we will only be able to generate a formula that denotes exactly
$A = \{a\}$ (i.e., a RE that uniquely refers to $a$) if there is no
other individual $b$ to which $a$ is similar; otherwise, any formula
that is satisfied by $a$ is also satisfied by $b$.  Conversely, if we
know that $a$ is not similar to any other individual, then there is a
formula that is satisfied by $a$ and not by anything else; this
formula can serve as a unique singular RE.  In other words, we can
reduce the $\gL$-GRE problem for a given model to the problem of
computing the $\gL$-similarity sets of this model.  Notice that this
use of similarity sets can be seen as a generalization of van
Deemter's \shortcite{deemter02:_gener_refer_expres} ``satellite sets''
to relational descriptions.

In the rest of this section, we will present algorithms that compute
the similarity sets of a given model for \alc\ and \el, together with
characteristic formulas that denote them.  In the \alc\ case, we adapt
a standard algorithm from the literature for computing
\emph{simulation classes}; we will then further adapt this algorithm
for \el.  In effect, both algorithms compute REs for all individuals
in some model at the same time -- very efficiently and without any
danger of infinite regress.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{pic-dale-haddock}
  \caption{(a) The \newcite{dale91:_gener_refer_expres_invol_relat}
    scenario; (b) the \newcite{Stone1998a} scenario.}
  \label{fig:dale-haddock}
\end{figure}

\subsection{Computing similarity sets}

It can be shown that for \alc, the similarity sets of a finite model
coincide exactly with the \emph{simulation classes} of this model.
Simulation classes have been studied extensively in the literature
(see e.g. \newcite{blac:moda01}; \newcite{kurt:expr99}), and there are several
efficient algorithms for computing \alc-simulation classes
\cite{hopc:algo71,paig:thre87,dovier04:_effic_algor_for_comput_bisim_equiv}.
However, these algorithms will only compute the simulation classes
themselves. Here we extend the \newcite{hopc:algo71} algorithm such
that it computes, along with each set, also a formula that denotes
exactly this set.  We can then use these formulas as representations
of the referring expressions.

The pseudocode for our \alc\ algorithm is shown as
Algorithm~\ref{algo:bisim-l} (with $\gL=\alc$) and
Algorithm~\ref{algo:bisim-add-alc}.  Given a model $\gM=(\Delta,
\interp{\cdot})$, the algorithm computes a set $\RE$ of $\alc$
formulas such that $\{\interp{\varphi} \mid \varphi \in \RE\}$ is the
set of \alc-similarity sets of $\gM$.  The algorithm starts with $\RE
= \{\top\}$ (where $\interp{\top} = \Delta$), and successively refines
$\RE$ by making its elements denote smaller and smaller sets.  It
maintains the invariant that at the start and end of every iteration,
$\{\interp{\varphi} \mid \varphi \in \RE\}$ is always a partition of
$\Delta$.  The algorithm iterates over all propositional and
relational symbols in $\prop$ and $\rel$ to construct new formulas
until either all formulas in $\RE$ denote singletons (i.e., there is
only one individual that satisfies them), or no progress has been made
in the previous iteration.  In each iteration, it calls the procedure
add$_\alc$($\varphi$, $\RE$), which intersects $\varphi$ with any
formula $\psi \in \RE$ which does not denote a singleton and which is
not equivalent to $\varphi$ and to $\neg \varphi$. In this case, it
replaces $\psi$ in $\RE$ by $\psi \sqcap \varphi$ and $\psi \sqcap
\neg \varphi$.

The \alc\ algorithm computes the \alc-similarity sets of the model in
time $O(n^3)$, where $n$ is the number of individuals in the domain.
However, it will freely introduce negations in the case distinctions,
which can make the resulting formula hard to realize (see also
Section~\ref{sec:discussion-realization}).  This is why we also
present an algorithm for the \el-similarity sets; \el\ corresponds to
positive relational REs, which are generally much easier to realize.

We obtain the \el\ algorithm by replacing the call to add$_{\alc}$ in
Algorithm~\ref{algo:bisim-l} by a call to add$_{\el}$, which is
defined in Algorithm~\ref{algo:bisim-add-el}.  As before, the
algorithm maintains a set $\RE = \{\varphi_1,\ldots,\varphi_n\}$ of
formulas (this time of \el) such that $\interp{\varphi_1} \cup \ldots
\cup \interp{\varphi_n} = \Delta$, and which it refines iteratively.
However, where the \alc\ algorithm maintains the invariant that
$\interp{\varphi_1},\ldots,\interp{\varphi_n}$ is a partition of
$\Delta$, we weaken this invariant to the requirement that there are
no $m \geq 2$ pairwise different indices $1 \leq i_1,\ldots,i_m \leq
n$ such that $\interp{\varphi_{i_1}} = \interp{\varphi_{i_2}} \cup
\ldots \cup \interp{\varphi_{i_m}}$.  We call $\varphi_{i_1}$
\emph{subsumed} if such a decomposition exists.

Because it maintains a weaker invariant, the set $\RE$ may contain
more formulas at the same time in the \el\ algorithm than in the \alc\
algorithm.  Given that $\Delta$ has an exponential number of subsets,
there is a risk that the \el\ algorithm might have worst-case
exponential runtime (although we are not aware of such worst-case
examples).  We leave a more careful runtime analysis for future work.

\begin{algorithm}[t]
\dontprintsemicolon
\caption{Computing the $\mathcal{L}$-similarity sets}
\label{algo:bisim-l}
\KwIn{A model $\gM = (\Delta, \interp{\cdot})$}
\KwOut{A set \RE of formulas  such that
$\{\interp{\varphi} \mid \varphi \in \RE\}$ is the set of
$\mathcal{L}$-similarity 
sets of $\gM$.}

$\RE \leftarrow \{\top\}$

\For{$p \in \prop$}{
      add$_\mathcal{L}(p,\RE)$
   }

\While{exists some $\varphi \in \RE, |\interp{\varphi}|^\gM>1$}{
   \For{$\varphi \in \RE, R \in \rel$}{
         add$_\mathcal{L}(\exists R.\varphi,\RE)$
   }
   \If{made no changes to \RE}{
      exit
      }
}
\end{algorithm}

Notice that both algorithms try to refine $\RE$ first according to the
propositional symbols, and then by relational expressions of
increasing depth.  The algorithms would remain correct if we
interleaved the addition of propositional and relational symbols;
however, the algorithms would become slower because the propositional
refinements would have to be percolated through the relational chains
in subsequent iterations.

\subsection{Some examples}\label{sec:examples}

Let's try our algorithms on some examples.  We first run the \el\
algorithm on the model shown in Fig.~\ref{fig:dale-haddock}a, which is
taken from \newcite{dale91:_gener_refer_expres_invol_relat}.  The
algorithm starts with $\RE = \{\top\}$.  In the first loop, it adds
the formulas $\mathsf{floor}$, $\mathsf{bowl}$, $\mathsf{cup}$, and
$\mathsf{table}$, and then removes $\top$ because it is now subsumed.
Not all of these formulas denote singletons; for instance,
$\interp{\mathsf{cup}}$ contains two individuals.  So we iterate over
the relations to refine our formulas.  After the first iteration over
the relations, we have $\RE = \{ \mathsf{floor}, \mathsf{bowl} \sqcap
\exists \mathsf{on}.\mathsf{floor}, \mathsf{bowl} \sqcap \exists
\mathsf{on}.\mathsf{table}, \mathsf{cup}, \mathsf{table} \}$. Notice
that $\mathsf{bowl}$ has become subsumed, but we haven't distinguished
the cups and tables further.

\begin{algorithm}[t]
\caption{add$_\alc(\varphi,\RE)$}
\label{algo:bisim-add-alc}
\For{$\psi \in \RE$ with $|\interp{\psi}| > 1$}{
   \If{$\interp{\psi \sqcap \varphi} \not = \emptyset$ and
       $\interp{\psi \sqcap \neg \varphi} \not = \emptyset$}{
         add $\psi \sqcap \varphi$ and
               $\psi \sqcap \neg \varphi$ to \RE \;
         remove $\psi$ from \RE \;
      }
   }
\end{algorithm}
%
\begin{algorithm}[t]
\dontprintsemicolon
\caption{add$_\el$($\varphi$, $\RE$)}
\label{algo:bisim-add-el}
\For{$\psi \in \RE$ with $|\interp{\psi}| > 1$}{
  \If{$\psi \sqcap \varphi$ is not subsumed in $\RE$ {\bf and}
    $\interp{\psi \sqcap \varphi} \neq \emptyset$ {\bf and}
    $\interp{\psi \sqcap \varphi} \neq \interp{\psi}$}{
    add $\psi \sqcap \varphi$ to $\RE$ \;
    remove subsumed formulas from $\RE$\;
  }
}
\end{algorithm}

Now we can use the split between the bowls to distinguish the cups in
the second iteration.  The result of this is $\RE = \{ \mathsf{floor},
\mathsf{bowl} \sqcap \exists \mathsf{on}.\mathsf{floor}, \mathsf{bowl}
\sqcap \exists \mathsf{on}.\mathsf{table}, \mathsf{cup} \sqcap \exists
\mathsf{in}. (\mathsf{bowl} \sqcap \exists
\mathsf{on}.\mathsf{floor}), \mathsf{cup} \sqcap \exists
\mathsf{in}. (\mathsf{bowl} \sqcap \exists
\mathsf{on}.\mathsf{table}), \mathsf{table} \}$.  At this point, all
formulas except $\mathsf{table}$ denote singletons, and further
iterations don't allow us to refine $\mathsf{table}$; so the algorithm
terminates.  Each formula with a singleton extension $\{a\}$ is a
unique description of $a$; for instance, $\mathsf{cup} \sqcap \exists
\mathsf{in}. (\mathsf{bowl} \sqcap \exists
\mathsf{on}.\mathsf{table})$ is only satisfied by $c_2$, so we may
refer to $c_2$ as ``the cup in the bowl on the table''.  Notice that
the algorithm didn't focus on any particular individual; it
simultaneously generated REs for all individuals except for the two
tables (which are similar to each other).

The \el\ algorithm has a harder time with the example in
Fig.~\ref{fig:dale-haddock}b \cite{Stone1998a}.  While it will
correctly identify $r_1$ as ``the rabbit in the hat'' and $f_2$ as
``the flower in the hat'', it will not be able to compute a RE for $f_1$
because $f_1$ is \el-similar to $f_2$.  Indeed, the algorithm
terminates with $\RE$ containing both $\mathsf{flower}$ and
$\mathsf{flower} \sqcap \exists \mathsf{in}.\mathsf{hat}$.  This is a
typical pattern for asymmetrical cases of similarity in \el: If there
are two formulas $\varphi_1$ and $\varphi_2$ in the output set with
$\interp{\varphi_1} \subseteq \interp{\varphi_2}$, then there is
generally some individual $b \in \interp{\varphi_2} -
\interp{\varphi_1}$ such that all individuals in $\interp{\varphi_1}$
are similar to $b$, but not vice versa.  By contrast, the \alc\
algorithm can exploit the greater expressivity of \alc\ to split
$\mathsf{flower}$ into the two new formulas $\mathsf{flower} \sqcap
\exists \mathsf{in}.\mathsf{hat}$ and $\mathsf{flower} \sqcap \neg
\exists \mathsf{in}.\mathsf{hat}$, generating a unique RE for $f_1$ as
well.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "dl-gre-08"
%%% End: