### 2008-05-20 trunk/Papers/rmrs-08/abstract.tex
first dummy version

\begin{abstract}
  la la la
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rmrs-08"
%%% End: