### 2008-06-05 trunk/Papers/experiences/experiments.tex

\section{Experiments} \label{sec:experiments}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: