### 2008-06-05 trunk/Papers/experiences/discussion.tex
wrote something about the experiments

\section{Discussion} \label{sec:discussion}

planning competition makes for great starting point for users

grounding only makes sense if the plans are long, compared to the
domain size

all sorts of minor problems: discovered bugs both in SGPLAN and FF;
FF input files limited to about 10.000 lines

typed domain is not a silver bullet -- sometimes we just have lots of
things of the same kind

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: