### 2008-06-05 trunk/Papers/experiences/conclusion.tex
now we have something in every section

\section{Conclusion} \label{sec:conclusion}

In this paper, we have introduced two novel planning domains related
to problems in natural language generation: the sentence generation
problem and the GIVE navigation problem.  We have then reported on a
number of experiments in which we applied off-the-shelf planners to
these domains.  Our main results were that while modern planners do
pretty well in controlling the complexity of search, they suffer from
practical problems that limit their performance in unexpected places.
In particular, in both domains the grounding step that both planners
perform tends to take much more time than the search itself.

\todo{say something clever about the future}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: