### 2008-06-19 trunk/Papers/experiences/experiences.tex
changes to author block and introduction

\documentclass[letterpaper]{article}

\usepackage{aaai} 
\usepackage{times} 
\usepackage{helvet} 
\usepackage{courier}

\usepackage{graphicx}

\newcommand{\todo}[1]{\textbf{(#1)}}

\title{Experiences with planning for natural language generation}
\author{Alexander Koller \and Ronald Petrick \\
School of Informatics \\
University of Edinburgh \\
Edinburgh \ EH8 9AB, Scotland, UK \\
\texttt{\{a.koller,r.petrick\}@ed.ac.uk}}

\nocopyright

\begin{document}

\maketitle

\begin{abstract}
  lkajsd
\end{abstract}

\input introduction
\input domains
\input experiments
\input discussion
\input conclusion

\bibliography{bibliography}
\bibliographystyle{aaai}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: