### 2008-06-19 trunk/Papers/experiences/introduction.tex

\section{Introduction}
\label{sec:introduction}

Natural language generation (NLG; \citeauthor{reiter00building}
\citeyear{reiter00building}) is one of the major subfields of natural
language processing, concerned with computing natural language sentences or
texts that convey a given piece of information to the user. Intuitively,
this task can be viewed as a problem involving actions, beliefs, and goals:
an agent communicating with another agent tries to change the mental state
of the hearer by applying actions which correspond to the utterance of
words or sentences. This characterization of NLG suggests obvious parallels
to automated planning---a view which has a long tradition in NLG
\cite{appelt:planning,young94dpocl}.  More recently, there has been a
resurgence of interest in applying modern planning techniques to NLG
\cite{Steedman-Petrick:07,KolSto07,benotti08b}.  While efficiency has not
been the main focus of NLG, problems in NLG are often more complex than
their counterparts in other areas of natural language processing (such as
parsing \cite{KolStr02}), giving rise to computationally challenging
domains. These new approaches combine an interest in planning as a
modelling tool for natural language, with a hope of improved efficiency by
exploiting modern approaches and algorithms.

%Natural language generation (NLG; \citeauthor{reiter00building} 2000)
%is one of the major subfields of natural language processing.  It is
%concerned with computing natural language sentences or texts that
%convey a given piece of information to the user.  Problems in NLG are
%often more complex than their counterparts in parsing \cite{KolStr02},
%and indeed efficiency has not traditionally been the main focus of
%NLG.  On the other hand, NLG has some obvious parallels with planning:
%One tries to change the mental state of the hearer by applying actions
%which correspond to uttering words or sentences.  This parallel has a
%long tradition in NLG \cite{appelt:planning,young94dpocl}, and there
%has recently been a resurgence of interest in applying modern planning
%techniques to NLG \cite{Steedman-Petrick:07,KolSto07,benotti08b}.
%These new approaches combine an interest in planning as a modelling
%tool for natural language with a hope for improved efficiency by
%exploiting modern planning algorithms to help with the search.

%%%%%
%The purpose of this paper is twofold.

In this paper, we present two recent planning domains that arise in
the context of applying planning to natural language generation.  We
will first review the sentence generation domain, which was cast as a
planning problem by Koller and Stone \citeyear{KolSto07}.  In this
domain, the goal is to generate a single sentence that expresses a
given piece of meaning; the plan encodes this sentence, and the
actions correspond to uttering individual words.  We will then
introduce the GIVE domain (``Generating Instructions in Virtual
Environments'').  GIVE is a new shared task that was recently posed as
a challenge for the NLG community
\cite{alexander07:_shared_task_propos}.  It uses planning as one
module of a NLG system which generates natural-language instructions
that will guide the user towards performing a given task in a virtual
environment.

We will then discuss some experiences with using off-the-shelf
planners in these domains.  In particular, we explore the efficiency
of FF \cite{HoffmannNebel01} and SGPLAN
\cite{hsu06:_new_featur_in_sgplan_for} on instances of our planning
domains of varying sizes.  Our findings are mixed.  On the one hand,
it turns out that modern planners handle the search problems that
arise in NLG quite easily.  On the other hand, they spend tremendous
amounts of time on preprocessing; for instance, FF spends 90\% of its
runtime in the sentence generation domain on grounding out the
literals and actions, and ends up slower than an ad-hoc Java
implementation of GraphPlan that avoids grounding.  We argue that for
domains like ours, which are dominated by the number of actions and the
universe size rather than the combinatorics of the search problem, the
overall runtime of a planner could be improved by grounding more
selectively.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: