### 2008-06-19 trunk/Papers/experiences/introduction.tex
more changes to the introduction

\section{Introduction}
\label{sec:introduction}

Natural language generation (NLG; \citeauthor{reiter00building}
\citeyear{reiter00building}) is one of the major subfields of natural
language processing, concerned with computing natural language sentences or
texts that convey a given piece of information to the user. Intuitively,
this task can be viewed as a problem involving actions, beliefs, and goals:
an agent communicating with another agent tries to change the mental state
of the hearer by applying actions which correspond to the utterance of
words or sentences. This characterization of NLG suggests obvious parallels
to automated planning---a view which has a long tradition in NLG
\cite{appelt:planning,young94dpocl}.  More recently, there has been a
resurgence of interest in applying modern planning techniques to NLG
\cite{Steedman-Petrick:07,KolSto07,benotti08b}.  While efficiency has not
been the main focus of NLG, problems in NLG are often more complex than
their counterparts in other areas of natural language processing (such as
parsing \cite{KolStr02}), giving rise to computationally challenging
domains. These new approaches combine an interest in planning as a
modelling tool for natural language, with a hope of improved efficiency by
exploiting modern approaches and algorithms.

The focus of this paper is twofold. We begin by presenting two recent
planning domains that arise in the context of natural language generation:
the sentence generation domain and the GIVE domain. In the sentence
generation task, the goal is to generate a single sentence that expresses a
given meaning. \citeauthor{KolSto07}~\shortcite{KolSto07} cast this domain
as a planning problem, where the plan encodes the necessary sentence and
the actions correspond to uttering individual words. In the GIVE domain
(``Generating Instructions in Virtual Environments''), we describe a new
shared task that was recently posed as a challenge for the NLG community
\cite{alexander07:_shared_task_propos}. GIVE uses planning as one module of
an NLG system that generates natural-language instructions to guide a
user performing a given task in a virtual environment.

We will then discuss some of our experiences using off-the-shelf planners
in these domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN \cite{hsu06:_new_featur_in_sgplan_for} on
varying-sized instances of our planning domains. We also compare these
results against an ad-hoc Java implementation of GraphPlan.  Our findings
are mixed. On the one hand, it turns out that modern planners handle some
of the search problems that arise in NLG quite easily. (Large problem
instances remain a challenge.) On the other hand, these same planners spend
tremendous amounts of time on preprocessing. For instance, FF spends 90\%
of its runtime in the sentence generation domain on grounding out literals
and actions, and ends up slower than our version of GraphPlan that avoids
grounding. For domains like ours, which are dominated by the number of
actions and the universe size rather than the combinatorics of the search
problem, this observation suggests that the overall runtime of a planner
could be improved by grounding more selectively.

%Natural language generation (NLG; \citeauthor{reiter00building} 2000)
%is one of the major subfields of natural language processing.  It is
%concerned with computing natural language sentences or texts that
%convey a given piece of information to the user.  Problems in NLG are
%often more complex than their counterparts in parsing \cite{KolStr02},
%and indeed efficiency has not traditionally been the main focus of
%NLG.  On the other hand, NLG has some obvious parallels with planning:
%One tries to change the mental state of the hearer by applying actions
%which correspond to uttering words or sentences.  This parallel has a
%long tradition in NLG \cite{appelt:planning,young94dpocl}, and there
%has recently been a resurgence of interest in applying modern planning
%techniques to NLG \cite{Steedman-Petrick:07,KolSto07,benotti08b}.
%These new approaches combine an interest in planning as a modelling
%tool for natural language with a hope for improved efficiency by
%exploiting modern planning algorithms to help with the search.

%In this paper, we present two recent planning domains that arise in the
%context of applying planning to natural language generation.  We will
%first review the sentence generation domain, which was cast as a planning
%problem by \citeauthor{KolSto07}~\shortcite{KolSto07}.  In this domain,
%the goal is to generate a single sentence that expresses a given piece of
%meaning; the plan encodes this sentence, and the actions correspond to
%uttering individual words.  We will then introduce the GIVE domain
%(``Generating Instructions in Virtual Environments'').  GIVE is a new
%shared task that was recently posed as a challenge for the NLG community
%\cite{alexander07:_shared_task_propos}.  It uses planning as one module of
%a NLG system which generates natural-language instructions that will guide
%the user towards performing a given task in a virtual environment.

%We will then discuss some experiences with using off-the-shelf planners in
%these domains.  In particular, we explore the efficiency of FF
%\cite{HoffmannNebel01} and SGPLAN \cite{hsu06:_new_featur_in_sgplan_for} on
%instances of our planning domains of varying sizes.  Our findings are
%mixed.  On the one hand, it turns out that modern planners handle the
%search problems that arise in NLG quite easily.  On the other hand, they
%spend tremendous amounts of time on preprocessing; for instance, FF spends
%90\% of its runtime in the sentence generation domain on grounding out the
%literals and actions, and ends up slower than an ad-hoc Java implementation
%of GraphPlan that avoids grounding.  We argue that for domains like ours,
%which are dominated by the number of actions and the universe size rather
%than the combinatorics of the search problem, the overall runtime of a
%planner could be improved by grounding more selectively.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: