### 2008-06-23 trunk/Papers/experiences/experiments.tex
changes to experiments section (experiment 1)

\section{Experiments} \label{sec:experiments}

We now describe the results of a set of experiments designed to evaluate
the suitability of particular planners (FF, SGPLAN, and an ad-hoc
implementation of GraphPlan) in our two NLG domains. As a first impression,
our results indicate that planning is a promising tool for both domains. In
the GIVE domain, SGPLAN 5.2.2 computes a domain plan from the initial state
to the goal in 0.3 seconds, which is fast enough for moderately-sized
problems instances in the application.\footnote{All runtimes were
  measured on a Pentium 4 CPU running at 3 GHz. Java programs were allowed
  to ``warm up'', i.e.\ the planner was run five times and the first four
  measurements discarded to ensure that the JVM had just-in-time compiled
  all relevant bytecode.}
In the sentence generation domain, FF dramatically outperforms the best
previously known algorithm for the same problem (a reimplementation of
\cite{Stone2003a}), although the latter is a greedy search algorithm with a
heuristic that is hand-tailored to the domain.

We also observe that although FF manages the search for a plan very
efficiently, it spends comparatively large amounts of time computing
instantiations of predicates and actions, most of which are then never used
during the search. As a result, FF's ``grounding time'' dominates its
overall planning time, leading to some conflicting results. Our experiments
below seek to improve our understanding of this situation.

\subsection{Experiment 1: Sentence generation}
\label{sec:exper-1:-sent}

In the first experiment, we generate a series of sentence generation
problems which require the planner to compute a plan representing the
sentence ``Mary likes the Adj$_1$ \ldots Adj$_n$ rabbit.''  Each problem
instance assumes a certain number $m$ of rabbits that were distinguished by
$n \leq m$ different properties, such that all $n$ properties are required
to distinguish the target referent from all other rabbits.  The $n$
properties are realized as $n$ different adjectives, in any order.  This
setup allows us to control the plan length (a plan with $n$ properties will
have length $n+4$) and the universe size (the universe will contain $m+1$
individuals in addition to the differently-typed individuals used to encode
the grammar).

\begin{figure}
  \centering
  \includegraphics[width=1\columnwidth]{pic-runtime-modifiers-with-sgplan}
  \caption{Runtimes in the sentence generation domain. The horizontal
  axis represents parameters $(m,n)$ from $(1,1)$ to $(10,10)$ in
  lexicographical order.}
  \label{fig:runtimes-crisp}
\end{figure}

The results of this experiment are shown in
Fig.~\ref{fig:runtimes-crisp}. The input parameters $(m,n)$ are plotted
(in lexicographic order) on the horizontal axis and the runtime is shown in
milliseconds on the vertical axis. These results reveal a number of
interesting insights. First, FF significantly outperforms
SGPLAN in this domain.\footnote{Experiments with SGPLAN use a
 pre-release version of SGPLAN, kindly provided by Chih-Wei Hsu. The
 release version of SGPLAN 5.2.2 had a bug, causing it to crash on some
 instances.}
Second, FF's runtime is dominated by its initial grounding step, in which
it computes the ground instances of all predicates and actions used in the
planning problem to avoid unnecessary instantiations during search. In
particular, the ratio of grounding time to total runtime is generally above
85\%, and rises to above 99\% at $m=11$, which is still a small universe in
this application.\footnote{The
  ``grounding'' time reported here is what FF reports as ``time spent:
  instantiating action templates''.}  This is also reflected in the
much higher sensitivity of FF to the domain size: if we fix $n=1$, FF takes
60 ms to compute a plan at $m=1$, but 2.4 seconds at $m=10$. By contrast,
our naive GraphPlan implementation takes 30 ms at $m=1$ and still only 50
ms at $m=10$.

Our results illustrate that FF is outperformed quite clearly by an ad-hoc
Java implementation of GraphPlan which only computes instances of
predicates and actions as they are discovered, while the planning graph is
being built. On the other hand, the runtime that FF actually spends on
search is again considerably lower than that of our GraphPlan
implementation, and for any given $m$, GraphPlan's runtime grows much
faster with $n$ (i.e., the plan size) than FF's. As a result, for
moderately-sized instances of the sentence planning domain FF does very
well on search, but spends about ten times as much runtime again on
grounding. Larger (but still realistically-sized) instances of the sentence
generation problem are still problematic for the planners we tested.

\subsection{Experiment 2: Minimal GIVE worlds}
\label{sec:exper-2:-minim}

To evaluate the performance of the planners on the planning problems
that arise in the GIVE domain, we construct a sequence of test worlds
as illustrated in Fig.~\ref{fig:give-minimal}.  These worlds consist
of a $2n$ by $h$ grid of positions, such that there are buttons at
positions $(2i-1,1)$ and $(2i,h)$ for $1 \leq i \leq n$.  The player
starts in position $(1,1)$ and must press all buttons in order to
complete the game.  The world is generated as a GIVE world
description, and then converted into a planning problem by the GIVE
software. \todo{at what level of detail should we describe this?}

\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{pic-buttons}
  \caption{Minimal GIVE world.}
  \label{fig:give-minimal}
\end{figure}

Some results from this experiment (for $h=20$, with $n$ ranging from
$1$ to $40$) are shown in Fig.~\ref{fig:give-runtime-minimal}.  The
most obvious result is that FF is unable to solve any problems beyond
$n=13$ on the experimentation machine within the memory limit of 1 GB.
SGPLAN, on the other hand, solves instances beyond $n=40$ without
major problems.  The time spent on grounding is not a major factor in
either planner, probably because the planners need more time to
actually compute the plan -- at $n=40$, the optimal plan has a length
of about 1600!

\begin{figure}
  \centering
  \includegraphics[width=1\columnwidth]{pic-runtime-buttons}
  \caption{Runtimes of FF and SGPLAN on the minimal GIVE worlds for $h=20$. The horizontal axis is $n$, ranging from 1 to 40.}
  \label{fig:give-runtime-minimal}
\end{figure}

\subsection{Experiment 3: GIVE worlds with extra positions}
\label{sec:experiment-3:-give}

Finally, we varied the GIVE world in order to judge the effect of the
universe size in the GIVE domain (see Fig.~\ref{fig:give-junk}).  We
added another $w$ by $h$ positions to the right of a minimal world as
in Experiment 2.  Other than that, we left the initial state and goal
untouched, i.e.\ the new positions were not needed in any plan; this
approximates the situation in the actual GIVE domain, where most grid
positions never need to be used.  As before, we generate a GIVE world
description and then convert it into a planning problem.

\begin{figure}
  \centering
  \includegraphics[width=1\columnwidth]{pic-empty-buttons}
  \caption{GIVE world with extra positions.}
  \label{fig:give-junk}
\end{figure}

Some results for $h=20$ and $n=5$ (with $w$ ranging from $1$ to $70$)
are shown in Fig.~\ref{fig:give-runtime-junk}.  Again, FF runs out of
memory, this time at $w=17$, whereas SGPLAN happily solved inputs
beyond $w=70$.  However, unlike in Experiment 2, both planners now
spend a substantial propertion of their time on grounding again.  In
SGPLAN, this translates to a ``parsing time'' (which
\todo{presumably} includes grounding) which grows from 180 ms to 21.7
seconds as $w$ grows from $1$ to $75$; the rest of the runtime (which
includes the search) only grows from 400 ms to 2.3 seconds.  This
difference is particularly dramatic given that the actual optimal plan
is an identical plan of about 100 steps for each $w$.

\begin{figure}
  \centering
  \includegraphics[width=1\columnwidth]{pic-runtime-empty-world}
  \caption{Runtimes of FF and SGPLAN on the GIVE worlds with junk
    positions for $h=20$ and $n=5$. The horizontal axis is $w$,
    ranging from 1 to 70.}
  \label{fig:give-runtime-junk}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: