### 2008-06-23 trunk/Papers/experiences/domains.tex
I'm essentially happy now

\section{Planning in NLG} \label{sec:domains}

In this section, we present our planning domains: the sentence
generation domain and the instruction giving domain.

\subsection{Sentence generation as planning}

Sentence generation is the problem of computing, from a grammar and a
semantic representation, a single sentence that expresses this piece of
meaning.  This problem is traditionally split into several steps
\cite{reiter00building}.  In a first step, called \emph{sentence planning},
the semantic representation is first enriched with more information; for
instance, \emph{referring expressions}, which refer to individuals
that we want to talk about, are determined at this point. In a second
step, called \emph{surface realization}, this enriched representation is
then translated into a natural-language sentence, using the grammar.

In practice, the process of determining referring expressions
typically interacts with the realization step and so it turns out to
be beneficial to perform both of these steps together. This was the
goal of the SPUD system \cite{Stone2003a}, which did this using a
top-down generation algorithm based on tree-adjoining grammars
\cite{joshi;etal1997} whose lexical entries were equipped with
semantic and pragmatic information. Unfortunately, SPUD suffered from
having to explore a huge search space, and had to resort to a
non-optimal greedy search strategy to retain reasonable efficiency.
To improve on the efficiency of SPUD,
\citeauthor{KolSto07}~\shortcite{KolSto07} translate the sentence
generation problem into a planning problem and use a planning
algorithm for generation.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{pic-grammar}
  \caption{The example grammar.}
  \label{fig:white-rabbit-sleeps-grammar}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{pic-derivation}
  \caption{Derivation of ``The white rabbit sleeps.''}
  \label{fig:white-rabbit-sleeps-deriv}
\end{figure}

We illustrate this process on a simplified example.  Consider a
knowledge base containing the individuals $e$, $r_1$ and $r_2$, and a
set of attributes encoding the fact that $r_1$ and $r_2$ are rabbits,
$r_1$ is white and $r_2$ is brown, and $e$ is an event in which $r_1$
sleeps.  Say that we want to express the information
$\{\mathsf{sleep}(e,r_1)\}$ using the tree-adjoining grammar shown in
Figure~\ref{fig:white-rabbit-sleeps-grammar}. This grammar consists of
\emph{elementary trees} (i.e., the disjoint trees in the figure), each
of which contributes certain \emph{semantic content}. We can
instantiate these trees by substituting individuals for \emph{semantic
  roles}, such as $\mathsf{self}$ and $\mathsf{subj}$, and then
combine the tree instances as shown in
Figure~\ref{fig:white-rabbit-sleeps-deriv} to obtain the sentence
``The white rabbit sleeps''.

We compute this grammatical derivation top-down by starting with the
elementary tree for ``sleeps''. This tree satisfies the need to convey
the semantic information, but introduces a need to generate a noun
phrase (NP) for the subject; this NP must refer uniquely to the target
referent $r_1$. In a second step, we substitute the tree for ``the
rabbit'' into the open NP leaf, which makes the derivation
grammatically complete. Since there are two different individuals that
could be described as ``the rabbit''---technically, $r_2$ is still a
\emph{distractor} (i.e., based on the description ``the rabbit'', the
hearer might erroneously think that we're talking about $r_2$ and not
$r_1$)---we are still not finished. To complete the derivation, the
tree for ``white'' is added to the existing structure by the
\emph{adjunction} operation, making the derivation syntactically and
semantically complete.

\begin{figure}[t]
\centering
{\small%
\begin{verbatim}
(:action add-sleeps
   :parameters (?u - node
                ?xself - individual
                ?xsubj - individual)
   :precondition (and (subst S ?u)
      (referent ?u ?xself)
      (sleep ?xself ?xsubj))
   :effect (and (not (subst S ?u))
      (expressed sleep ?xself ?xsubj)
      (subst NP (subj ?u))
      (referent (subj ?u) ?xsubj)
      (forall (?y - individual)
         (when (not (= ?y ?xself))
            (distractor (subj ?u) ?y)))))

(:action add-rabbit
   :parameters (?u - node
                ?xself - individual)
   :precondition (and (subst NP ?u)
      (referent ?u ?xself)
      (rabbit ?xself))
   :effect (and (not (subst NP ?u))
      (canadjoin N ?u)
      (forall (?y - individual)
         (when (not (rabbit ?y))
           (not (distractor ?u ?y))))))

(:action add-white
   :parameters (?u - node
                ?xself - individual)
   :precondition (and (canadjoin N ?u)
      (referent ?u ?xself)
      (rabbit ?xself))
   :effect (forall (?y - individual)
      (when (not (white ?y))
         (not (distractor ?u ?y)))))
\end{verbatim}}%
\caption{PDDL actions for generating the sentence ``The white rabbit
sleeps.''}
%\caption{Generating ``The white rabbit sleeps'' as a planning problem.}
\label{fig:white-rabbit-as-planning}
\end{figure}

The process described above has clear parallels to planning: we manipulate
a state by applying actions in order to achieve a goal. We can make this
connection even more precise by translating the SPUD problem into a
planning problem. For instance, Figure~\ref{fig:white-rabbit-as-planning}
shows the corresponding PDDL actions for the above generation task, where
each action corresponds to an operation that adds a single elementary tree
to the derivation.  In each case, the first parameter of the action is a
node name in the derivation tree, and the remaining parameters stand for
the individuals to which the semantic roles will be instantiated.  The
syntactic preconditions and effects are encoded using $\mathsf{open}$ and
$\mathsf{canadjoin}$ literals; the status of each referring expression is
tracked using $\mathsf{distractor}$ literals. Notice that the action
effects contain terms of the form $\mathsf{subj}(u)$, which construct new
node names. In order to meet the syntactic requirements of PDDL, these
terms can be eliminated by estimating an upper bound $n$ for the plan
length, making $n$ copies of each action, ensuring that copy $i$ can only
be applied in step $i$, and replacing the term $\mathsf{subj}(u)$ in an
action copy by the constant $\mathsf{subj}_i$.

Once we've defined appropriate actions, we can solve the generation
problem as an ordinary planning problem. For instance, the following plan
solves our previous example:
%
\begin{enumerate}
\item $\mathsf{sleeps}(\mathsf{root}, r_1)$
\item $\mathsf{rabbit}(\mathsf{subj}(\mathsf{root}), r_1)$
\item $\mathsf{white}(\mathsf{subj}(\mathsf{root}), r_1)$
\end{enumerate}
%
Using this plan, the grammatical derivation in
Figure~\ref{fig:white-rabbit-sleeps-deriv}, and therefore the generated
sentence, can be systematically reconstructed. Thus, we can solve the
sentence generation problem via the detour through planning and bring
current search heuristics for planning to bear on generation.

\subsection{Planning in instruction giving}

The object of the GIVE Challenge (``Generating Instructions in Virtual
Environments''; Koller et al.\ \citeyear{alexander07:_shared_task_propos})
is to build an NLG system which is able to produce natural-language
instructions which will guide a human user in performing some task in a
virtual environment. From an NLG perspective, GIVE makes for an
interesting challenge because it is a theory-neutral task that exercises
all components of an NLG system, and emphasizes the study of communication
in a (simulated) physical environment. It also has the advantage that the
user and the NLG system can be physically in different places, as long as
the 3D client and the NLG system are connected over a network. This makes
it possible to evaluate GIVE NLG systems on a large scale over the
Internet. The first GIVE evaluation will take place in late 2008;
currently eight research teams from five countries are working on
developing systems to participate in the challenge.\footnote{See the GIVE
 website for more details about this project:
 \url{homepages.inf.ed.ac.uk/v1akolle/proj/give/}.}

\begin{figure}
\centering
%\includegraphics[width=1 \columnwidth]{give_world_no_expl}
\includegraphics[width=1 \columnwidth]{give_world_2}
\caption{Map of the GIVE development world.}
  \label{fig:give-development-world}
\end{figure}

A map of an example GIVE world is shown in
Figure~\ref{fig:give-development-world}.  In this world, the user's
task is to pick up a trophy in the top left room.  The trophy is
hidden in a safe behind a picture, so the user must first move the
picture out of the way and open the safe by pushing a sequence of
buttons (the small square boxes on the walls) in the correct order.
In order to get to all these buttons, the user must also open the door
in the centre of the map and deactivate the alarm tile by pushing
further buttons.  To simplify both the planning and the NLG task, the
world is discretized into tiles of equal size; the user can turn by 90
degree steps in either direction, and can move from the centre of one
tile to the centre of the next. Figure~\ref{fig:give-planning} shows
some of the available actions in PDDL syntax.

\begin{figure}[t]
{\small%
\begin{verbatim}
(:action move
   :parameters (?from - position
                ?to - position
                ?ori - orientation)
   :precondition (and (player-pos ?from) 
      (adjacent ?from ?to ?ori) 
      (player-orient ?ori)
      (not-blocked ?from ?to)
      (not-alarmed ?to))
   :effect (and (not (player-pos ?from))
      (player-pos ?to)))

(:action turn-left
   :parameters (?ori - orientation
                ?newOri - orientation)
   :precondition (and (player-orient ?ori)
      (next-orient-left ?ori ?newOri))
   :effect (and (not (player-orient ?ori))
      (player-orient ?newOri)))

(:action turn-right
   :parameters (?ori - orientation
                ?newOri - orientation)
   :precondition (and (player-orient ?ori)
      (next-orient-right ?ori ?newOri))
   :effect (and (not (player-orient ?ori))
      (player-orient ?newOri)))

(:action manipulate-b1-off-on
   :parameters (?pos - position)
   :precondition (and (state b1 off)
      (player-pos ?pos)
      (position b1 ?pos))
   :effect (and (not (state b1 off))
      (state b1 on)
      (not (state d1 closed))
      (state d1 open) 
      (not (blocked pos_6_5 pos_6_4))
      (not (blocked pos_6_4 pos_6_5))))
\end{verbatim}}%
\caption{PDDL actions for the GIVE domain.}
\label{fig:give-planning}
\end{figure}

The plans generated in the GIVE domain are often nontrivial. For instance,
in the example above the shortest plan consists of 108 action steps; the
first few steps are as follows:
%
\begin{enumerate}
\item $\mathsf{turn}\textsf{-}\mathsf{left}(\mathsf{north}, \mathsf{west})$
\item $\mathsf{move}(\mathsf{pos\_5\_2}, \mathsf{pos\_4\_2}, \mathsf{west})$
\item $\mathsf{manipulate}\textsf{-}\mathsf{b1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_5\_2})$
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{west}, \mathsf{north})$
\end{enumerate}

Most of the actions in this plan are ``move'' actions.  This makes the
bulk of the GIVE problem very similar to the Gridworld problem, which
also involves finding a route through a world with discrete
positions---but with the additional need to press buttons in the right
order, reason about many more objects in the world, and navigate
somewhat more complicated room shapes.

The GIVE task comes with a number of interesting research challenges
both for the NLG and the planning communities.  On the NLG side, one
major problem is that the natural language instruction sequences can't
simply be obtained by verbalising each action instance in the plan in
turn.  For instance, the NLG system could generate an instruction
sequence starting with ``turn left; walk forward; press the button;
turn right; walk forward; walk forward; turn right; walk forward; walk
forward; turn left; walk forward''.  However, this would be clumsy and
boring, and it is to be expected that frustrated users will quickly
cancel an interaction with such a system, resulting in a bad
evaluation score.  Having just pressed the button, it would be much
better to say ``turn around and walk through the door''; i.e.,
multiple action instances should be merged into a single
instruction. Conversely, it may also be necessary to express a single
planning step with several instructions.  Having just entered the top
left room, it may be easier for the user to understand the instruction
``walk to the centre of the room; turn right; now press the green
button in front of you'' than the instruction ``press the green button
on the wall to your right''.  To plan the referring expression ``the
green button in front of you'' at a time when the button is not yet in
front of the user, the NLG system must keep track of the hypothetical
changes in what objects are visible to the user.  Thus the acts of
referring and instructing, and the modules for planning and NLG, must
be tightly integrated.

From a planning perspective, the use of planning in GIVE imposes very
strict runtime requirements on the planner: Planning must happen in
real time, because if the system takes too much time between deciding
to give an instruction and actually giving this instruction, the user
may have walked or turned away, thus making the instruction invalid.
Furthermore, plan execution monitoring also plays an important role in
the GIVE problem.  At a high level, the system needs to monitor a
user's actions and compare them against the generated instruction set
to determine if the user has correctly followed directions or not. In
the case of the latter, new instructions may have to be generated. In
practice, the situation can be quite complicated since the mental
state of the user is not known and so the system must observe the
user's actions in real time. For example, a user directed to ``turn
around and walk through the door'' may not necessarily perform these
actions to the letter, i.e., immediately turning 180 degrees and
proceeding directly to the door. Instead, the user might take a
roundabout route through the room, eventually exiting out the
door. Although the user's actions do not match the generated
instructions exactly, they meet the intended goal. The system must be
able to identify such ``equivalent'' plans and not immediately
generate new instructions as soon as the user's actions have gone off
course. Furthermore, a user can communicate certain intentions to the
system, both through action and inaction. For instance, the system
should infer that a user has failed to follow instructions if the user
exits a room when given a directive to ``walk to the centre of the
room''. However, the system should also make a similar conclusion if a
user simply does nothing when given the instruction.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: