### 2008-06-23 trunk/Papers/experiences/introduction.tex

\section{Introduction}
\label{sec:introduction}

Natural language generation (NLG; \citeauthor{reiter00building}
\citeyear{reiter00building}) is one of the major subfields of natural
language processing, concerned with computing natural language
sentences or texts that convey a given piece of information to the
user. Intuitively, this task can be viewed as a problem involving
actions, beliefs, and goals: an agent communicating with another agent
tries to change the mental state of the hearer by applying actions
which correspond to the utterance of words or sentences. This
characterization of NLG suggests obvious parallels to automated
planning---a view which has a long tradition in NLG
\cite{appelt:planning,young94dpocl}.  More recently, there has been a
resurgence of interest in applying modern planning techniques to NLG
\cite{Steedman-Petrick:07,KolSto07,benotti08b}.  While efficiency has
not always been the main focus of NLG, problems in NLG are often more
complex than their counterparts in other areas of natural language
processing (such as parsing \cite{KolStr02}), giving rise to
computationally challenging domains. These new approaches combine an
interest in planning as a modelling tool for natural language, with a
hope of improved efficiency by exploiting modern approaches and
algorithms.

The focus of this paper is twofold. We begin by presenting two recent
planning domains that arise in the context of natural language generation:
the sentence generation domain and the GIVE domain. In the sentence
generation task, the goal is to generate a single sentence that expresses a
given meaning. \citeauthor{KolSto07}~\shortcite{KolSto07} cast this domain
as a planning problem, where the plan encodes the necessary sentence and
the actions correspond to uttering individual words. In the GIVE domain
(``Generating Instructions in Virtual Environments''), we describe a new
shared task that was recently posed as a challenge for the NLG community
\cite{alexander07:_shared_task_propos}. GIVE uses planning as one module of
an NLG system that generates natural-language instructions to guide a
user performing a given task in a virtual environment.

We will then discuss some of our experiences using off-the-shelf planners
in these domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN
\cite{hsu06:_new_featur_in_sgplan_for}---planners that have been successful
on traditional benchmarks, and in the International Planning Competition
(IPC)---on a range of problem instances from our planning domains.  We also
compare these results against an ad-hoc Java implementation of GraphPlan
\cite{Blum1997}.  Our findings are mixed. On the one hand, it turns out
that modern planners handle some of the search problems that arise in NLG
quite easily (although large problem instances remain a challenge). On the
other hand, these same planners spend tremendous amounts of time on
preprocessing. For instance, FF spends 90\% of its runtime in the sentence
generation domain on grounding out literals and actions, and ends up slower
than our version of GraphPlan that avoids grounding. For domains like ours,
which are dominated by the number of actions and the universe size, rather
than the combinatorics of the search problem, this observation suggests
that the overall runtime of a planner could be improved by grounding more
selectively. We therefore offer these domains as challenges for the
planning community.

%Natural language generation (NLG; \citeauthor{reiter00building} 2000)
%is one of the major subfields of natural language processing.  It is
%concerned with computing natural language sentences or texts that
%convey a given piece of information to the user.  Problems in NLG are
%often more complex than their counterparts in parsing \cite{KolStr02},
%and indeed efficiency has not traditionally been the main focus of
%NLG.  On the other hand, NLG has some obvious parallels with planning:
%One tries to change the mental state of the hearer by applying actions
%which correspond to uttering words or sentences.  This parallel has a
%long tradition in NLG \cite{appelt:planning,young94dpocl}, and there
%has recently been a resurgence of interest in applying modern planning
%techniques to NLG \cite{Steedman-Petrick:07,KolSto07,benotti08b}.
%These new approaches combine an interest in planning as a modelling
%tool for natural language with a hope for improved efficiency by
%exploiting modern planning algorithms to help with the search.

%In this paper, we present two recent planning domains that arise in the
%context of applying planning to natural language generation.  We will
%first review the sentence generation domain, which was cast as a planning
%problem by \citeauthor{KolSto07}~\shortcite{KolSto07}.  In this domain,
%the goal is to generate a single sentence that expresses a given piece of
%meaning; the plan encodes this sentence, and the actions correspond to
%uttering individual words.  We will then introduce the GIVE domain
%(``Generating Instructions in Virtual Environments'').  GIVE is a new
%shared task that was recently posed as a challenge for the NLG community
%\cite{alexander07:_shared_task_propos}.  It uses planning as one module of
%a NLG system which generates natural-language instructions that will guide
%the user towards performing a given task in a virtual environment.

%We will then discuss some experiences with using off-the-shelf planners in
%these domains.  In particular, we explore the efficiency of FF
%\cite{HoffmannNebel01} and SGPLAN \cite{hsu06:_new_featur_in_sgplan_for} on
%instances of our planning domains of varying sizes.  Our findings are
%mixed.  On the one hand, it turns out that modern planners handle the
%search problems that arise in NLG quite easily.  On the other hand, they
%spend tremendous amounts of time on preprocessing; for instance, FF spends
%90\% of its runtime in the sentence generation domain on grounding out the
%literals and actions, and ends up slower than an ad-hoc Java implementation
%of GraphPlan that avoids grounding.  We argue that for domains like ours,
%which are dominated by the number of actions and the universe size rather
%than the combinatorics of the search problem, the overall runtime of a
%planner could be improved by grounding more selectively.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: