### 2008-07-03 trunk/Papers/rmrs-08/abstract.tex
3rd July

\begin{abstract}
  One way to construct semantic
  representations in a robust manner is to enhance shallow language
  processors with semantic components (e.g., \cite{frank:2004,
    briscoe:etal:2006}).  Here, we provide a model theory for a
  semantic formalism 
  that is designed to enable this,
  namely Robust Minimal Recursion Semantics ({\sc rmrs},
  \cite{copestake:2003}).  We show that {\sc rmrs} supports a notion
  of entailment 
  that has a number of desirable properties: specifically, it allows
  {\sc rmrs} to form the basis for comparing the semantic output of
  different parses of the same string, even when the parses vary in
  the depth of their analysis, and in particular to test the
  compatibility of the semantic output of a shallow parser with that
  of a deep parser (where such deep parses exist).
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rmrs-08"
%%% End: