### 2008-07-20 trunk/Papers/experiences/domains.tex

\section{Planning in NLG} \label{sec:domains}

We begin by presenting our two planning domains: the sentence generation
domain and the instruction giving domain.

\subsection{Sentence generation as planning}

Sentence generation is the problem of computing, from a grammar and a
semantic representation, a single sentence that expresses this piece of
meaning.  This problem is traditionally split into several steps
\cite{reiter00building}.  In a first step, called \emph{sentence planning},
the semantic representation is first enriched with more information; for
instance, \emph{referring expressions}, which refer to individuals
that we want to talk about, are determined at this point. In a second
step, called \emph{surface realization}, this enriched representation is
then translated into a natural-language sentence, using the grammar.

In practice, the process of determining referring expressions typically
interacts with the realization step and so it turns out to be beneficial to
perform both of these steps together. This was the goal of the SPUD system
\cite{Stone2003a}, which performed this task using a top-down generation
algorithm based on tree-adjoining grammars \cite{joshi;etal1997} whose
lexical entries were equipped with semantic and pragmatic information.
Unfortunately, SPUD suffered from having to explore a huge search space,
and had to resort to a non-optimal greedy search strategy to retain
reasonable efficiency.  To improve on the efficiency of SPUD,
\citeauthor{KolSto07}~\shortcite{KolSto07} translate the sentence
generation problem into a planning problem and use a planning algorithm for
generation.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{pic-grammar}
  \caption{The example grammar.}
  \label{fig:white-rabbit-sleeps-grammar}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{pic-derivation}
  \caption{Derivation of ``The white rabbit sleeps.''}
  \label{fig:white-rabbit-sleeps-deriv}
\end{figure}

We illustrate this process on a simplified example.  Consider a
knowledge base containing the individuals $e$, $r_1$ and $r_2$, and a
set of attributes encoding the fact that $r_1$ and $r_2$ are rabbits,
$r_1$ is white and $r_2$ is brown, and $e$ is an event in which $r_1$
sleeps.  Say that we want to express the information
$\{\mathsf{sleep}(e,r_1)\}$ using the tree-adjoining grammar shown in
Figure~\ref{fig:white-rabbit-sleeps-grammar}. This grammar consists of
\emph{elementary trees} (i.e., the disjoint trees in the figure), each
of which contributes certain \emph{semantic content}. We can
instantiate these trees by substituting individuals for \emph{semantic
  roles}, such as $\mathsf{self}$ and $\mathsf{subj}$, and then
combine the tree instances as shown in
Figure~\ref{fig:white-rabbit-sleeps-deriv} to obtain the sentence
``The white rabbit sleeps''.

We compute this grammatical derivation in a top-down manner, starting with
the elementary tree for ``sleeps''. This tree satisfies the need to convey
the semantic information, but introduces a need to generate a noun phrase
(NP) for the subject; this NP must refer uniquely to the target referent
$r_1$. In a second step, we substitute the tree for ``the rabbit'' into the
open NP leaf, which makes the derivation grammatically complete. Since
there are two different individuals that could be described as ``the
rabbit''---technically, $r_2$ is still a \emph{distractor} (i.e., based on
the description ``the rabbit'', the hearer might erroneously think that
we're talking about $r_2$ and not $r_1$)---we are still not finished. To
complete the derivation, the tree for ``white'' is added to the existing
structure by an \emph{adjunction} operation, making the derivation
syntactically and semantically complete.

\begin{figure}[t]
\centering
{\small%
\begin{verbatim}
(:action add-sleeps
   :parameters (?u - node
                ?xself - individual
                ?xsubj - individual)
   :precondition (and (subst S ?u)
      (referent ?u ?xself)
      (sleep ?xself ?xsubj))
   :effect (and (not (subst S ?u))
      (expressed sleep ?xself ?xsubj)
      (subst NP (subj ?u))
      (referent (subj ?u) ?xsubj)
      (forall (?y - individual)
         (when (not (= ?y ?xself))
            (distractor (subj ?u) ?y)))))

(:action add-rabbit
   :parameters (?u - node
                ?xself - individual)
   :precondition (and (subst NP ?u)
      (referent ?u ?xself)
      (rabbit ?xself))
   :effect (and (not (subst NP ?u))
      (canadjoin N ?u)
      (forall (?y - individual)
         (when (not (rabbit ?y))
           (not (distractor ?u ?y))))))

(:action add-white
   :parameters (?u - node
                ?xself - individual)
   :precondition (and (canadjoin N ?u)
      (referent ?u ?xself)
      (rabbit ?xself))
   :effect (forall (?y - individual)
      (when (not (white ?y))
         (not (distractor ?u ?y)))))
\end{verbatim}}%
\caption{PDDL actions for generating the sentence ``The white rabbit
sleeps.''}
%\caption{Generating ``The white rabbit sleeps'' as a planning problem.}
\label{fig:white-rabbit-as-planning}
\end{figure}

The process described above has clear parallels to planning: we
manipulate a state by applying actions in order to achieve a goal. We
can make this connection even more precise by translating the SPUD
problem into a planning problem.\footnote{See
  \url{http://code.google.com/p/crisp-nlg/} for the CRISP system, in
  which this conversion is implemented.}  For instance,
Figure~\ref{fig:white-rabbit-as-planning} shows the corresponding PDDL
actions for the above generation task, where each action corresponds
to an operation that adds a single elementary tree to the derivation.
In each case, the first parameter of the action is a node name in the
derivation tree, and the remaining parameters stand for the
individuals to which the semantic roles will be instantiated.  The
syntactic preconditions and effects are encoded using $\mathsf{open}$
and $\mathsf{canadjoin}$ literals; the status of each referring
expression is tracked using $\mathsf{distractor}$ literals. Notice
that the action effects contain terms of the form $\mathsf{subj}(u)$,
which construct new node names. In order to meet the syntactic
requirements of PDDL, these terms can be eliminated by estimating an
upper bound $n$ for the plan length, making $n$ copies of each action,
ensuring that copy $i$ can only be applied in step $i$, and replacing
the term $\mathsf{subj}(u)$ in an action copy by the constant
$\mathsf{subj}_i$.

Once the appropriate actions are defined, we can solve the generation
problem as an ordinary planning problem. For instance, the following plan
solves our previous example:
%
\begin{enumerate}
\item $\mathsf{sleeps}(\mathsf{root}, r_1)$
\item $\mathsf{rabbit}(\mathsf{subj}(\mathsf{root}), r_1)$
\item $\mathsf{white}(\mathsf{subj}(\mathsf{root}), r_1)$
\end{enumerate}
%
Using this plan, the grammatical derivation in
Figure~\ref{fig:white-rabbit-sleeps-deriv}, and therefore the
generated sentence ``the white rabbit sleeps'', can be systematically
reconstructed. Thus, we can solve the sentence generation problem via
the detour through planning and bring current search heuristics for
planning to bear on generation.

\subsection{Planning in instruction giving}

The object of the GIVE Challenge (``Generating Instructions in Virtual
Environments''; Koller et al.\
\citeyear{alexander07:_shared_task_propos}) is to build an NLG system
which is able to produce natural-language instructions which will
guide a human user in performing some task in a virtual
environment. From an NLG perspective, GIVE makes for an interesting
challenge because it is a theory-neutral task that exercises all
components of an NLG system, and emphasizes the study of communication
in a (simulated) physical environment. It also has the advantage that
the user and the NLG system can be in physically different places, as
long as the 3D client and the NLG system are connected over a
network. This makes it possible to evaluate GIVE NLG systems on a
large scale over the Internet. The first GIVE evaluation will take
place in late 2008; currently eight research teams from five countries
are working on developing systems to participate in the
challenge.\footnote{See the GIVE website for more details about this
  project: \url{http://www.give-challenge.org/}.}

\begin{figure}
\centering
%\includegraphics[width=1 \columnwidth]{give_world_no_expl}
\includegraphics[width=1 \columnwidth]{give_world_2}
\caption{Map of the GIVE development world.}
  \label{fig:give-development-world}
\end{figure}

A map of an example GIVE world is shown in
Figure~\ref{fig:give-development-world}.  In this world, the user's
task is to pick up a trophy in the top left room.  The trophy is
hidden in a safe behind a picture, so the user must first move the
picture out of the way and open the safe by pushing a sequence of
buttons (the small square boxes on the walls) in the correct order.
In order to get to all these buttons, the user must also open the door
in the centre of the map and deactivate the alarm tile by pushing
further buttons.  To simplify both the planning and the NLG task, the
world is discretized into tiles of equal size; the user can turn by 90
degree steps in either direction, and can move from the centre of one
tile to the centre of the next. Figure~\ref{fig:give-planning} shows
some of the available actions in PDDL syntax.

\begin{figure}[t]
{\small%
\begin{verbatim}
(:action move
   :parameters (?from - position
                ?to - position
                ?ori - orientation)
   :precondition (and (player-pos ?from) 
      (adjacent ?from ?to ?ori) 
      (player-orient ?ori)
      (not-blocked ?from ?to)
      (not-alarmed ?to))
   :effect (and (not (player-pos ?from))
      (player-pos ?to)))

(:action turn-left
   :parameters (?ori - orientation
                ?newOri - orientation)
   :precondition (and (player-orient ?ori)
      (next-orient-left ?ori ?newOri))
   :effect (and (not (player-orient ?ori))
      (player-orient ?newOri)))

(:action turn-right
   :parameters (?ori - orientation
                ?newOri - orientation)
   :precondition (and (player-orient ?ori)
      (next-orient-right ?ori ?newOri))
   :effect (and (not (player-orient ?ori))
      (player-orient ?newOri)))

(:action manipulate-b1-off-on
   :parameters (?pos - position)
   :precondition (and (state b1 off)
      (player-pos ?pos)
      (position b1 ?pos))
   :effect (and (not (state b1 off))
      (state b1 on)
      (not (state d1 closed))
      (state d1 open) 
      (not (blocked pos_6_5 pos_6_4))
      (not (blocked pos_6_4 pos_6_5))))
\end{verbatim}}%
\caption{PDDL actions for the GIVE domain.}
\label{fig:give-planning}
\end{figure}

The plans generated in the GIVE domain are often nontrivial. For instance,
in the example above the shortest plan consists of 108 action steps; the
first few steps are as follows:
%
\begin{enumerate}
\item $\mathsf{turn}\textsf{-}\mathsf{left}(\mathsf{north}, \mathsf{west})$
\item $\mathsf{move}(\mathsf{pos\_5\_2}, \mathsf{pos\_4\_2}, \mathsf{west})$
\item $\mathsf{manipulate}\textsf{-}\mathsf{b1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_5\_2})$
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{west}, \mathsf{north})$
\end{enumerate}

Most of the actions in this plan are ``move'' actions.  This makes the
bulk of the GIVE problem very similar to the Gridworld problem, which
also involves finding a route through a world with discrete
positions---but with the additional need to press buttons in the right
order, reason about many more objects in the world, and navigate
somewhat more complicated room shapes.

The GIVE task comes with a number of interesting research challenges
both for the NLG and the planning communities.  On the NLG side, one
major problem is that the natural language instruction sequences can't
simply be obtained by verbalising each action instance in a plan in
turn.  For instance, the NLG system could generate an instruction
sequence starting with ``turn left; walk forward; press the button;
turn right; walk forward; walk forward; turn right; walk forward; walk
forward; turn left; walk forward''.  However, this would be clumsy and
boring, and it is to be expected that frustrated users will quickly
cancel an interaction with such a system, resulting in a bad
evaluation score.  Having just pressed the button, it would be much
better to say ``turn around and walk through the door''. That is,
multiple action instances should be merged into a single instruction;
plans must be \emph{summarised} in order to present them to the
user. Conversely, it may also be necessary to express a single
planning step with several instructions.  Having just entered the top
left room, it may be easier for the user to understand the instruction
``walk to the centre of the room; turn right; now press the green
button in front of you'' than the instruction ``press the green button
on the wall to your right''.  To plan the referring expression ``the
green button in front of you'' at a time when the button is not yet in
front of the user, the NLG system must keep track of the hypothetical
changes in what objects are visible to the user.  Thus the acts of
referring and instructing, and the modules for planning and NLG, must
be tightly integrated.

From a planning perspective, GIVE imposes very strict runtime requirements
on the planner and its associated modules: planning must happen in real
time and the system must respond to a user in a timely fashion.  If the
system takes too much time deliberating over an instruction to give, rather
than actually giving this instruction, the user may have walked or turned
away, thus making the instruction invalid.  Furthermore, plan execution
monitoring also plays an important role in the GIVE problem.  At a high
level, the system needs to monitor a user's actions and compare them
against the generated instruction set to determine if the user has
correctly followed directions or not. In the case of the latter, new
instructions may have to be generated. In practice, the situation can be
quite complicated since the mental state of the user is not known and so
the system must observe the user's actions in real time. For example, a
user directed to ``turn around and walk through the door'' may not
necessarily perform these actions to the letter, i.e., immediately turning
180 degrees and proceeding directly to the door. Instead, the user might
take a roundabout route through the room, eventually exiting out the door.
Although the user's actions do not match the generated instructions
exactly, they meet the intended goal. The system must be able to identify
such ``equivalent'' plans and not immediately generate new instructions as
soon as the user's actions have gone off course. Furthermore, a user can
communicate certain intentions to the system, both through action and
inaction. For instance, the system should infer that a user has failed to
follow instructions if the user exits a room when given a directive to
``walk to the centre of the room''. The system should also make a
similar conclusion if a user simply does nothing when given the
instruction.

%between deciding to give an instruction and

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: