### 2008-07-24 trunk/Papers/experiences/conclusion.tex
new graphs; small changes to acknowledgements

\section{Conclusion} \label{sec:conclusion}

In this paper, we introduced two novel planning domains arising from
problems in natural language generation: the sentence generation
domain and the GIVE navigation domain. We also reported on the results
of a number of experiments in which we applied off-the-shelf planners
to these domains.

Our results were mixed. While modern planners do a pretty good job of
controlling the complexity of search, they also suffer from practical
problems that limit their performance in unexpected ways. In
particular, in both domains the grounding step performed by both
planners dominates the time it takes to perform the search itself. In
future work, it would be interesting to explore planners such as SHOP2
\cite{DBLP:journals/jair/NauAIKMWY03}, which do not perform the
initial grounding step.  We are also considering to propose some of
our domains as IPC benchmarks.

The NLG community's recent interest in planning presents a valuable
opportunity for planning researchers: provided some of the challenges
we have highlighted can be addressed, projects like GIVE (and other
NLG-inspired problem domains) offer a constructive platform for the
planning community to showcase their techniques to a wider
audience---and to improve the quality of their tools for real-world
planning tasks.  Furthermore, GIVE highlights planning-related
challenges such as plan execution monitoring and plan presentation
and summarisation and offers a platform on which such technologies can
be evaluated in experiments with human users.

\section{Acknowledgements}
This work arose in the context of the Planning and Language Interest Group
at the University of Edinburgh. We thank all members of this group,
especially Hector Geffner and Mark Steedman, for interesting discussions.
This work was partially supported by the DFG Research Fellowship ``CRISP:
Efficient integrated realization and microplanning'' and by the European
Commission through the PACO-PLUS project (FP6-2004-IST-4-27657).

%\paragraph{Acknowledgments.} This work is a collaboration that arose
%in the context of the ``planning and language'' group at the
%University of Edinburgh.  We thank all members of this group,
%especially Hector Geffner and Mark Steedman, for interesting
%discussions.  This work was supported by the DFG Research Fellowship
%``CRISP: Efficient integrated realization and microplanning'' and the
%EU project XYZ PACO-PLUS.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: