### 2008-07-25 trunk/Papers/experiences/discussion.tex

\section{Discussion} \label{sec:discussion}

The positive conclusion we can draw from the simple experiments
reported above is that modern planners are fairly good at controlling
the search for many of the moderately-sized NLG problems we looked
at. This is particularly true for FF in the sentence generation
domain, which generates nontrivial 14-word sentences in about two seconds.
There is still room for improvement, however: special-purpose
algorithms generate much larger referring expressions in
milliseconds \cite{AreKolStr08}.  Although search efficiency is not the
main concern in these domains, it nevertheless becomes a factor in
scenarios like Experiment~2, where the world is almost completely
unconstrained and (unsurprisingly) the search becomes much slower.

The most restrictive bottleneck in the two domains we investigated is
the initial grounding step that many modern planners perform. While
this may be a good strategy for traditional planning benchmarks and
IPC domains---where plan length often dominates universe size, and an
initial investment into grounding pays off in saved instantiations
during the search---this approach is less effective in the domains
we've considered. As our experiments show, there are natural
planning domains in which relatively short plans must be computed in
large universes. For instance, in the case of the sentence generation
domain, it is not unrealistic to look for plans of length 20 over a
universe consisting of thousands of domain individuals and tens of
thousands of actions, some of which require three domain individuals
as parameters. The common strategy of splitting the universe over
several types of individuals will not be very effective here.

The recent trend in planning research has (rightfully) focused on the
development of algorithms that control search in sophisticated ways,
resulting in a host of planners that are more powerful and more
successful than their predecessors. However, the common strategy of
grounding out sets of predicates and actions in implementations of these
algorithms has a very pronounced effect on domains such as those described
above. More sophisticated techniques based on reachability analysis may
have the potential to help improve this situation, provided such
approaches can avoid the drawbacks resulting from unnecessary grounding.
Since our domains are not that unusual in their structure and composition,
we hope that the lessons learnt from our experiences can help improve the
performance of current systems.  We believe that such improvements are
necessary if planning is to become a more mature technology that can offer
tools to a wider community of users.  At the end of the day, real-world
users will care about the \emph{total} runtime of a planner, and this is
more than just the search time.
%has the potential to help---we could gain a lot by only computing
%reachable instances---, but this too is typically done on the grounded
%problem.  \todo{check if this is true}

By and large, our experiences with the planning community from the point of
view of a ``customer'' (one of the authors is not a planning researcher)
have been relatively pleasant. Thanks to the planning competitions, it is
easy to identify and download a fast implementation, at least for Linux.
However, in the course of our experiments we found (and reported) bugs in
both SGPLAN and FF. We also discovered that deciding between the range of
available planners is not always straightforward. As our experiments have
shown, even planners as closely related as FF and SGPLAN can differ
significantly in their performance on different domains.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End: