### 2008-09-05 trunk/Papers/rmrs-08/abstract.tex
language more concise and partial conclusion added

\begin{abstract}
  One way to construct semantic
  representations in a robust manner is to enhance shallow language
  processors with semantic components.  Here, we provide a model
  theory for a 
  semantic formalism 
  that is designed for this,
  namely Robust Minimal Recursion Semantics ({\sc rmrs},
  \cite{copestake:2003}).  We show that {\sc rmrs} supports a notion
  of entailment 
  that allows it
  to form the basis for comparing the semantic output of
  different parses of varying depth.
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rmrs-08"
%%% End: