### 2008-09-05 trunk/Papers/rmrs-08/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper we have motivated and described \rmrs---a semantic
framework that can be used to specify
semantic components to robust shallow parsers.  Accordingly, \rmrs\
is designed to be
maximally flexible on the type of semantic information
that can be left underspecified, so that the semantic output of a
shallow parser needn't over-determine or under-determine the semantic
information that can be extracted from the shallow syntactic analysis.

We defined its syntax and model-theoretic semantics, and following the
tradition from dominance constraints, we defined a syntactic criterion
for characterising entailment.  The syntactic notion of {\em solved
  forms} is useful because it means that even when an underspecified
\rmrs\ corresonds to an infinite number of possible fully-specific
interpretations in context, there is a finite way of identifying all
the `relevant' differences among those fully-specific interpretations.
We showed that the model theory supports theorems that allow one to
use \rmrs\ representations as a way of comparing and even integrating
the semantic output of different parses of the same natural language
string, even when they vary in depth.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rmrs-08"
%%% End: