### 2008-10-06 trunk/Papers/rmrs-08/conclusion.tex
Monday update

\section{Conclusion}
\label{sec:conclusion}

In this paper, we have motivated and defined \rmrs---a semantic
framework that has been used to represent, compare, and combine
semantic information computed from deep and shallow parsers.  \rmrs\
is designed to be maximally flexible on the type of semantic
information that can be left underspecified, so that
the semantic output of a shallow parser needn't over-determine or
under-determine the semantics that can be extracted from
the shallow syntactic analysis.  This lays the formal foundations for
a formalism that is emerging as a standard in
robust semantic processing.

We then defined the concept of \emph{solved forms} of an \rmrs, in
analogy to dominance graphs.  Solved forms abstract away from
irrelevant differences between models, and have been a crucial concept
in the development of efficient solvers for dominance graphs.  In the
future, we hope to draw upon that body of literature to develop
efficient algorithms for processing \rmrs s, including solvers which
will compute fully specific semantic representations from an \rmrs,
efficient inference algorithms, and algorithms for entailment
checking.  We have argued that entailment is a central notion in
comparing the outputs of different {\sc nlp} tools, and provided a syntactic
characterisation for \rmrs\ entailment, which we hope will allow us to
develop efficient entailment algorithms in the future.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rmrs-08"
%%% End: