### 2008-10-08 trunk/Papers/rmrs-08/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper, we have motivated and defined \rmrs---a semantic
framework that has been used to represent, compare, and combine
semantic information computed from deep and shallow parsers.  \rmrs\
is designed to be maximally flexible on the type of semantic
information that can be left underspecified, so that the semantic
output of a shallow parser needn't over-determine or under-determine
the semantics that can be extracted from the shallow syntactic
analysis.  The key contribution of this paper was to lay the formal
foundations for a formalism that is emerging as a standard in robust
semantic processing.

Although we have not directly provided new tools for modelling or
processing language, we believe that a cleanly defined model theory
for \rmrs\ is a crucial prerequisite for the future development of
such tools.  We hope that future research will build upon this paper
to develop efficient algorithms and implementations for solving \rmrs
s, performing inferences that enrich \rmrs s from shallow analyses
with deeper information, and checking consistency of \rmrs s that were
obtained from different parsers.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rmrs-08"
%%% End: