### 2009-04-27 trunk/Papers/ci-spark/manuscript.tex
added Ron's CI submission

%%%
%%% Computational Intelligence manuscript based on SPARK'08 paper
%%%
%%% Alexander Koller and Ronald Petrick
%%% 
%%% Updated: 2008/12/22
%%%
\documentclass[letterpaper]{article}
\usepackage{times}
\usepackage{graphicx}
\usepackage{url}
\usepackage[round]{natbib}
\usepackage{setspace}
\usepackage[paper=letterpaper,margin=1.0in]{geometry}

\title{Experiences with Planning for Natural Language Generation}
\author{\textbf{Alexander Koller} \\
Exzellenzcluster / FR 4.7 Computerlinguistik \\
Saarland University \\
Postfach 151150 \\
66041 Saarbrücken \\
Germany \\
Phone: +49 681 302 70040 \\
Fax: +49 681 302 4351 \\
E-mail: \url{koller@mmci.uni-saarland.de}
\and
\textbf{Ronald P. A. Petrick} \\
School of Informatics \\
University of Edinburgh \\
Informatics Forum, 10 Crichton Street \\
Edinburgh \ EH8 9AB \\
Scotland, United Kingdom \\
Phone: +44 131 650 4426 \\
Fax: +44 131 650 6899 \\
E-mail: \url{rpetrick@inf.ed.ac.uk}}

%\date{22 December 2008}
\date{}

\pagestyle{plain}
\pagenumbering{arabic}
\doublespacing

\bibpunct{(}{)}{;}{a}{}{,}

\begin{document}

\maketitle

%%%
%%% Abstract
%%%
\pagebreak
\begin{abstract}
We investigate the application of modern planning techniques to domains
arising from problems in natural language generation (NLG). In particular,
we consider two novel NLG-inspired planning problems, the sentence
generation domain and the GIVE (``Generating Instructions in Virtual
Environment'') domain, and investigate the efficiency of the FF and SGPLAN
planners in these domains. We also compare our results against an ad-hoc
implementation of GraphPlan in Java. Our results are mixed. While modern
planners are able to quickly solve many moderately-sized instances of our
problems, the overall planning time is dominated by the grounding step that
these planners perform, rather than search. This has a pronounced effect on
our domains which require relatively short plans but have large universes.
We share our experiences and offer these domains as challenges for the
planning community.
\end{abstract}

\bigskip\noindent
\textbf{Keywords:} natural language generation, planning

%%%
%%% Main text
%%%
\pagebreak
\section{Introduction}
\label{sec:introduction}

Natural language generation (NLG; \citealp{reiter00building}) is one of the
major subfields of natural language processing, concerned with computing
natural language sentences or texts that convey a given piece of
information to the user. Intuitively, this task can be viewed as a problem
involving actions, beliefs, and goals: an agent communicating with another
agent tries to change the mental state of the hearer by applying actions
which correspond to the utterance of words or sentences.
This characterization of NLG suggests obvious parallels to automated
planning---a view which has a long tradition in NLG.

Early approaches to generation as planning
\citep{perrault80,appelt:planning,hovy88,young94dpocl} focused primarily on
planning high-level structures, such as speech acts and discourse
relations.  While these approaches suffered from the inefficiency of the
planners that were available at the time, there has recently been a
resurgence of interest in applying modern planning techniques to NLG.
These newer papers revisit speech act planning based on the more expressive
planning formalisms available today \citep{Steedman-Petrick:07,benotti08b}
and exploit the improved efficiency of modern planners for an integrated
model of sentence planning and surface realization \citep{KolSto07}.
Problems in NLG are often computationally challenging (see, e.g.,
\citep{KolStr02}), and these new approaches combine an interest in planning
as a modelling tool for natural language, with a hope of improved
efficiency by exploiting modern approaches and algorithms.

The focus of this paper is twofold. First, we present two novel planning
domains that arise in the context of natural language generation: the
sentence generation domain and the GIVE domain. In the sentence generation
task, the goal is to generate a single sentence that expresses a given
meaning. \citet{KolSto07} cast this domain as a planning problem, where the
plan encodes the necessary sentence and the actions correspond to uttering
individual words.  In the GIVE domain (``Generating Instructions in Virtual
Environments''), we describe a new shared task that was recently posed as a
challenge for the NLG community \citep{alexander07:_shared_task_propos}.
GIVE uses planning as part of an NLG system that generates natural-language
instructions to guide a user in performing a given task in a virtual
environment.

Second, we discuss some of our experiences using off-the-shelf
planners in our two NLG-inspired domains. In particular, we explore the
efficiency of FF \citep{HoffmannNebel01} and SGPLAN
\citep{hsu06:_new_featur_in_sgplan_for}: two planners that are readily
available, support an expressive subset of the Planning Domain Definition
Language (PDDL; McDermott et al.~\citeyear{PDDL}) for encoding domains, and
have been successful on traditional benchmarks and problems from the
International Planning Competition (IPC)\footnote{See
 \url{http://ipc.icaps-conference.org/} for information about past
 editions of the IPC.}.
We test these planners on a range of problem instances in our planning
domains, and compare our results against an ad-hoc Java implementation of
GraphPlan \citep{Blum1997}. Overall, our findings are mixed. On the one
hand, we demonstrate that modern planners can easily handle some of the
moderately-sized search problems that arise in NLG. In certain cases, the
planners we tested are more efficient than specialized NLG algorithms
applied to the same problem. On the other hand, these same planners spend
tremendous amounts of time on preprocessing. For instance, in the sentence
generation domain, FF spends 90\% of its runtime grounding out literals and
actions, and ends up being slower than our version of GraphPlan that avoids
grounding. For domains like ours, which are dominated by the number of
actions and the universe size, rather than the combinatorics of the search
problem, this observation suggests that the overall runtime of a planner
could be improved by grounding more selectively. This is particularly true
for large problem instances, which still remain a challenge for the
planners we tested, and where the effects of the grounding approach become
quite pronounced. We therefore offer our NLG domains as challenges for the
planning community.

\section{Planning in NLG}
\label{sec:domains}

We begin by presenting our two planning domains: the sentence generation
domain and the instruction giving domain.

\subsection{Sentence generation as planning}

Sentence generation is the problem of computing from a grammar and a
semantic representation, a single sentence that expresses this piece of
meaning. Traditionally, this problem is split into several steps
\citep{reiter00building}. In a first step, called \emph{sentence planning},
the semantic representation is enriched with additional information. For
instance, \emph{referring expressions}, which refer to individuals (e.g.,
objects, people, etc.) in the domain of discourse, are determined at this
point. In a second step, called \emph{surface realization}, this enriched
representation is translated into a natural-language sentence, using the
grammar.

In practice, the process of determining referring expressions typically
interacts with the realization step and so it often turns out to be
beneficial to perform these two steps together. For instance, the SPUD
system (``Sentence Planning Using Description''; \citealt{Stone2003a})
performed this task using a top-down generation algorithm based on
tree-adjoining grammars \citep{joshi;etal1997} whose lexical entries were
equipped with semantic and pragmatic information. Unfortunately, SPUD had
to resort to a non-optimal greedy search strategy to cope with its large
search space and retain reasonable efficiency. To improve the efficiency
of SPUD, \citet{KolSto07} translate the sentence generation problem into a
planning problem and use a planning algorithm to perform the generation
task.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.75\columnwidth]{pic-grammar}
  \caption{An example grammar in the sentence generation domain.}
  \label{fig:white-rabbit-sleeps-grammar}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.75\columnwidth]{pic-derivation-2}
  \caption{Derivation of ``The white rabbit sleeps.''}
  \label{fig:white-rabbit-sleeps-deriv}
\end{figure}

We illustrate this process on a simplified example.  Consider a
knowledge base containing the individuals $e$, $r_1$ and $r_2$, and a
set of attributes encoding the fact that $r_1$ and $r_2$ are rabbits,
$r_1$ is white and $r_2$ is brown, and $e$ is an event in which $r_1$
sleeps.  Say that we want to express the information
$\{\mathsf{sleep}(e,r_1)\}$ using the tree-adjoining grammar shown in
Figure~\ref{fig:white-rabbit-sleeps-grammar}. This grammar consists of
\emph{elementary trees} (i.e., the disjoint trees in the figure), each
of which contributes certain \emph{semantic content}. We can
instantiate these trees by substituting individuals for \emph{semantic
  roles}, such as $\mathsf{self}$ and $\mathsf{subj}$, and then
combine the tree instances as shown in
Figure~\ref{fig:white-rabbit-sleeps-deriv} to obtain the sentence
``The white rabbit sleeps''.

We compute this grammatical derivation in a top-down manner, starting with
the elementary tree for ``sleeps''. This tree satisfies the need to convey
the semantic information, but introduces a need to generate a noun phrase
(NP) for the subject; this NP must refer uniquely to the target referent
$r_1$. In a second step, we substitute the tree for ``the rabbit'' into the
open NP leaf, which makes the derivation grammatically complete. Since
there are two different individuals that could be described as ``the
rabbit''---technically, $r_2$ is still a \emph{distractor} (i.e., based on
the description ``the rabbit'', the hearer might erroneously think that
we're talking about $r_2$ and not $r_1$)---we are still not finished. To
complete the derivation, the tree for ``white'' is added to the existing
structure by an \emph{adjunction} operation, making the derivation
both syntactically and semantically complete.

\begin{figure}[p]
\centering
\begin{minipage}{0.5\textwidth}
{\small%
\begin{verbatim}
(:action add-sleeps
   :parameters (?u - node
                ?xself - individual
                ?xsubj - individual)
   :precondition
       (and (subst S ?u)
            (referent ?u ?xself)
            (sleep ?xself ?xsubj))
   :effect 
       (and (not (subst S ?u))
            (expressed sleep ?xself ?xsubj)
            (subst NP (subj ?u))
            (referent (subj ?u) ?xsubj)
            (forall (?y - individual)
                (when (not (= ?y ?xself))
                    (distractor (subj ?u) ?y)))))

(:action add-rabbit
   :parameters (?u - node
                ?xself - individual)
   :precondition 
       (and (subst NP ?u)
            (referent ?u ?xself)
            (rabbit ?xself))
   :effect 
       (and (not (subst NP ?u))
            (canadjoin N ?u)
            (forall (?y - individual)
                (when (not (rabbit ?y))
                    (not (distractor ?u ?y))))))

(:action add-white
   :parameters (?u - node
                ?xself - individual)
   :precondition 
       (and (canadjoin N ?u)
            (referent ?u ?xself)
            (rabbit ?xself))
   :effect 
       (forall (?y - individual)
           (when (not (white ?y))
               (not (distractor ?u ?y)))))
\end{verbatim}}%
\end{minipage}
\caption{PDDL actions for generating the sentence ``The white rabbit
sleeps.''}
\label{fig:white-rabbit-as-planning}
\end{figure}

The process described above has clear parallels to planning: we manipulate
states by applying sequences of actions to achieve a goal. We can make this
connection even more precise by translating the SPUD problem into a
planning problem.\footnote{See
  \url{http://code.google.com/p/crisp-nlg/} for the CRISP system, in
  which this conversion is implemented.}  For instance,
Figure~\ref{fig:white-rabbit-as-planning} shows the corresponding PDDL
actions for the above generation task, where each action corresponds
to an operation that adds a single elementary tree to the derivation.
In each case, the first parameter of the action is a node name in the
derivation tree, and the remaining parameters stand for the
individuals to which the semantic roles will be instantiated.  The
syntactic preconditions and effects are encoded using $\mathsf{open}$
and $\mathsf{canadjoin}$ literals; the status of each referring expression
is tracked using $\mathsf{distractor}$ literals. The resulting action
effects contain terms of the form $\mathsf{subj}(u)$, which construct new
node names, but aren't strictly permitted by PDDL. In order to meet the
syntactic requirements of PDDL, these terms can be eliminated by estimating
an upper bound $n$ for the plan length, making $n$ copies of each action,
ensuring that copy $i$ can only be applied in step $i$, and replacing the
term $\mathsf{subj}(u)$ in an action copy by the constant
$\mathsf{subj}_i$.

Once the appropriate actions are defined, we can solve the generation
problem as an ordinary planning problem. For instance, the following plan
solves our previous example:
%
\begin{enumerate}
\item $\mathsf{add}\textsf{-}\mathsf{sleeps}(\mathsf{root}, r_1)$,
\item $\mathsf{add}\textsf{-}\mathsf{rabbit}(\mathsf{subj}(\mathsf{root}),r_1)$,
\item $\mathsf{add}\textsf{-}\mathsf{white}(\mathsf{subj}(\mathsf{root}),r_1)$.
\end{enumerate}
%
Using this plan, the grammatical derivation in
Figure~\ref{fig:white-rabbit-sleeps-deriv}, and therefore the
generated sentence ``the white rabbit sleeps'', can be systematically
reconstructed. Thus, we can solve the sentence generation problem via
the detour through planning and bring current search heuristics for
planning to bear on generation.

\subsection{Planning in instruction giving}

The object of the GIVE Challenge (``Generating Instructions in Virtual
Environments''; \citealt{alexander07:_shared_task_propos}) is to build an
NLG system which is able to produce natural-language instructions which
will guide a human user in performing some task in a virtual environment.
From an NLG perspective, GIVE makes for an interesting challenge because it
is a theory-neutral task that exercises all components of an NLG system,
and emphasizes the study of communication in a (simulated) physical
environment. It also has the advantage that the user and the NLG system can
be in physically different places, as long as the 3D client and the NLG
system are connected over a network. This makes it possible to evaluate
GIVE NLG systems on a large scale over the Internet.
%
The first GIVE evaluation is currently underway. Five NLG systems from
three countries are in the process of being evaluated. At the time of
writing, 1028 different evaluation runs---each consisting of one user
interacting with one instance of the virtual environment and being guided
by an NLG system---have been collected, making GIVE the largest-scale
NLG evaluation effort ever.\footnote{See the GIVE website at
 \url{http://www.give-challenge.org/} for more details about this
 challenge.}

\begin{figure}[t]
\centering
%\includegraphics[width=1 \columnwidth]{give_world_no_expl}
\includegraphics[width=0.75\columnwidth]{give_world_2}
\caption{Map of the GIVE development world.}
  \label{fig:give-development-world}
\end{figure}

An example GIVE world map is shown in
Figure~\ref{fig:give-development-world}. To simplify both the planning and
the NLG task, the world is discretised into a set of tiles of equal size.
The user can turn by 90 degree steps in either direction, and can move from
the centre of one tile to the centre of the next tile, provided the path
between two tiles is not blocked. The world also consists of a set of
objects which can be manipulated by the user in various ways. For
instance, in the example world the user's task is to pick up a trophy in
the top left room. The trophy is hidden in a safe behind a picture, so the
user must first move the picture out of the way and open the safe by
pushing a sequence of buttons (the small square boxes on the walls) in the
correct order. In order to get to all these buttons, the user must also
open the door in the centre of the map and deactivate the alarm tile by
pushing further buttons. Figure~\ref{fig:give-planning} shows the encoding
of some of the available GIVE domain actions, in PDDL syntax.

\begin{figure}[p]
\centering
\begin{minipage}{0.5\textwidth}
{\small%
\begin{verbatim}
(:action move
   :parameters (?from - position
                ?to - position
                ?ori - orientation)
   :precondition 
       (and (player-pos ?from) 
            (adjacent ?from ?to ?ori) 
            (player-orient ?ori)
            (not-blocked ?from ?to)
            (not-alarmed ?to))
   :effect 
       (and (not (player-pos ?from))
            (player-pos ?to)))

(:action turn-left
   :parameters (?ori - orientation
                ?newOri - orientation)
   :precondition 
       (and (player-orient ?ori)
            (next-orient-left ?ori ?newOri))
   :effect 
       (and (not (player-orient ?ori))
            (player-orient ?newOri)))

(:action turn-right
   :parameters (?ori - orientation
                ?newOri - orientation)
   :precondition 
       (and (player-orient ?ori)
            (next-orient-right ?ori ?newOri))
   :effect 
       (and (not (player-orient ?ori))
            (player-orient ?newOri)))

(:action manipulate-b1-off-on
   :parameters (?pos - position)
   :precondition 
       (and (state b1 off)
            (player-pos ?pos)
            (position b1 ?pos))
   :effect 
       (and (not (state b1 off))
            (state b1 on)
            (not (state d1 closed))
            (state d1 open) 
            (not (blocked pos_6_5 pos_6_4))
            (not (blocked pos_6_4 pos_6_5))))
\end{verbatim}}%
\end{minipage}
\caption{PDDL actions for the GIVE domain.}
\label{fig:give-planning}
\end{figure}

The plans generated in the GIVE domain are often nontrivial. For instance,
in the above example the shortest plan consists of 108 action steps, with
the first few steps as follows:
%
\begin{enumerate}
\item $\mathsf{turn}\textsf{-}\mathsf{left}(\mathsf{north},
\mathsf{west})$,
\item $\mathsf{move}(\mathsf{pos\_5\_2}, \mathsf{pos\_4\_2}, \mathsf{west})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{b1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_5\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{west}, \mathsf{north})$.
\end{enumerate}

Most of the actions in the above plan are ``move'' actions. This makes the
bulk of the GIVE problem very similar to that of a Gridworld problem (see,
e.g., \citep{Tovey-Koenig:2000}), which also involves finding a route
through a two-dimensional world map with discrete positions. GIVE tasks are
usually much more complex, however, as illustrated by the above example
map: a successful plan must include steps for pressing buttons in the right
order, reasoning about large numbers of world objects, and navigating
through complicated room shapes.

The GIVE task comes with a number of interesting research challenges for
both the NLG and planning communities. On the NLG side, one major problem
is that verbalising each action instance in a plan makes for very
poor natural-language instructions. For instance, the NLG system could
generate an instruction sequence starting with ``turn left; walk forward;
press the button; turn right; walk forward; walk forward; turn right; walk
forward; walk forward; turn left; walk forward''. However, directions of
this form are linguistically clumsy and uninteresting, and it is to be
expected that frustrated users will quickly cancel an interaction with such
a system, resulting in a bad evaluation score. Having just pressed the
button, it would be much better to say ``turn around and walk through the
door''. That is, the system should \emph{summarise} plans by merging
multiple action instances into single instructions before presenting them
to the user.\footnote{This task is similar to the construction of
  ``macro'' actions, but is primarily intended for plan presentation
  purposes. See, e.g., \citep{Botea-etal:05} which contains a survey of
  recent approaches for generating macro actions in planning.}
Conversely, it may also be necessary to \emph{elaborate} on a single
planning step by expressing it with several instructions. Having just
entered the top left room, it may be easier for the user to understand the
instruction ``walk to the centre of the room; turn right; now press the
green button in front of you'', rather than the instruction ``press the
green button on the wall to your right''. Furthermore, in order to plan the
referring expression ``the green button in front of you'' at a time when
the button is not yet in front of the user, the NLG system must keep track
of the hypothetical changes in what objects are visible to the user. Thus,
the NLG acts of referring and instructing must be tightly integrated with
the modules for plan generation.

From a planning perspective, GIVE imposes very strict runtime
requirements on the plan generation task: planning must
happen in real time and the system must respond to a user in a timely
fashion.  If the system takes too much time deliberating over an
instruction to give, rather than actually giving this instruction, the
user may have walked or turned away, thus making the instruction
invalid.  Furthermore, plan execution monitoring also plays an
important role in the GIVE problem.  At a high level, the system needs
to monitor a user's actions and compare them against the generated
instruction set to determine if the user has correctly followed
directions or not. In the case of the latter, new instructions may
have to be generated. In practice, the situation can be quite
complicated since the mental state of the user is not known and so the
system must be inferred from observing the user's actions in real
time. For example, a user directed to ``turn around and walk through
the door'' may not necessarily perform these actions to the letter,
i.e., immediately turning 180 degrees and proceeding directly to the
door. Instead, the user might take a roundabout route through the
room, eventually exiting out the door.  Although the user's actions do
not match the generated instructions exactly, they meet the intended
goal. The system must be able to identify such ``equivalent'' plans
and not immediately generate new instructions as soon as the user's
actions have gone off course. Furthermore, a user can communicate
certain intentions to the system, both through action and
inaction. For instance, the system should infer that a user has failed
to follow instructions if the user exits a room when given a directive
to ``walk to the centre of the room''. The system should also make a
similar conclusion if a user simply does nothing when given the
instruction.

\section{Experiments}
\label{sec:experiments}

We now describe the results of four simple experiments designed to evaluate
the performance of three planners---FF \citep{HoffmannNebel01}, SGPLAN
5.2.2 \citep{hsu06:_new_featur_in_sgplan_for}, and an ad-hoc implementation
of GraphPlan \citep{Blum1997} written in Java---in our NLG domains. The FF
and SGPLAN planners were chosen due to their success in the International
Planning Competitions. Our implementation of GraphPlan is included as a
counterpoint to the other two, more complex, planners. Most notably, our
version of GraphPlan does not perform a preprocessing ``grounding'' step
like FF and SGPLAN but is much more selective in its predicate and action
instantiations.

On first impressions, our results indicate that planning is a promising
tool for both domains.  In the sentence generation domain, FF dramatically
outperforms the best previously known algorithm for the same problem (a
reimplementation of \citep{Stone2003a}), although the latter approach uses
a greedy search algorithm with a heuristic that is hand-tailored to the
domain. In the GIVE domain, SGPLAN computes a domain plan from the initial
state to the goal in 0.3 seconds, which is fast enough for moderately-sized
problems instances in the application.\footnote{All
  runtimes were measured on a Pentium 4 CPU running at 3 GHz. Java programs
  were allowed to ``warm up'', i.e., the planner was run five times and the
  first four measurements discarded to ensure that the JVM had just-in-time
  compiled all relevant bytecode.}

We also observe that although FF and SGPLAN manage the search for plans
very efficiently, they both spend comparatively large amounts of time
computing predicate and action instantiations in their preprocessing
stages, many of which are then never used during the search. FF's
``grounding time'' in particular dominates its overall planning time,
leading to some conflicting results. In some cases, our Java implementation
of GraphPlan significantly outperforms both state-of-the-art planners. The
experiments described below seek to improve our understanding of this
situation.

\subsection{Experiment 1: Sentence generation}
\label{sec:exper-1:-sent}

In the first experiment, we construct a series of sentence generation
problems which require the planner to compute a plan representing the
sentence ``Mary likes the Adj$_1$ \ldots Adj$_n$ rabbit.''  Each problem
instance assumes a certain number $m$ of rabbits that were distinguished by
$n \leq m$ different properties, such that all $n$ properties are required
to distinguish the target referent from all other rabbits.  The $n$
properties are realized as $n$ different adjectives, in any order.  This
setup allows us to control the plan length (a plan with $n$ properties will
have length $n+4$) and the universe size (the universe will contain $m+1$
individuals in addition to the differently-typed individuals used to encode
the grammar).

\begin{figure}
  \centering
  \includegraphics[width=0.75\columnwidth]{graph-exp1}
  %\includegraphics[width=1\columnwidth]{pic-runtime-modifiers-with-sgplan}
  \caption{Results for the sentence generation domain. The
    horizontal axis represents parameters $(m,n)$ from $(1,1)$ to
    $(10,10)$ in lexicographical order. The vertical axis is the
    runtime in milliseconds.}
  \label{fig:runtimes-crisp}
\end{figure}

The results of this experiment are shown in the graph in
Figure~\ref{fig:runtimes-crisp}. The input parameters $(m,n)$ are plotted
in lexicographic order on the horizontal axis and the runtime is shown in
milliseconds on the vertical axis. These results reveal a number of
interesting insights. First, FF significantly outperforms SGPLAN in this
domain, sometimes by a factor of 10 or more.\footnote{Experiments with
 SGPLAN use a pre-release version of SGPLAN, kindly provided to us by
 Chih-Wei Hsu.The release version of SGPLAN 5.2.2 had a bug which caused it
 to crash on some of our problem instances.}
Second, FF's runtime is dominated by its initial grounding step, in which
it computes the ground instances of all predicates and actions used in the
planning problem in order to avoid unnecessary instantiation during search.
In particular, the ratio of grounding time to total runtime is generally
above 85\%, and rises to above 99\% at $m=11$, which is still a relatively
small universe for this application.\footnote{The
  ``grounding'' time reported here is what FF reports as ``time spent:
  instantiating action templates''.}

In our testing examples, the time spent by FF on grounding is such that it
is consistently outperformed by our Java implementation of GraphPlan, which
only computes instances of predicates and actions as they are discovered
during the construction of the planning graph. Although FF is consistently
much faster as far as pure search time is concerned, our results indicate
that FF's performance is much more sensitive to the domain size: if we fix
$n=1$, FF takes 60 ms to compute a plan at $m=1$, but 2.4 seconds to
compute the same plan at $m=10$. By comparison, our GraphPlan
implementation takes 30 ms at $m=1$ and still only requires 50 ms at
$m=10$. Conversely, GraphPlan's runtime grows much faster with the plan
size (i.e., with growing values of $n$ for a fixed $m$). Larger, but still
realistically-sized, instances of the sentence generation problem are still
problematic for all the planners we tested.

\subsection{Experiment 2: Minimal GIVE worlds}
\label{sec:exper-2:-minim}

In the second experiment, we evaluate the performance of the planners on
problems arising in the GIVE domain. We construct a series of test worlds,
similar to the one illustrated in Figure~\ref{fig:give-minimal}. These
worlds consist of a $2n$ by $h$ grid of positions, such that there are
buttons at positions $(2i-1,1)$ and $(2i,h)$ for $1 \leq i \leq n$. The
player starts in position $(1,1)$ and must press all the buttons
to successfully complete the game. The world is generated as a GIVE world
description, and then automatically converted into a planning problem by
the GIVE software. For instance, Figure~\ref{fig:give-planning} shows some
of the actions available in the GIVE domain in PDDL syntax.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\columnwidth]{pic-buttons}
  \caption{Minimal GIVE world.}
  \label{fig:give-minimal}
\end{figure}

Results for the $h=20$ case, with $n$ ranging from $1$ to $40$, are shown
in Figure~\ref{fig:give-runtime-minimal}. The most obvious result is that
FF is unable to solve any problems beyond $n=13$ on our experimentation
machine within the memory limit of 1 GB. SGPLAN, on the other hand, solves
instances beyond $n=40$ without major problems. In this case, our
implementation of GraphPlan is unable to find any of these plans. The time
spent on grounding is not a major factor in either planner, probably
because the planners need more time to actually compute the plan---for
instance, the optimal plan for the problem $n=40$ has a length of about
1600 steps. As a concrete example, the following is a minimal plan for the
case of a $2$ by $2$ grid with $2$ buttons for the player to press (i.e.,
$n=1$, $h=2$):

\begin{enumerate}
\item $\mathsf{move}(\mathsf{pos\_1\_1},\mathsf{pos\_1\_2}, \mathsf{north})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{u1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_1\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{north}, \mathsf{east})$,
\item $\mathsf{move}(\mathsf{pos\_1\_2}, \mathsf{pos\_2\_2},
  \mathsf{east})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{east}, \mathsf{south})$,
\item $\mathsf{move}(\mathsf{pos\_2\_2}, \mathsf{pos\_2\_1}, \mathsf{south})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{l1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_2\_1})$.
\end{enumerate}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.75\columnwidth]{graph-exp2}
  %\includegraphics[width=1\columnwidth]{pic-runtime-buttons}
  \caption{Results on the minimal GIVE
    domain for $h=20$. The horizontal axis is $n$. The vertical axis
    is the runtime in seconds.}
  \label{fig:give-runtime-minimal}
\end{figure}

\subsection{Experiment 3: GIVE worlds with extra positions}
\label{sec:experiment-3:-give}

In the third experiment, we vary the structure of the GIVE world in order
to judge the effect that universe size has on the planning problem in this
domain. Starting with the ordinary GIVE world described in Experiment~2, we
extend the world map by adding another $w$ by $h$ empty ``junk'' positions
to the right of the minimal world, as shown in Figure~\ref{fig:give-junk}.
These new positions are not actually required in any plan, but extend the
size of the state space and approximate the situation in the actual GIVE
domain where most grid positions are never used. We leave the initial state
and goal untouched. As before, we generate a GIVE world description and
then convert it into a planning problem in PDDL.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.80\columnwidth]{pic-empty-buttons}
  \caption{GIVE world with extra ``junk'' positions.}
  \label{fig:give-junk}
\end{figure}

Results for the $h=20$, $n=5$ case with $w$ ranging from $1$ to $70$ are
shown in Figure~\ref{fig:give-runtime-junk}. As in Experiment~2, FF again
runs out of memory, this time at $w=17$, while SGPLAN easily solves inputs
beyond $w=70$. As in Experiment~2, our implementation of GraphPlan is
unable to generate any of these plans. Unlike Experiment 2, however, both
FF and SGPLAN now spend a substantial proportion of their time on
grounding. In SGPLAN, this translates to a ``parsing time'' (which we
assume includes grounding) which grows from 180 ms to 21.7 seconds as $w$
grows from $1$ to $75$. The rest of the runtime, which also includes the
search time, only grows from 400 ms to 2.3 seconds. This difference is
particularly dramatic given that the actual optimal plan in each case is an
identical plan of about 100 steps (i.e., the one we would have found in
Experiment 2 for the $h=20$, $n=5$ case). The planning times for these
instances are also concerning since times over a couple seconds will
negatively affect the overall response time of the system, which must react
in real time to user actions.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.75\columnwidth]{graph-exp3}
  %\includegraphics[width=1\columnwidth]{pic-runtime-empty-world}
  \caption{Results on the GIVE domain with junk
    positions for $h=20$ and $n=5$. The horizontal axis is $w$.
    The vertical axis is the runtime in seconds.}
  \label{fig:give-runtime-junk}
\end{figure}

\subsection{Experiment 4: GIVE worlds with inaccessible positions}
\label{sec:experiment-4:-give}

In the final experiment, we consider a variation on the GIVE worlds from
Experiment~3. As before, we begin with a base grid of $2n$ by $h$
positions, with buttons at positions $(2i-1,1)$ and $(2i,h)$ for $1 \leq i
\leq n$. We then generate a second grid of $w$ by $h$ positions and place
an additional button in this grid. Unlike Experiment~3, the second grid
does not extend the first grid but is ``disconnected'' so its positions are
inaccessible from the first grid, as shown in
Figure~\ref{fig:give-junk-nosoln}. Since the geometry of the grid makes it
impossible to construct a plan for pressing all the buttons in the world,
we instead investigate the time it takes a planner to arrive at the
conclusion that the problem cannot be solved.

Results for the $h=20$, $n=5$ case with $w$ ranging from 1 to 50 are shown
in Figure~\ref{fig:give-runtime-nosoln}.\footnote{FF and SGPLAN do
 not display runtimes when they fail to construct a plan so an external timing
 program was used to generate these results.}
In this case, the runtimes for FF rise sharply around $w=28$, producing
times that are unacceptable for an NLG system in practice. FF is unable
to complete problem instances above $w=35$. SGPLAN, by comparison, performs
exceptionally well and is able to complete problems instances up to $w=50$
in about 0.3 seconds and $w=80$ in about 0.6 seconds. These results
demonstrate that SGPLAN has a significant advantage over FF in this
experiment. 
%Our implementation of GraphPlan is unable to complete any of
%these problem instances.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.85\columnwidth]{pic-empty-inaccessible}
  \caption{GIVE world with inaccessible positions.}
  \label{fig:give-junk-nosoln}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.85\columnwidth]{graph-exp4}
  %\includegraphics[width=1\columnwidth]{pic-runtime-empty-world}
  \caption{Results on the GIVE domain with inaccessible
    positions for $h=20$ and $n=5$. The horizontal axis is $w$.
    The vertical axis is the runtime in seconds.}
  \label{fig:give-runtime-nosoln}
\end{figure}

\section{Discussion}
\label{sec:discussion}

The positive conclusion we draw from our simple set of experiments is that
modern planners are fairly good at controlling the search for many of the
moderately-sized NLG problems we looked at. This is particularly true for
FF in the sentence generation domain, which generates nontrivial 14-word
sentences in about two seconds. Our results also indicate there is still
room for improvement, however: special-purpose algorithms are able to
generate much larger referring expressions in milliseconds
\citep{AreKolStr08}. Although search efficiency is not the main concern in
these domains, it nevertheless becomes a factor in scenarios like
Experiment~2, where the world is almost completely unconstrained and
(unsurprisingly) the search becomes much slower.

The most restrictive bottleneck in the two NLG domains we investigated is
the initial grounding step that many modern planners perform. While this
may be a good strategy for traditional planning benchmarks and IPC
domains---where plan length often dominates universe size and an initial
investment into grounding pays off in saved instantiations during
search---this approach has a very pronounced effect in domains like ours.
As our experiments show, there are natural planning domains in which
relatively short plans must be computed in large universes. For instance,
in the sentence generation domain, it is not unrealistic to look for plans
of length 20 over a universe consisting of thousands of domain individuals
and tens of thousands of possible action instantiations, some of which
require three domain individuals as parameters. Planners like FF and SGPLAN
are less effective in such domains, to the point that grounding
often dominates the total running time of these planners. (In the case of
Experiment~1, our ad-hoc implementation of GraphPlan outperforms both
planners in the sentence generation task.)

Since our domains are not that unusual in their structure and composition,
we hope that the lessons learned from our experiences can help improve the
performance of current planning systems. For instance, the common strategy
of splitting the universe into several types of individuals will not be a
very effective solution in our NLG domains given the large number of
individuals that could exist for a given type. On the other hand,
sophisticated techniques based on reachability analysis may have the
potential to improve this situation, provided such approaches can avoid the
drawbacks of unnecessary grounding.\footnote{See
  \citep{Bryce:07} for a good survey on reachability heuristics for planning
  graph-based approaches to plan generation.}
While recent trends in planning research have (rightfully) focused on the
development of algorithms that control search in sophisticated ways,
resulting in state-of-the-art planners that are more powerful and
successful than their predecessors, we nevertheless believe that
improvements are necessary if planning is to become a more mature
technology that can offer tools to a wider community of users and
researchers. At the end of the day, real-world users will care about the
\emph{total} runtime of a planner, which is more than just the search time.

By and large, our experiences with the planning community from a user's
point of view have been favourable. Thanks to the planning competitions, it
is easy to identify and download a fast state-of-the-art planner, at least
for Linux. However, in the course of our experiments we found and reported
bugs in both FF and SGPLAN. We also discovered that deciding as to which
planner is appropriate for a particular task is not always straightforward,
as the best performing planner in the IPC is not necessarily suitable for
every planning task. As our experiments have shown, even planners as
closely related as FF and SGPLAN can differ significantly in their
performance on different domains (for instance, our results in
Experiment~4).

\section{Conclusion}
\label{sec:conclusion}

In this paper, we introduced two novel planning domains arising from
problems in natural language generation: the sentence generation domain and
the Generating Instructions in Virtual Environments (GIVE) challenge
domain. We also reported the results of a set of experiments designed to
evaluate the suitability of off-the-shelf planners (particularly FF and
SGPLAN) in these domains.

Our results were mixed. While the planners we tested do an impressive job
of controlling the complexity of search in many domains, they also suffer
from practical problems that limit their performance in unexpected ways: in
both of our NLG domains the grounding step performed by FF and SGPLAN
dominates the time it takes to perform the search itself. As future work,
we hope to extend our analysis to other planners that do not perform the
initial grounding step, planners such as SHOP2
\citep{DBLP:journals/jair/NauAIKMWY03} that have traditionally performed
well on a wide range of domains, and new planners introduced in the latest
edition of the IPC.\footnote{See
 \url{http://ipc.informatik.uni-freiburg.de/} for details of the 2008
 International Planning Competition (deterministic track).}

The NLG community's recent interest in planning presents a valuable
opportunity for planning researchers. Domains like GIVE highlight certain
challenges, such as plan execution monitoring and plan presentation (i.e.,
summarisation and elaboration), but also offer a platform on which such
technologies can be evaluated in experiments with human users. Although we
have focused on classical planning problems in this work, research related
to reasoning under uncertainty, resource management, and planning with
knowledge and sensing, can also be investigated in these settings. As such,
we believe our domains would provide suitable challenges for planners
entered in future editions of the IPC. Provided some of the challenges we
have identified can be addressed, NLG-inspired problem domains offer a
constructive platform for the planning community to showcase their
techniques to a wider audience---and to improve the quality of their tools
for real-world planning tasks.

\section*{Acknowledgements}

This work arose in the context of the Planning and Language Interest Group
at the University of Edinburgh. The authors would like to thank all members
of this group, especially Hector Geffner and Mark Steedman, for interesting
discussions. This work was partially supported by the DFG Research
Fellowship ``CRISP: Efficient integrated realization and microplanning''
and by the European Commission through the PACO-PLUS project
(FP6-2004-IST-4-27657).

\bibliography{references}
\bibliographystyle{ci}

\pagebreak
\listoffigures

\end{document}