### 2009-05-07 trunk/Papers/ci-spark/discussion.tex

\section{Discussion}
\label{sec:discussion}

The positive conclusion we draw from our simple set of experiments is that
modern planners are fairly good at controlling the search for many of the
moderately-sized NLG problems we looked at. This is particularly true for
FF in the sentence generation domain, which generates nontrivial 14-word
sentences in about two seconds. Our results also indicate there is still
room for improvement, however: special-purpose algorithms are able to
generate much larger referring expressions in milliseconds
\citep{AreKolStr08}. Although search efficiency is not the main concern in
these domains, it nevertheless becomes a factor in scenarios like
Experiment~2, where the world is almost completely unconstrained and
(unsurprisingly) the search becomes much slower.

The most restrictive bottleneck in the two NLG domains we investigated is
the initial grounding step that many modern planners perform. While this
may be a good strategy for traditional planning benchmarks and IPC
domains---where plan length often dominates universe size and an initial
investment into grounding pays off in saved instantiations during
search---this approach has a very pronounced effect in domains like ours.
As our experiments show, there are natural planning domains in which
relatively short plans must be computed in large universes. For instance,
in the sentence generation domain, it is not unrealistic to look for plans
of length 20 over a universe consisting of thousands of domain individuals
and tens of thousands of possible action instantiations, some of which
require three domain individuals as parameters. Planners like FF and SGPLAN
are less effective in such domains, to the point that grounding
often dominates the total running time of these planners. (In the case of
Experiment~1, our ad-hoc implementation of GraphPlan outperforms both
planners in the sentence generation task.)

Since our domains are not that unusual in their structure and composition,
we hope that the lessons learned from our experiences can help improve the
performance of current planning systems. For instance, the common strategy
of splitting the universe into several types of individuals will not be a
very effective solution in our NLG domains given the large number of
individuals that could exist for a given type. On the other hand,
sophisticated techniques based on reachability analysis may have the
potential to improve this situation, provided such approaches can avoid the
drawbacks of unnecessary grounding.\footnote{See
  \citep{Bryce:07} for a good survey on reachability heuristics for planning
  graph-based approaches to plan generation.}
While recent trends in planning research have (rightfully) focused on the
development of algorithms that control search in sophisticated ways,
resulting in state-of-the-art planners that are more powerful and
successful than their predecessors, we nevertheless believe that
improvements are necessary if planning is to become a more mature
technology that can offer tools to a wider community of users and
researchers. At the end of the day, real-world users will care about the
\emph{total} runtime of a planner, which is more than just the search time.

By and large, our experiences with the planning community from a user's
point of view have been favourable. Thanks to the planning competitions, it
is easy to identify and download a fast state-of-the-art planner, at least
for Linux. However, in the course of our experiments we found and reported
bugs in both FF and SGPLAN. We also discovered that deciding as to which
planner is appropriate for a particular task is not always straightforward,
as the best performing planner in the IPC is not necessarily suitable for
every planning task. As our experiments have shown, even planners as
closely related as FF and SGPLAN can differ significantly in their
performance on different domains (for instance, our results in
Experiment~4).

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "manuscript"
%%% End: