### 2009-05-07 trunk/Papers/ci-spark/manuscript.tex

%%%
%%% Computational Intelligence manuscript based on SPARK'08 paper
%%%
%%% Alexander Koller and Ronald Petrick
%%% 
%%% Updated: 2008/12/22
%%%
\documentclass[letterpaper]{article}
\usepackage{times}
\usepackage{graphicx}
\usepackage{url}
\usepackage[round]{natbib}
\usepackage{setspace}
\usepackage[paper=letterpaper,margin=1.0in]{geometry}

\title{Experiences with Planning for Natural Language Generation}
\author{\textbf{Alexander Koller} \\
Exzellenzcluster / FR 4.7 Computerlinguistik \\
Saarland University \\
Postfach 151150 \\
66041 Saarbr\"ucken \\
Germany \\
Phone: +49 681 302 70040 \\
Fax: +49 681 302 4351 \\
E-mail: \url{koller@mmci.uni-saarland.de}
\and
\textbf{Ronald P. A. Petrick} \\
School of Informatics \\
University of Edinburgh \\
Informatics Forum, 10 Crichton Street \\
Edinburgh \ EH8 9AB \\
Scotland, United Kingdom \\
Phone: +44 131 650 4426 \\
Fax: +44 131 650 6899 \\
E-mail: \url{rpetrick@inf.ed.ac.uk}}

%\date{22 December 2008}
\date{}

\pagestyle{plain}
\pagenumbering{arabic}
\doublespacing

\bibpunct{(}{)}{;}{a}{}{,}

\newcommand{\todo}[1]{\textbf{(#1)}}

\begin{document}

\maketitle

\pagebreak
\input abstract

\pagebreak

\input introduction
\input nlg-as-planning
\input domains
\input experiments
\input discussion
\input conclusion

\bibliography{references}
\bibliographystyle{ci}

\pagebreak
\listoffigures

\end{document}