### 2009-05-07 trunk/Papers/ci-spark/manuscript.tex
changed to official CI style

%%%
%%% Computational Intelligence manuscript based on SPARK'08 paper
%%%
%%% Alexander Koller and Ronald Petrick
%%% 
%%% Updated: 2008/12/22
%%%
\documentclass[useAMS,usenatbib]{coin}

%\usepackage{times}
\usepackage{graphicx}
\usepackage{url}
\usepackage{natbib}
%\usepackage{setspace}
%\usepackage[paper=letterpaper,margin=1.0in]{geometry}

\title{Experiences with Planning for Natural Language Generation}

\author{{\sc Alexander Koller}\\
        {\it Cluster of Excellence, Saarland University, Saarbr\"ucken, Germany}\\ 
		\and
        {\sc Ronald P. A. Petrick}\\
        {\it School of Informatics, University of Edinburgh,
          Edinburgh, UK\thanks{Address correspondence to XXXX.}}
}

% \author{\textbf{Alexander Koller} \\
% Exzellenzcluster / FR 4.7 Computerlinguistik \\
% Saarland University \\
% Postfach 151150 \\
% 66041 Saarbr\"ucken \\
% Germany \\
% Phone: +49 681 302 70040 \\
% Fax: +49 681 302 4351 \\
% E-mail: \url{koller@mmci.uni-saarland.de}
% \and
% \textbf{Ronald P. A. Petrick} \\
% School of Informatics \\
% University of Edinburgh \\
% Informatics Forum, 10 Crichton Street \\
% Edinburgh \ EH8 9AB \\
% Scotland, United Kingdom \\
% Phone: +44 131 650 4426 \\
% Fax: +44 131 650 6899 \\
% E-mail: \url{rpetrick@inf.ed.ac.uk}}

%\date{22 December 2008}
\date{}

%\pagestyle{plain}
%\pagenumbering{arabic}
%\doublespacing

\bibpunct{(}{)}{;}{a}{}{,}

\newcommand{\todo}[1]{\textbf{(#1)}}
\newcommand{\addact}[1]{\mathsf{add}\textsf{-}\mathsf{#1}}

\begin{document}

%\pagebreak
\input abstract

\maketitle

%\pagebreak

\input introduction
\input nlg-as-planning
\input domains
\input experiments
\input discussion
\input conclusion

\bibliography{references}
\bibliographystyle{ci}

\pagebreak
\listoffigures

\end{document}