### 2009-05-31 trunk/Papers/ci-spark/nlg-as-planning.tex

\section{NLG as Planning} \label{sec:nlg-as-planning}

The task of generating natural language from semantic representations (NLG) is
typically split into two parts: the \emph{discourse planning} task, which
selects the information to be conveyed and structures it into sentence-sized
chunks, and the \emph{sentence generation} task, which then translates each of
these chunks into natural language sentences. The sentence generation task is
often divided into two parts of its own---the \emph{sentence planning} task,
which enriches the input by, e.g., determining object references and selecting
some lexical material, and the \emph{surface realization} task, which maps the
enriched meaning representation into a sentence using a grammar. The chain of
domain planning, sentence planning, and surface realization is sometimes called
the ``NLG pipeline'' \citep{reiter00building}.

Viewing generation as a planning problem has a long tradition in the NLG
literature. \citet{perrault80} presented an approach to discourse planning in
which the planning operators represented individual speech acts such as
``request'' and ``inform''. This idea was later expanded, e.g., by
\citet{young94dpocl}. On the other hand, researchers such as
\citet{appelt:planning} and \citet{hovy88} used techniques from hierarchical
planning to expand a high-level plan consisting of speech acts into more
detailed specifications of individual sentences. Their systems also covered some
aspects of sentence planning. However, these systems used very expressive
logics, which were designed to reason about beliefs and intentions, to represent
the planning state and the planning operators. Furthermore, most of them used
ad-hoc planning algorithms with rather naive search strategies, which did not
scale to realistic inputs. As a consequence, the NLG-as-planning approach was
mostly marginalized throughout the 1990s.

More recently, there has been a string of publications by various authors with a
renewed interest in the generation-as-planning approach, motivated by the
ongoing development of increasingly more efficient and expressive planners. For
instance, \cite{KolSto07} propose an approach to sentence generation (i.e., the
sentence planning and surface realization modules of the pipeline) as
planning---an approach we will explore in more detail below
(Section~\ref{sec:domain-crisp}). \citet{Steedman-Petrick:07} revisit the
analysis of indirect speech acts with modern planning technology, viewing the
problem as an instance of the general problem of planning with incomplete
information and sensing actions. In addition, \citet{benotti08b} uses planning
to explain the accommodation of presuppositions, and
\citet{brenner08:_contin_multiag_plann_approac_to_situat_dialog} use multi-agent
planning to model the joint problem-solving behaviour of agents in a situated
dialogue. While these approaches focus on different issues compared to the
1980's NLG-as-planning literature, they all apply \emph{existing},
well-understood planning approaches to linguistic problems, in order to utilise
the rich set of modelling tools provided by modern planners, and in the hope
that such planners can efficiently solve the hard search problems that arise in
NLG \citep{KolStr02}. This paper aims to investigate whether existing planners
achieve this latter goal.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "manuscript"
%%% End: