### 2009-06-02 trunk/Papers/ci-spark/abstract.tex
changes to abstract, section1, and section 2

\begin{abstract}
  Natural language generation (NLG) is a major subfield of
  computational linguistics with a long tradition as an application
  area of automated planning systems. While the connection between
  NLG and planning has not received much attention in current
  mainstream NLG research, several recent publications have sparked
  a renewed interest in this area. In this paper, we investigate the
  extent to which these new NLG approaches profit from the advances
  in planner expressiveness and efficiency. Our findings are mixed.
  While modern planners can readily handle the search problems that
  arise in our NLG experiments, their overall runtime is often
  dominated by the grounding step they perform as preprocessing.
  Furthermore, small changes in the structure of a domain can
  significantly shift the balance between search and preprocessing.
  Overall, our experiments show that the off-the-shelf planners we
  tested are unusably slow for nontrivial NLG problem instances. As
  a result, we offer our domains and experiences as challenges for
  the planning community.
\end{abstract}

\begin{keywords}
  natural language generation, planning
\end{keywords}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "manuscript"
%%% End: