### 2009-06-02 trunk/Papers/ci-spark/introduction.tex

\section{Introduction}
\label{sec:introduction}

\emph{Natural language generation} (NLG; \citealp{reiter00building})
is one of the major subfields of natural language processing,
concerned with computing natural language sentences or texts that
convey a given piece of information to an audience. While the output
of a generation task can take many forms, including written text,
synthesised speech, or embodied multimodal presentations, the
underlying NLG problem in each case can be modelled as a problem of
achieving a (communicative) goal by successively applying a set of
(communicative) actions. This view of NLG as goal-directed action has
clear parallels to \emph{automated planning}, which seeks to find
general techniques for efficiently solving the action sequencing
problem.

Treating generation as planning has a long history in NLG, ranging
from the initial attempts of the field to utilise early planning
approaches \citep{perrault80,appelt:planning,hovy88,young94dpocl}, to
a recent surge of research
\citep{Steedman-Petrick:07,KolSto07,brenner08:_contin_multiag_plann_approac_to_situat_dialog,benotti08b}
seeking to capitalise on the improvements modern planners offer in
terms of efficiency and expressiveness. This paper attempts to assess
the usefulness of current planning techniques to NLG by investigating
some representative generation problems, and by evaluating whether
automated planning has advanced to the point that it can provide
solutions to such NLG applications---applications that are not
currently being investigated by mainstream planning research.

To answer this question, we proceed in two steps. First, we present
two generation problems that have recently been cast as planning
problems: the sentence generation task and the GIVE task. In the
sentence generation task, we concentrate on generating a single
sentence that expresses a given meaning. In this case, a plan encodes
the necessary sentence with the actions in the plan corresponding to
the utterance of individual words \citep{KolSto07}. In the GIVE domain
(``Generating Instructions in Virtual Environments''), we describe a
new shared task that was recently posed as a challenge for the NLG
community \citep{ByrKolStrCasDalMooObe09}. GIVE uses planning as part
of a larger NLG system for generating natural-language instructions
that guide a human user in performing a given task in a virtual
environment.

Second, we evaluate the performance of several off-the-shelf planners
on the planning domains into which these two generation problems
translate. Among the planners we test, we explore the efficiency of FF
\citep{HoffmannNebel01}---a planner that has arguably had the greatest
impact on recent approaches to deterministic planning---and some of its
descendants, such as SGPLAN
\citep{hsu06:_new_featur_in_sgplan_for}. All of the planners we test
are freely available, support an expressive subset of the Planning
Domain Definition Language (PDDL; McDermott et al.~\citeyear{PDDL}),
and have been successful on both standard planning benchmarks and the
problems of the International Planning Competition (IPC).\footnote{See
  \url{http://ipc.icaps-conference.org/} for information about past
  editions of the IPC. Also see \citep{Hoffmann-Edelkamp:05} for a good
  overview of the deterministic track of the 2004 competition.}
Using these planners---together with an ad-hoc Java implementation of
GraphPlan \citep{Blum1997} serving as a baseline for certain
experiments---we perform a series of tests on a range of problem
instances in our NLG domains.

Overall, our findings are mixed. On the one hand, we demonstrate that
some planners can readily handle the \emph{search} problems that arise
in our testing domains on realistic inputs, which is promising given
the challenging nature of these tasks (e.g., the sentence generation
task is NP-complete; see \citealt{KolStr02}). On the other hand, these
same planners often spend tremendous amounts of time on
\emph{preprocessing} to analyse the problem domain in support of the
search. On many of our problem instances, the preprocessing time
overshadows the search time. (For instance, FF spends 90\% of its
runtime in the sentence generation domain on preprocessing.)
Furthermore, small changes in the structure of a planning domain can
dramatically shift the balance between preprocessing and search. As a
consequence, we are forced to conclude that the off-the-shelf planners
we investigated are generally too slow to be useful in real NLG
applications. It is also our hope, however, that these results will
spark an interest to improve the quality of planner
implementations---especially in the area of preprocessing
techniques---and to this end we offer our domains and experiences as
challenges for the planning community.

The remainder of this paper is structured as follows. In
Section~\ref{sec:nlg-as-planning}, we introduce the idea of NLG as
planning and briefly review the relevant literature. In
Section~\ref{sec:domains}, we describe a set of planning problems
associated with two NLG tasks: sentence planning and situated
instruction generation. In Section~\ref{sec:experiments}, we report on
our experiments with these planning problems.  In
Section~\ref{sec:discussion} we discuss our results and overall
experiences, and conclude in Section~\ref{sec:conclusion}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "manuscript"
%%% End: