### 2009-06-17 trunk/Papers/crisp-rtg-09/crisp-rtg-09.tex
ported npc proof and crisp encoding

\documentclass[11pt,a4]{article}

\usepackage{url}
\usepackage{graphicx}
\usepackage[round]{natbib}
\usepackage{amsfonts,amsthm}

\def\N{\mathbb{N}}
\newcommand{\sem}{\mathsf{sem}}
\newcommand{\self}{\mathsf{self}}
\newcommand{\produ}{\mathsf{prod}}
\newcommand{\roles}{\mathsf{roles}}
\newcommand{\Neq}{{:}}
\newcommand{\refr}{\mathsf{ref}}
\newcommand{\id}{\mathsf{id}}
\newcommand{\idsem}{\mathsf{idsem}}
\newcommand{\Vars}{\mathsf{Vars}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{kor}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{verm}{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{bem}{Remark}

\newcommand{\actionname}[3]{\mathbf{\mathtt{#1}\mbox{-}\mathtt{#2}\mbox{-}\mathtt{#3}}}
\newcommand{\action}[4]{\textbf{Action \texttt{#1}$(#2)$.}\\
\strut\quad   Precond:$\;$ \parbox[t]{12cm}{\ensuremath{#3}}\\
\strut\quad   Effect:$\;$ \parbox[t]{12cm}{\ensuremath{#4}}}
\newcommand{\actionc}[4]{\textbf{Action \texttt{#1}$(#2)$.}
Precond:$\;$ \parbox[t]{3cm}{\ensuremath{#3}}\\
\strut\quad   Effect:$\;$ \parbox[t]{12cm}{\ensuremath{#4}}}

\newcommand{\open}{\ensuremath{\mathsf{open}}}
\newcommand{\gensym}{\ensuremath{\mathsf{gensym}()}}

\title{Sentence generation with RTGs}
\author{Alexander Koller \\ Saarland University \\
  \url{koller@mmci.uni-saarland.de}}
\date{\today\ (v3.0)}

\begin{document}

\maketitle

\section{Introduction}

\section{The generation problem of regular tree grammars}
\label{sec:gener-probl-regul}

We start by defining the generation problem of regular tree
grammars. We will first show how RTGs can be equipped with semantic
information; we call the result a \emph{regular generation grammar
  (RGG)} (see Def.~\ref{def:rgg}). Then we will define the
grammatically correct derivations of an RGG, and a \emph{successful}
derivation as one that also achieves a given communicative goal in
Def.~\ref{def:rgg-deriv}. We will then discuss an example in
Section~\ref{sec:rgg-example}.

\subsection{Definitions}

I write $\N_n$ for the set $\{1,\ldots,n\}$.  I write $T_A(B)$ for the
set of terms $f(a_1,\ldots,a_n)$ with $f \in A$ and $a_1,\ldots,a_n
\in B$.  I also assume that the set of nodes in a tree is a set of
strings in $\N^*$ that is closed under left sister (i.e., if $ui$ is a
node of some tree and $i>1$, then $u(i-1)$ is also a node) and prefix
(i.e., if $ui$ is a node of the tree, then $u$ is too).

\begin{definition} \label{def:rgg}
  A \emph{regular generation grammar (RGG)} is a tuple $G =
  (N,\Sigma,S,Pred,R,L)$ consisting of a terminal alphabet $\Sigma$, a
  nonterminal alphabet $N$, a start symbol $S \in N$, a set $Pred$ of
  \emph{predicate symbols}, a set $R$ of \emph{roles}, and a finite
  set $L$ of \emph{lexicon entries} $l = (P,\sem,r)$, where $P = A
  \rightarrow f(B_1,\ldots,B_n)$ is an RTG production rule over $N$
  and $\Sigma$, $r:\N_n \rightarrow R$ an assignment of roles to
  right-hand nonterminal occurrences that assigns the role $\self$ to
  exactly one number, and $\sem \subseteq
  T_{Pred}(\{r(1),\ldots,r(n)\})$ is the semantic representation.
\end{definition}

We write $A \rightarrow f(B_1\Neq r_1,\ldots,B_n \Neq r_n)$ to
abbreviate a production rule with a role assignment that maps $i$ to
$r_i$ for each $1 \leq i \leq n$. We write $\produ(l)$, $\roles(l)$,
and $\sem(l)$ for the production rule, the set of roles
$\{r(1),\ldots,r(n)\}$, and the $\sem$ value of the lexicon entry
respectively.

We will now define the derivations of RGGs. Derivations are sequences
of lexicon entries. It is intuitively clear that the sequence of RTG
production rules in each of these lexicon entries should constitute an
ordinary RTG derivation. Furthermore, we require that there is a
function $\id$ that maps (the position of) a lexicon entry in the
derivation and a role to a variable, and a variable assignment $\refr$
that maps variables to individuals in the universe. For a derivation
to be correct, we insist that the variables assigned by $\id$ are
chosen consistently, i.e.\ if we use some rule $\produ(l_k)$ to expand
a right-hand nonterminal $A \Neq r$ introduced by the rule
$\produ(l_i)$, then the $\self$ variable of $k$ must be the same as
the $r$ variable of $i$.

\begin{definition} \label{def:rgg-deriv}
  Assume a universe $U$ and an infinite set $\Vars$ of variables. A
  \emph{derivation} of a regular generation grammar $G$ is a sequence
  $d = l_1,\ldots,l_n$ of lexicon entries from $G$ such that there are
  functions $\id:\N_n \times R \rightsquigarrow \Vars$ and
  $\refr:\Vars \rightarrow U$ such that
  \begin{enumerate}
  \item grammaticality: the sequence $\produ(l_1),\ldots,\produ(l_n)$
    is a complete RTG derivation that maps $S$ into a tree of terminal
    symbols;
  \item definedness: $\id(i,r)$ is defined iff $r \in
    \roles(l_i)$;
  \item consistent reference: if $\produ(l_k)$ expands the right-hand
    nonterminal $A:r$ in $\produ(l_i)$, then $\id(k,\self) =
    \id(j,r)$;
  \item naming apart: if $r_1,r_2 \in \roles(l_i)$ and $r_1\neq r_2$,
    then $\id(i,r_1) \neq \id(i,r_2)$.
  \end{enumerate}
\end{definition}

We write $\idsem(l_i) = \{P(x_1,\ldots,x_m) \;|\;
\mbox{$P(r_1,\ldots,r_m) \in \sem(l_i)$ and $x_k = \id(i,r_k)$ for all
  $k$}\}$, and take $||S||_M$ to be the set of all variable
assignments $\Vars \rightarrow U$ that satisfy $S$ in all models over
$U$ that satisfy all formulas in the set $M$.

Then we can define a \emph{successful derivation} as a derivation that
also achieves a given communicative goal. We assume two knowledge
bases $SKB$ and $HKB$, which are sets of ground atoms representing the
speaker's and hearer's knowledge, respectively. Then we insist that
the generated statement is truthful (i.e., the semantic
representations of all used lexicon entries are supported by the SKB),
the communicative goal was achieved, and all referring expressions can
be resolved uniquely by the hearer.

\begin{definition} \label{def:rgg-succ-deriv} A \emph{successful
    derivation} of $G$ for the communicative goal $C \subseteq
  T_{Pred}(U)$, the target referent $e \in U$, and the speaker and
  hearer knowledge bases $SKB, HKB \subseteq T_{Pred}(U)$ is a
  derivation $l_1,\ldots,l_n$ of $G$ such that:

  \begin{enumerate}
  \item truthfulness: for all $i$, $\refr(\idsem(l_i)) \subseteq SKB$;
  \item communicative goal achieved: $C \subseteq \cup_{k=1}^n \refr(\idsem(l_k))$;
  \item unique reference: for all $1 \leq i \leq n$ and $r \in
    \roles(l_i)$, $|| \cup_{k=1}^n \idsem(l_k) ||_{HKB}(\id(i,r)) =
    \{\refr(\id(i,r))\}$.
  \end{enumerate}

  The \emph{sentence generation problem} of RGGs is to decide for a
  grammar $G$ and the communicative goal, target referent, and
  knowledge bases, whether there is a successful derivation.
\end{definition}

Compared to a system like SPUD \citep{Stone2003a}, the mapping from
the semantic roles in the lexicon entries to individuals in the
universe is relatively indirect: The semantic roles are first mapped
to variables by the $\id$ function, and then the variables are
assigned individuals by $\refr$. The point about this two-step process
is that it makes it possible to specify the ``unique reference''
condition. By mapping the roles to variables first, we can ask for the
set of all variable assignments that satisfy the combined semantic
representations of all lexicon entries (given the hearer's
knowledge). We can then convince ourselves that the information in the
semantic representations was precise enough that the hearer can
identify the referent uniquely by checking that all these variable
assignments assign the same individual to the variable for the
referring expression, and that this individual is indeed the target
referent. If we had mapped roles to individuals directly, rather than
distinguishing $\refr$ and $\i$, we could have verified that the
assignment satisfies all the semantic representations, but we couldn't
have stated that it is the \emph{only} such assignment.

\subsection{An example} \label{sec:rgg-example}

Let's go through an example to illustrate these definitions.

Consider the following grammar, which is defined over the nonterminals
$\{S, NP, Det, Adj\}$, start symbol $S$, roles $\{\self,
\mathsf{subj}, \mathsf{obj}\}$, and the following lexicon
entries:\footnote{I assume a predicate ``$\cdot = \mathsf{peter}$'' of
  which  $\self = \mathsf{peter}$ is an instance.}

$$\begin{array}{lp{1cm}l}
S \rightarrow \mathrm{likes}(NP \Neq \mathsf{subj}, NP \Neq
\mathsf{obj}) 
&&
Det \rightarrow \mathrm{the} \\
\mathsf{sem} = \{ \mathsf{likes}(\self, \mathsf{subj}, \mathsf{obj})
\} 
&&
\mathsf{sem} = \emptyset \\
&&\\
%
%
NP \rightarrow \mathrm{Peter} 
&&
Adj \rightarrow \mathrm{white} \\
\mathsf{sem} = \{ \self = \mathsf{peter} \}
&&
\mathsf{sem} = \{ \mathsf{white}(\self) \} \\
&&\\
%
%
NP \rightarrow \mathrm{rabbit}(Det \Neq \self, Adj \Neq \mathsf{self})
&&
Adj \rightarrow \epsilon \\
\mathsf{sem} = \{ \mathsf{rabbit}(\self) \}
&&
\mathsf{sem} = \emptyset
\end{array}
$$

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{pic-rdg-crisp}
  \caption{Verbalization of the example. Each node displays its
    terminal, nonterminal, and $\id$ function, in this order. $\refr$
    is $\{x \mapsto e, y \mapsto p, z \mapsto r_1\}$.}
  \label{fig:rdg-crisp}
\end{figure}

There are several different derivations for this grammar. Identifying
words with lexicon entries, we can write two of these as:
$$
\begin{array}{lll}
  d_1 &= &(\mathrm{likes}, \mathrm{Peter}, \mathrm{rabbit},
  \mathrm{the}, \mathrm{white}) \\
  d_2 &= &(\mathrm{likes}, \mathrm{Peter}, \mathrm{rabbit},
  \mathrm{the}, \epsilon) 
\end{array}
$$

In each of these cases, we can assume an $\id$ function that looks as
follows:
$$\begin{array}{ll|l}
i&r&\id(i,r) \\\hline
1&\self&x_1 \\
1&\mathsf{subj}&x_2 \\
1&\mathsf{obj}&x_3 \\
2&\self&x_2 \\
3&\self&x_3 \\
4&\self&x_3 \\
5&\self&x_3
\end{array}
$$

The tree represented by the derivation $d_1$ is shown in
Fig.~\ref{fig:rdg-crisp}.  The values of $\refr$ are irrelevant at
this point.

Now let's determine which of these derivations is
\emph{successful}. We assume the following knowledge bases, the
communicative goal $C=\{\mathsf{likes}(e,p,r_1)\}$, and the target
referent $e$.

$$\begin{array}{lll}
  SKB &=
  &\{\mathsf{likes}(e,p,r_1), p = \mathsf{peter}, \mathsf{rabbit}(r_1),
  \mathsf{rabbit}(r_2), \mathsf{white}(r_1)\} \\
  HKB &= &\{p = \mathsf{peter}, \mathsf{rabbit}(r_1),
  \mathsf{rabbit}(r_2), \mathsf{white}(r_1)\} 
\end{array}
$$

Then both $d_1$ and $d_2$ are truthful and convey the communicative
goal for $\refr$ being $\{x_1 \mapsto e, x_2 \mapsto p, x_3 \mapsto
r_1\}$. However, the reference in $d_2$ is not unique: $||\cup
\idsem(l_k)||(x_3)$ is $\{r_1,r_2\}$, i.e.\ the hearer might mistake
the object of the liking as the rabbit $r_2$. Therefore $d_2$ is not
successful. On the other hand, the unique reference condition is
satisfied in $d_1$, and so it is a successful derivation. Thus we can
read off the sentence ``Peter likes the white rabbit'' as a correct
verbalization of $C$ from Fig.~\ref{fig:rdg-crisp}.

\section{RTG generation as planning} 
\label{sec:rtg-generation-as-planning}

It is straightforward to extend the encoding of TAG sentence
generation to planning \citep{KolSto07} to regular tree grammars.
Notice that this makes the \citet{KolSto07} encoding a special case of
the algorithm presented here, as the derivation trees of a given TAG
grammar form regular tree languages
\citep{schmitz08:_featur_in_tag_deriv_trees} and can therefore be
described by RTGs.\footnote{This is not strictly speaking true,
  because we have not shown how to represent the pragmatic conditions
  and semantic requirements in an RGG, in order to keep the
  presentation simple; but it is obvious how these could be added.}

Let $l$ be a lexicon entry with production rule $A \rightarrow
\sigma(B_1 \Neq r_1, \ldots, B_n \Neq r_n)$.  Then the grammaticality
condition is encoded by the precondition and effects of a planning
operator as follows:\footnote{I write $r(u)$ to indicate the
  construction of a new symbol representing the $r$-child of $u$. I
  take $\self(u) = u$ as a special case. This cannot be expressed in
  PDDL, but can be encoded with the same planning-step-counting
  construction as in the ACL paper.}

\strut\\
\action{$\sigma$/$L$}{u}
{\open(u,A)}
{\neg \open(u,A), \open(r_1(u),,A_1), \ldots, \open(r_n(u),A_n)}\\

Notice that I write $\open$ to encode the unprocessed nodes in the
tree we are constructing, rather than $\mathsf{subst}$ and
$\mathsf{adj}$ as in the paper. Those were TAG-specific predicates,
encoding TAG's substitution and adjunction operation, and are no
longer necessary here. Effectively, what happens here is that the plan
encodes a derivation of the underlying RTG grammar; an atom
$\open(u,A)$ in the current planning state encodes the fact that the
current partial RTG derivation contains a leaf $u$ that is labeled
with the nonterminal symbol $A$ (and still needs to be expanded in the
future).

The translation of the semantic and reference conditions then proceeds
essentially as in the paper.  However, one crucial difference is that
we no longer distinguish between substitution and adjunction, and thus
it is not completely obvious where distractors for new referring
expressions get introduced. This used to happen exactly at the
substitution nodes. Now we can get the same effect by introducing
distractors only for those nonterminal occurrences on the right-hand
side that are labeled with a semantic role that is not $\self$.

Note that the encoding of RGG generation as a planning problem is not
entirely correct.  The difference is that in the problem as defined in
Section~\ref{sec:gener-probl-regul}, we took $a$ to be a possible
distractor for a new referring expression if there was any tuple
$(a_1,\ldots,a,\ldots,a_n)$ that satisfied the semantic condition of
the lexicon entry that introduced the RE.  In the encoding into
planning in the ACL paper, however, we only accept $a$ as a possible
distractor if $a_1,\ldots,a_n$ are the intended referents for the
other semantic roles in the lexicon entry. If we continue using this
encoding here, we end up with a planning problem that generates too
few distractors, and therefore can compute referring expressions that
are not unique. This is something that needs to be fixed at some
point, but for now the planning encoding seems to work pretty well in
practice.

\section{NP-completeness of the RTG generation problem}
\label{sec:np-completeness}

Finally, we can show that the sentence generation problem of RGGs is
NP-complete. This means that it is combinatorially hard, and hence
solving it with a worst-case exponential-time algorithm like a planner
is appropriate.

The NP-hardness proof below works by reducing the HAMILTONIAN-CYCLE
problem \cite{garey79:_comput_and_intrac} to the sentence generation
problem. HAMILTONIAN-CYCLE is the (NP-complete) problem of deciding
whether a directed graph has a Hamiltonian cycle.  A Hamiltonian cycle
is a path $(u_1,u_2,\ldots,u_n,u_1)$ in the graph such that the graph
contains an edge $(u_i,u_{i+1})$ for every $1 \leq i <n$ and the edge
$(u_n,u_1)$ -- i.e., it is a cycle --, and such that the cycle visits
every node in the graph exactly once. In the example graph in
Fig.~\ref{fig:npc}a, $(1,3,2,1)$ is a Hamiltonian cycle; $(1,2,1)$ is
a cycle that is not Hamiltonian because it doesn't visit 3.

\begin{prop} \label{prop:npc} The sentence generation problem of
  regular generation grammars is NP-complete.
\end{prop}
\begin{proof}
  It is obvious that the problem is in NP; once we have
  nondeterministically guessed a candidate tree, we can test whether
  it satisfies all conditions in polynomial time.

  For the NP-hardness, let's say we have a graph $G =
  (\{1,\ldots,n\},E)$ with $m$ edges of the form $(i,k)$ of which we
  want to check whether it has a Hamiltonian cycle.  We construct a
  regular generation grammar $G_G$ with nonterminals $N = \{S,
  A_{11}\} \cup \{A_{ik} \;|\; 2 \leq i \leq n+1, 1 \leq k \leq n\}$,
  start symbol $S$, terminal alphabet $\Sigma=\{f,a\}$, role alphabet
  $R = \{\mathsf{r}, \self\}$, and the following lexicon entries:

  $$\begin{array}{l}
    S \rightarrow f(A_{11} \Neq \mathsf{r}) \\
    \mathsf{sem} = \{ \mathsf{start}(\self,\mathsf{r})\} \\
    \\
    A_{ij} \rightarrow f(A_{(i+1)k} \Neq \self) \quad \mbox{for all
      $(j,k) \in E$ and $1 \leq i \leq n$} \\ 
    \mathsf{sem} = \{ \mathsf{neq}_j(\self) \} \\
    \\
    A_{(n+1)1} \rightarrow a \\
    \mathsf{sem} = \emptyset
  \end{array}
  $$

  Now the target referent $e$ and the communicative goal $\emptyset$
  have a successful derivation given $G_G$ and the knowledge bases
  $SKB = HKB = \{\mathsf{start}(e,1)\} \cup \{ \mathsf{neq}_i(k) \;|\;
  1 \leq i,k \leq n \;\mbox{and}\; i \neq k \}$ iff $G$ has a
  Hamiltonian cycle.  To see this, notice that any tree generated by
  $G_G$ must be a linear tree with nonterminal symbols $S, A_{11},
  A_{2u_2}, \ldots, A_{(n+1)1}$ from top to bottom, and for any two
  subsequent nonterminals $A_{ij}$ and $A_{(i+1)k}$ in this sequence,
  there must be an edge $(j,k)$ in the graph.  So the tree encodes a
  cycle of length $n$.  Now, if any node in the graph, say $j$, was
  not visited on this cycle, then the semantics in the subtree of the
  referring expression with nonterminal $A_{11}$ does not contain the
  atom $\mathsf{neq}_j(1)$.  In other words, the tree does not contain
  sufficient information to distinguish 1 (the target referent of the
  referring expression) from $j$, so the referring expression is not
  unique, and the derivation is not successful.  Hence, the cycle
  visits every node exactly once and is therefore Hamiltonian.
\end{proof}

\begin{figure}
  \centering
  \begin{tabular}{cp{2cm}c}
  \includegraphics[scale=0.5]{pic-npc-graph}
  &&
  \includegraphics[scale=0.5]{pic-npc} \\
  (a) && (b)
  \end{tabular}
  \caption{(a) A graph with a Hamiltonian cycle, and (b) a tree
    encoding this cycle according to the construction in
    Prop.~\ref{prop:npc}.} 
  \label{fig:npc}
\end{figure}

An example of this construction (i.e., a graph together with the tree
that encodes its Hamiltonian cycle and that is generated by the RGG)
is shown in Fig.~\ref{fig:npc}. Notice that nothing in this
construction particularly hinges in the generalization of the TAG
generation problem to RGGs. The proof can thus be ported directly to a
proof of the NP-completeness of the TAG sentence generation problem.

\section{Conclusion}

\section*{Version History}

\begin{tabular}{lll}
  v 3.0 & \today & Completely revised \\
  v 2.0 & 29/10/07 & Version for Hector Geffner
\end{tabular}

\bibliographystyle{plainnat}
\bibliography{gen}

\end{document}