### 2009-12-09 trunk/Papers/crisp-ff-10/crisp.tex
ICAPS submission for CRISP as challenge problem

\section{Sentence generation as planning}
\label{sec:crisp}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: