### 2009-12-09 trunk/Papers/reasoning-challenges/abstract.tex
ICAPS submission for Reasoning Challenges 2.0

\begin{abstract}
\begin{quote}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following 
instructions. 
\end{quote}
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: