### 2009-12-09 trunk/Papers/reasoning-challenges/conclusion.tex

\section{Conclusion} \label{sec:conclusion}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: