### 2009-12-11 trunk/Papers/crisp-ff-10/experiments.tex

\section{Experiments} 
\label{sec:experiments}

.. describe before/after picture, claim that now the performance is
``for the moment ok for this application'', but not for foreseeable
future (bigger grammers etc)

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: