### 2009-12-11 trunk/Papers/crisp-ff-10/abstract.tex
abstract revised

\begin{abstract}
We present a planning domain that encodes the problem of generating
natural language sentences. This domain has a number of features that
make it significantly different from existing IPC benchmark
domains. In particular, we show how several techniques used in the FF
planner, which has been highly successful on IPC problems, are led
astray. We discuss possible fixes to FF's shortcomings, some of which
we have already implemented, but some of which pose interesting
challenges to the AI Planning community. The performance of our
modified FF is up to three orders of magnitude better than that of the
original FF. While this makes FF much more practical for this
application, we show that it is still very limited with respect to
problem parameters such as the complexity of the underlying
grammar. We outline how other planning domains might share the
features that make the generation domain difficult. {\bf do we? how?}
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: