### 2009-12-11 trunk/Papers/crisp-ff-10/main.tex
author order

\documentclass[letterpaper]{article}
\usepackage{aaai}

\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\usepackage{graphicx}

\pdfinfo{
/Title (Formatting Instructions for Authors)
/Subject (Proceedings of the AAAI Press Conference)
/Author (AAAI)}

\setcounter{secnumdepth}{0}

\begin{document}
\title{Using Planning to Generate Natural-Language Sentences:\\
  Experiences and Challenges}

\author{
Alexander Koller \\ Saarland University, Germany \\ koller@mmci.uni-saarland.de
\And
Ronald Petrick \\ University of Edinburgh, UK \\ rpetrick@inf.ed.ac.uk
\And
J\"org Hoffmann \\ INRIA, France \\ joerg.hoffmann@inria.fr
}

\maketitle

\input abstract
\input introduction
\input crisp
\input crisp-ff
\input experiments
\input conclusion

\bibliography{bibliography}
\bibliographystyle{aaai}

\end{document}