### 2009-12-13 trunk/Papers/reasoning-challenges/abstract.tex
changes to abstract, introduction

\begin{abstract}
We revisit the structure of the International Planning Competition (IPC)
and propose a new ``rolling'' competition, as a permanent fixture for the
planning community. A central competition server would continually evaluate
planners and generate performance data, while members of the community
could directly upload new planners and problem domains, with the server
automatically reevaluating existing planners as necessary. Real-time
results would continuously be available, while more formal competitions
could be ``run'' at regular intervals by producing snapshots of current
planners and problems. Researchers would have access to a central
repository with the latest domains and performance statistics, running on a
common hardware platform, plus a facility for related activities such as
posting challenges and demonstrating new technologies. Since much of the
proposed infrastructure is not tied to a specific research area, this
approach offers the possibility of joint ventures with other research
communities.
\end{abstract}