### 2009-12-13 trunk/Papers/reasoning-challenges/infrastructure.tex

\section{Infrastructure}
\label{sec:infrastructure}

Talk a little bit about the current planning competition and (online)
competitions from other research areas (GIVE, CP?, others?)... Outline the
infrastructure we're proposing. We need some references here...