### 2009-12-16 trunk/Papers/reasoning-challenges/conclusion.tex
Some more details on infrastructure, some references (not cited yet)

\section{Conclusion}
\label{sec:conclusion}

\subsubsection*{Acknowledgements}
The original idea to organize a ``rolling'' competition in the CP community is due to Christian Schulte and Peter Stuckey.