### 2009-12-18 trunk/Papers/crisp-ff-10/main.tex

\documentclass[letterpaper]{article}
\usepackage{aaai}

\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\usepackage{graphicx}

\newcommand{\todo}[1]{\textbf{(#1)}}
\usepackage{url}

\pdfinfo{
/Title (Using Planning to Generate Natural-Language Sentences)
/Subject (Proceedings of ICAPS, hopefully)
/Author (Anonymous authors)}

\setcounter{secnumdepth}{0}

\begin{document}
\title{Using Planning to Generate Natural-Language Sentences:\\
  Experiences and Challenges}

\author{
Alexander Koller \\ Saarland University, Germany \\ koller@mmci.uni-saarland.de
%\And
%Ronald Petrick \\ University of Edinburgh, UK \\ rpetrick@inf.ed.ac.uk
\And
J\"org Hoffmann \\ INRIA, France \\ joerg.hoffmann@inria.fr
}

\maketitle

\input abstract
\input introduction
\input crisp
\input crisp-ff
\input experiments
\input conclusion

\bibliography{bibliography}
\bibliographystyle{aaai}

\end{document}