### 2009-12-18 trunk/Papers/crisp-ff-10/introduction.tex
revision of text reg ff and challenges

\section{Introduction} \label{sec:introduction}

Natural language generation (NLG) -- the problem of computing a
sentence or text that communicates a given piece of information -- has
a long-standing connection to automated planning
\cite{perrault80,appelt:planning}. In the past few years, the idea has
been picked up again in the NLG community, both to use planning as an
approach for modeling language problems
\cite{Steedman-Petrick:07,benotti08b} and to tackle the search
problems in existing NLG problems by using techniques from planning
\cite{KolSto07}.

Existing planners are relatively good at solving modest instances of
the planning problems that arise in this context. However, Koller and
Petrick \shortcite{koller10:_exper_with_plann_for_natur_languag_gener}
recently showed that as the planning problem instances from Koller and
Stone \shortcite{KolSto07} scale up, off-the-shelf planners become too
slow to be practically useful. They noted that there are major sources
of inefficiency both in the search itself and in the commonly used
preprocesses, which are put under pressure by several features of the
domain (e.g.\ comparatively large predicate and operator arities) that
diverge from most IPC benchmarks.
%the sheer number of
%ground instances of facts and actions that come up in this domain.

In this paper, we revisit this situation, analyze the sources of
inefficiency in more detail in the context of the FF planner
\cite{HoffmannNebel01}, and show how FF can be modified to perform
dramatically better. We achieve this by tweaking the implementation of
the FF preprocessor, and by acting on some simple but shocking
observations regarding the behavior of some of FF's search techniques,
in particular its goal ordering technique and the relaxed plan
heuristic. The changes we make are all simple, but make FF perform
well enough to be useful in practical NLG applications.

%the enforced hillclimbing and best-first search strategies on the NLG
%domain.  Neither of these changes degrades FF's performance on other
%domains.

%Nevertheless

Our performance gains with simple FF configuration changes certainly
point out that automatic planner configuration may be quite
useful. Apart from that, our domain remains an interesting challenge
for the development of better heuristics. There still are realistic
instances of the NLG planning domain on which our modified FF becomes
too slow for practical use. Overcoming this weakness would entail
overcoming the observed weaknesses of FF's search techniques --
variants of which are used in many planners today.

%As FF is still a competitive
%planner, this suggests that NLG might be interesting as a future
%challenge domain. 
%
% In particular, we show that by modifying the
%linguistic grammar from which we generate, we can vary the planning
%problems quite widely and emphasize different aspects that make our
%domain hard.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: