### 2009-12-18 trunk/Papers/crisp-ff-10/abstract.tex
shorter, new title

\begin{abstract}
We present a planning domain that encodes the problem of generating
natural language sentences. This domain has a number of features that
provoke fairly unusual behavior in existing planners. In particular,
we show how several techniques used in the FF planner, which has been
highly successful on IPC problems, are led astray. We discuss possible
fixes to FF's shortcomings, some of which we have already implemented,
but some of which pose interesting challenges to the AI Planning
community. The performance of our modified FF is up to several orders
of magnitude better than that of the original FF. While this makes FF
much more practical for this application, we show that it is still
quite limited with respect to the complexity of the underlying
grammar.
%We outline how other planning domains might share the
%features that make the generation domain difficult. {\bf do we? how?}
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: