### 2009-12-18 trunk/Papers/reasoning-challenges/levels.tex

\section{Levels of community involvement} \label{sec:levels}

We present our proposal for a new planning competition infrastructure
in terms of an ascending scale of levels, each of which makes it
easier or more rewarding for the community to participate in the
competition process.  As we will discuss below, we believe that the
competition format we propose is applicable in fields beyond planning
as well.  Therefore we will talk about \emph{systems}, \emph{problem
  classes}, and \emph{problem instances} instead of planners, planning
domains, and problems.

\subsection{Level 0: The current IPC}

As background and starting point of our proposal, consider the process
that underlies the current IPC competition.  Every two years, the IPC
committee appoints one or two organizers for each track of that year's
IPC competition.  These organizers assemble a number of problem
classes and instances, by developing them themselves or soliciting
them from the IPC committee or the community at large; sometimes this
requires extending the Planning Domain Description Language (PDDL,
\cite{PDDL}) to more advanced classes of problems.  After the
participants have developed their planners, they submit the source
code to the organizers, who then compile and run the planners on the
test machines.  Finally, the organizers prepare and present a report,
and the best-performing systems win awards.

\todo{Ron, can you fix this so it becomes true? - ak}

\subsection{Level 1: Collecting problems}

While this format for the IPC has been highly successful in the past
ten years, there is also a number of ways in which it can be improved
and streamlined.  One such direction concerns the way in which the
problem classes on which an IPC runs are collected.  There is no doubt
that the selection of IPC domains is a fundamental driving force for
the development of new planning algorithms.  Unfortunately, the
current process makes it difficult for users of planning technology,
who may come from a different community, to get a new
application-motivated domain accepted as an IPC benchmark.  This is
unsurprising, given that an IPC organizer must be able to justify
their selection of domains, and picking untested domains from other
communities is not the best way to do this.  However, this leads to a
selection process that is less dynamic and open to suggestions from
users than it could be.  A converse movement is that to keep the
number of domains that must be tested in each round of the competition
manageable, older domains are sometimes ``retired'', i.e.\ removed
from subsequent competitions.  This comes with a risk of a false sense
of progress, in which retired domains are remembered as ``solved''
although they were removed for completely independent reasons.
Finally, it has been argued \cite{ToughNuts} that some IPC organizers
tend to favor the development of new types of planning problems and
new dialects of PDDL over trying to find challenging benchmark
problems for existing versions of PDDL.

We propose that one way in which these problems can be alleviated is a
web repository in which any user can upload problem classes.  This
takes up Hoffmann's \shortcite{ToughNuts} idea to create a
community-driven repository for ``tough nuts'' -- i.e., instances that
were shown to be hard to solve in the past --, and takes it one step
further: In our proposal, \emph{all} IPC problem classes are suggested
by the community, much in line with the proposal by Roberts and Howe
\shortcite{roberts07:_call_for_partic}. This approach offers a
principled, publicly documented and accessible avenue for proposing
new problem classes, which benefits both application-oriented users of
planning technology and planning researchers who come up with tough
challenges for existing PDDL dialects. It also makes it easy to record
retired problem classes alongside more current ones, and makes it easy
to keep them accessible.

Submitting a problem class should involve three pieces of data. First,
the user must submit a collection of instances of the problem class or
a generator for such instances.  Second, the user must specify some
metadata including the class name and a human-readable description,
maybe with links to the application literature, complexity results,
and so on.  Third, the user must specify the \emph{type} of the
problem class, which determines what systems the data is suitable for
-- i.e., the language in which the problem is specified (such as a
particular dialect of PDDL or EUROPA \todo{cite}) and the type of
planning problem this is (propositional, temporal, uncertainty,
etc.).  If it is possible to convert between one specification
language and another automatically, this conversion could be performed
each time a problem is submitted.

Problem classes and instances can be submitted by any registered
user. In order to avoid abuse, the number of instances should be
limited (e.g.\ per hour), and a simple validity check should be
possible (such as running a parser if the format is PDDL).

\subsection{Level 2: Continuous evaluation}

But given that all the domains are available on a server, why stop
there?  We propose to make it possible to upload the \emph{planners}
to the same server as well, and then continuously and automatically
run the planners on the domains.  While the ability to upload domains
in Level 1 represents a conservative extension of the existing IPC
infrastructure, it is here that we depart more radically from the
existing format: Instead of a regular competition with organizers who
run the planner evaluation every other year, we propose a rolling
evaluation scheme, in which the current evaluation results would be
available online at any given time.

Of course, several issues need to be addressed to make this possible.
The first is how we can allow users to upload their systems and still
expect to run and maybe compile them automatically, without requiring
the manual intervention of an organizer if things break.  This problem
can be solved by imposing strict requirements on the execution
environment of the planner.  One way in which we could envision this
is by defining a single image for a virtual machine, such as
VirtualBox.\footnote{\url{http://www.virtualbox.org/}} When a planner
is evaluated, this happens in a fresh instance of this virtual
machine.  We can offer the virtual machine image for download; a
participant can then run the image on their own machine and ensure
that their planner works, in the exact environment in which it will be
evaluated.  This would even make it technically possible for
participants to submit not the source code, but a binary executable of
the planner (perhaps along with support files).  However, even if we
require the submission of source code, the move to a clean virtual
machine environment that contains only standard versions of tools and
libraries enforces at least a certain degree of portability from which
users will benefit when they download the planners.  Even when the
virtual machine image changes to reflect updates in the standard
development environments (new version of the C compiler, etc.), old
images could remain available, and so old planners would continue
running until the virtual machine emulator itself becomes unavailable
or incompatible.

We propose that every newly submitted planner is run automatically on
all available problem instances, and every newly submitted problem
instance is sent to all available planners.  The scheduling of this
process can be under the control of an automated system for continuous
integration, such as Hudson.\footnote{\url{http://hudson-ci.org/}}
Each run of a planner on a problem instance or problem class can take
place in a fresh instance of the virtual machine image.  The
evaluation infrastructure then inserts the problem data and the
planner into the virtual machine, runs the planner, measures
performance statistics (including which problems they could solve and
how much time this took them; see below), and stores the results in a
database.  Users will be able to view the results in dynamically
generated reports on the main website; there may also be functionality
to customize these reports to focus on certain types of problems.
Because all planners are run on all problem instances, problems in
which only a small and somewhat arbitrary selection of instances of a
certain problem are actually used in the evaluation \cite{pg2008002}
can be avoided.

In order to minimize security risks -- which arise when user-submitted
programs are being run automatically --, the individual virtual
machines need to be sandboxed quite tightly.  There is no need for an
individual virtual machine to be connected to any network, beyond
being able to enter performance measurements into a database.  By
further imposing resource limits on the number of available
processors, memory and processor time, and disk space, we believe the
potential for abuse is limited.  It can be limited further by only
allowing planner uploads to registered users.

While we see our proposed system's ability to generate up-to-date
reports automatically as a major strength, there may still be a
benefit in maintaining the current style of regular competitions at
discrete timepoints, e.g.\ to encourage the development of faster
planners through awards.  Such a competition could easily be based on
our system, by announcing a deadline and a specified subset of problem
classes. The selection of this subset could be informed, among other
things, by the report on problems that many planners can't
solve. Participants could then upload their planners at their leisure;
``running'' the competition amounts to taking a snapshot of the
performance reports on the selected problems when the deadline comes
and announcing the results.  By recording timestamps whenever a
performance measurements is made, this snapshot could remain available
on the website simply as one special report for a certain point in
time.  In such a case, the authors of planners should have the option
of marking them as ``non-competing'' to exclude them from the
competition snapshot, and maybe also from the up-to-date performance
reports.

\subsection{Level 3: Social networking}

As a final step, we justify the ``2.0'' in the title of our paper by
proposing the addition of mechanisms for social networking to the
continuous evaluation website.

The simplest such mechanism is to add Wiki pages to the evaluation
website.  This would allow the users who upload problem classes to
document these problems and link to existing literature; it would
allow the developers of planners to document these planners; and it
would make it easy for the organizers of regular competitions to add
more information to the snapshot reports.  The exchange of ideas and
opinions can be further facilitated by providing a discussion board
which links into these Wiki pages.

A second social networking tool which we believe would be useful in
the context of a planning evaluation website is to let users tag
problem classes and vote on them.  Any registered user can be allowed
to comment on any problem class and vote for how relevant, important,
or exciting they find it -- say, using a one to five star system as on
Amazon.  This has a number of crucial advantages.  First, an incoming
user or a competition organizer can take the average score as an
indication for the problems that are considered important in the
community.  Next to the difficulty of problems, this can be another
factor in the selection of problem classes for a competition snapshot.
Second, a relevance score for problem classes can also serve as a
guideline for the scheduling of planner runs: Assuming that there is
not enough computing power to run all incoming planners on all domains
in real time (e.g., just before a competition deadline), the runs on
popular domains can be prioritized.

\section{Implementation}

The obvious criticism that can be brought up against this proposal is
that of cost -- both in terms of implementing the necessary software
and in terms of running the hardware on which the continuous
evaluation is performed.

We believe that the actual hardware cost will be bearable.  While the
idea of running all planners on all domains sounds intimidating at
first, we don't anticipate that there will ever be more than a couple
dozen current versions of planners entered in the system.  This means
that any new domain that is entered will not be attempted to be solved
too many times.  Conversely, while there may (hopefully!) be a
significant number of domains in the system, and every new planner
must try to solve all domains whose types it claims to be able to
process, this cost can also be controlled by prioritizing popular or
hard domains.  A single computer should be sufficient for these needs;
if it turns out not to be, the virtual machines can easily be run on
machines anywhere.

implementing software

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: