### 2009-12-19 trunk/Papers/reasoning-challenges/levels.tex

\section{Levels of community involvement} \label{sec:levels}

We present our proposal for a new planning competition infrastructure
in terms of an ascending scale of levels, each of which makes it
easier or more rewarding for the community to participate in the
competition process. As we will discuss below, we believe that the
competition format we propose is applicable to fields beyond planning.  We
will therefore refer to \emph{systems}, \emph{problem classes}, and
\emph{problem instances}, instead of planners, planning domains, and
planning problems.

\subsection{Level 0: The current IPC}

As background, and a starting point for our proposal, consider the process
that underlies the current IPC. Every two years, the IPC committee looks
for volunteers to organize and run the various tracks of that year's
competition. These organizers are responsible for assembling a number of
problem classes and instances, by developing them themselves or soliciting
them from the wider community. Sometimes this process results in, or
coincides with, a new dialect of the Planning Domain Description Language
(PDDL, \cite{PDDL}), which describes for participants the supported
language for that year's competition. After the participants have developed
their systems, they submit the source code to the organizers, who then
compile and run the planners on the test machines.  Performance data is
collected and analysed.  Finally, the organizers prepare and present a
report, and the best-performing systems win awards.

\subsection{Level 1: Problem collection}

While this format for the IPC has been highly successful in the past
ten years, there is also a number of ways in which it can be improved and
streamlined.  One such direction concerns the way in which problem classes
are collected.  There is no doubt that the existing collection of IPC
domains is a fundamental driving force for the development of new planning
algorithms.  Unfortunately, the current selection process makes it
difficult for users of these systems, who may come from different
communities, to get a new application-motivated domain accepted as an IPC
benchmark.  This is unsurprising, given that IPC organizers must be able to
justify their selection of domains, and picking untested domains from other
communities is not the best way to do this.  However, this leads to a
selection process that is less dynamic than it could be, and more
restrictive for users wanting to contribute suggestions.  Conversely, to
keep the number of domains that must be tested in each round of the
competition manageable, older domains are sometimes ``retired'', i.e.\
removed from subsequent competitions.  This comes with a risk of a false
sense of progress, in which retired domains are remembered as ``solved''
although they were removed for completely independent reasons.  Finally, it
has been argued \cite{ToughNuts} that some IPC organizers tend to favor the
development of new types of planning problems and new dialects of PDDL over
trying to find challenging benchmark problems for existing versions of
PDDL. Such approaches run the risk of ``reinventing'' past problem classes
(albeit in different guises), or redirecting investigation away from
current challenges.

One way to alleviate these problems is by introducing a web repository
where users can upload problem classes. This revisits Hoffmann's
\shortcite{ToughNuts} idea to create a community-driven repository for
``tough nuts'' -- i.e., instances that were shown to be hard to solve in
the past -- and takes it one step further: in our proposal, \emph{all} IPC
problem classes are provided by the community, much in line with the
proposal by Roberts and Howe \shortcite{roberts07:_call_for_partic}. Such a
repository offers a principled, publicly documented and accessible avenue
for proposing new problem classes, which benefits both application-oriented
users of planning technology and researchers who invent tough challenges
for existing PDDL dialects. It also makes it easy to record retired problem
classes alongside active ones, and keep them all accessible.

Problem classes and instances can be submitted by any registered user. 
Submitting a problem class should involve three pieces of data. First,
the user should submit a collection of instances of the problem class or
a generator for such instances.  Second, the user should specify some
metadata including the class name and a human-readable description,
possibly with links to the applicable literature, complexity results, and
so on.  Third, the user should specify the \emph{type} of problem class,
which determines what systems the data is suitable for, i.e., the language
in which the problem is specified (such as a particular dialect of PDDL or
EUROPA \todo{cite}) and the type of planning problem (e.g., propositional,
temporal, uncertainty, etc.).  If it is possible to convert between one
specification language and another automatically, this conversion could be
performed each time a problem is submitted.
In order to avoid abuse, the number of instances should be limited (e.g.,
per hour), and a simple validity check should be possible (e.g., running a
parser to verify that the format is PDDL).

\subsection{Level 2: Continuous evaluation}

Given that all the domains are available on a server, why stop
there?  We propose to make it possible to upload the \emph{planners}
to the same server as well, and then continuously and automatically
run the planners on the domains.  While the ability to upload domains
in Level 1 represents a conservative extension of the existing IPC
infrastructure, it is here that we depart more radically from the
existing format: instead of a regular competition with organizers who
run the planner evaluation every other year, we propose a rolling
evaluation scheme, in which the current evaluation results would be
available online at any given time.

Of course, several issues need to be addressed to make this possible.
The first is how we can allow users to upload their systems and still
expect to run and maybe compile them automatically, without requiring
the manual intervention of an organizer if things break.  This problem
can be solved by imposing strict requirements on the execution
environment of the planner.  One way in which we could envision this
is by defining a single image for a virtual machine, such as
VirtualBox.\footnote{\url{http://www.virtualbox.org/}} When a planner
is evaluated, this happens in a fresh instance of this virtual
machine.  We can offer the virtual machine image for download; a
participant can then run the image on their own machine and ensure
that their planner works, in the exact environment in which it will be
evaluated.  This would even make it technically possible for
participants to submit not the source code, but a binary executable of
the planner (perhaps along with support files).  However, even if we
require the submission of source code, the move to a clean virtual
machine environment that contains only standard versions of tools and
libraries enforces at least a certain degree of portability from which
users will benefit when they download the planners.  Even when the
virtual machine image changes to reflect updates in the standard
development environments (new version of the C compiler, etc.), old
images could remain available, and so old planners would continue
running until the virtual machine emulator itself becomes unavailable
or incompatible.

We propose that every newly submitted planner is run automatically on
all available problem instances, and every newly submitted problem
instance is sent to all available planners.  The scheduling of this
process can be under the control of an automated system for continuous
integration, such as Hudson.\footnote{\url{http://hudson-ci.org/}}
Each run of a planner on a problem instance or problem class can take
place in a fresh instance of the virtual machine image.  The
evaluation infrastructure then inserts the problem data and the
planner into the virtual machine, runs the planner, measures
performance statistics (including which problems they could solve and
how much time this took them; see below), and stores the results in a
database.  Users will be able to view the results in dynamically
generated reports on the main website; there may also be functionality to
customize these reports to focus on certain types of problems.  Because all
planners are run on all problem instances, and users can ensure that
challenging problem instances are included in the competition, the chance
of systems being evaluated on only a small and somewhat arbitrary selection
of certain problems \cite{pg2008002} can be avoided.

In order to minimize security risks, which can arise when user-submitted
programs are being run automatically, the individual virtual
machines need to be sandboxed quite tightly.  There is no need for an
individual virtual machine to be connected to any network, beyond
being able to enter performance measurements into a database.  By
further imposing resource limits on the number of available
processors, memory and processor time, and disk space, we believe the
potential for abuse is limited.  It can be limited further by only
allowing planner uploads to registered users.

While we see our proposed system's ability to generate up-to-date
reports automatically as a major strength, there may still be a
benefit in maintaining the current style of regular competitions at
discrete timepoints, e.g., to encourage the development of faster
planners through awards.  Such a competition could easily be based on
our system, by announcing a deadline and a specified subset of problem
classes. The selection of this subset could be informed, among other
things, by the report on problems that many planners can't
solve. Participants could then upload their planners at their leisure;
``running'' the competition amounts to taking a snapshot of the
performance reports on the selected problems when the deadline comes
and announcing the results.  By recording timestamps whenever a
performance measurements is made, this snapshot could remain available
on the website simply as one special report for a certain point in
time.  In such a case, the authors of planners should have the option
of marking them as ``non-competing'' to exclude them from the
competition snapshot, and maybe also from the up-to-date performance
reports.

\subsection{Level 3: Social networking}

As a final step, we justify the ``2.0'' in the title of our paper by
proposing the addition of mechanisms for social networking to the
continuous evaluation website.

The simplest such mechanism is to add Wiki pages to the evaluation
website.  This would allow the users who upload problem classes to
document these problems and link to existing literature; it would
allow the developers of planners to document these planners; and it
would make it easy for the organizers of regular competitions to add
more information to the snapshot reports.  The exchange of ideas and
opinions can be further facilitated by providing a discussion board
which links into these Wiki pages.

A second social networking tool which we believe would be useful in
the context of a planning evaluation website is to let users tag
problem classes and vote on them.  Any registered user can be allowed
to comment on any problem class and vote for how relevant, important,
or exciting they find it -- say, using a one to five star system as on
Amazon.  This has a number of crucial advantages.  First, an incoming
user or a competition organizer can take the average score as an
indication for the problems that are considered important in the
community; imagine a ``top ten'' list of domains and planners on the
website's front page.  Next to the difficulty of problems, this can be
another factor in the selection of problem classes for a competition
snapshot.  Second, a relevance score for problem classes can also
serve as a guideline for the scheduling of planner runs: Assuming that
there is not enough computing power to run all incoming planners on
all domains in real time (e.g., just before a competition deadline),
the runs on popular domains can be prioritized.

\section{Evaluation measures}

The continuous evaluation infrastructure we just presented is
fundamentally agnostic to the type of concrete evaluation measures
that it collects.  In particular, traditional measures such as the
percentage of problems solved per system, average solution quality,
average runtime and memory use, and search tree sizes can be
collected -- with the caveat that the virtual machine must be
configured so as to allow comparable runtime measurements.
Furthermore, popularity scores from the social networking aspect can
be incorporated, or different measures could be combined into an
overall score.

\section{Implementation}

The obvious criticism that can be brought against this proposal is
that of cost -- both in terms of implementing the necessary software
and in terms of running the hardware on which the continuous
evaluation is performed.

We believe that the actual hardware cost will be bearable.  While the
idea of running all planners on all domains sounds intimidating at
first, we don't anticipate that there will ever be more than a couple
dozen current versions of planners entered in the system.  This means
that any new domain that is entered will not be attempted to be solved
too many times.  Conversely, while there may (hopefully!) be a
significant number of domains in the system, and every new planner
must try to solve all domains whose types it claims to be able to
process, this cost can also be controlled by prioritizing popular or
hard domains.  A single computer should be sufficient for these needs;
if it turns out not to be, the virtual machines can easily be run on
machines anywhere.

On the other hand, we expect that implementing the software for
running the continuous evaluation and presenting the results on a
website would not be a prohibitive effort either.  Most of the
components of the system (continuous integration server, virtual
machines, Wiki) already exist as open-source off-the-shelf
components.  We estimate that a prototype implementation of a complete
system could be built in a person-month of work; surely this is the
amount of work for which funding or volunteers could be found.

\section{Applicability and limitations}

One advantage of the evaluation infrastructure proposed above is that
its applicability is not limited to the evaluation of planners.
%RP: I don't see why this needs to be restricted to "reasoning" tasks
Many reasoning tasks, including constraint solving and theorem proving, are
being evaluated by running solvers over problem instances and reporting
performance measures.  At the level of abstraction at which our continuous
evaluation system could be implemented, there is no fundamental difference
between the problem type ``deterministic planning problem in PDDL 1.0'' and
the problem type ``\todo{something from CP or similar}'' -- in both cases,
the system can collect problem classes and instances of this type and run
them through solvers that declare themselves to accept such inputs.  This
means that there is a potential for sharing development and maintenance
costs for the evaluation platform among different communities.

However, there are of course limitations to the type of problem for
which it makes sense to run a community-driven continuous evaluation.
For one thing, it must be acceptable in the community that every
system runs on the same hardware and software configuration, as
implemented in the virtual machine images.  On the other hand, our
system makes most sense where the ratio of problem classes to systems
is relatively high.  Neither of the two is true, for instance, for
most natural language processing competitions: different research
groups tend to use a variety of programming languages and development
environments to build their systems, and there are typically many more
systems in any competition than there are problem classes (corpora).
However, our approach is completely consistent with the requirements
of a community where solving a single hard problem instance for the
first time counts as an achievement, such as in theorem proving.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: