### 2009-12-19 trunk/Papers/crisp-ff-10/introduction.tex

\section{Introduction} \label{sec:introduction}

Natural language generation (NLG) -- the problem of computing a
sentence or text that communicates a given piece of information -- has
a long-standing connection to automated planning \cite{perrault80}. In
the past few years, the idea has been picked up again in the NLG
community, both to use planning as an approach for modeling language
problems (e.g., \cite{Steedman-Petrick:07}) and to tackle the search
problems in NLG by using techniques from planning \cite{KolSto07}.

Existing planners are relatively good at solving modest instances of
the planning problems that arise in this context. However, Koller and
Petrick \shortcite{koller10:_exper_with_plann_for_natur_languag_gener}
recently showed that as the planning problem instances from Koller and
Stone \shortcite{KolSto07} scale up, off-the-shelf planners become too
slow to be practically useful. They noted that there are major sources
of inefficiency both in the search itself and in the commonly used
preprocesses, due to several features of the domain
(e.g.\ comparatively large predicate and operator arities) that
diverge from most IPC benchmarks.
%the sheer number of
%ground instances of facts and actions that come up in this domain.

In this paper, we revisit this situation, analyze the sources of
inefficiency in more detail in the context of the FF planner
\cite{HoffmannNebel01}, and show how to modify FF to obtain
dramatically better performance. We achieve this by tweaking the
implementation of FF's preprocessor, and by acting on some simple but
shocking observations regarding the behavior of some of FF's search
techniques, in particular its goal ordering technique and the relaxed
plan heuristic. The changes we make are all simple, but make FF
perform well enough to be useful in practical NLG applications.

%the enforced hillclimbing and best-first search strategies on the NLG
%domain.  Neither of these changes degrades FF's performance on other
%domains.

%Nevertheless

Our performance gains with simple FF configuration changes certainly
point out that automatic planner configuration may be quite
useful. Apart from that, our domain remains an interesting challenge
for the development of better heuristics. There still are realistic
instances of the NLG planning domain on which our modified FF is not
practical. Overcoming this weakness would entail overcoming the
observed weaknesses of FF's search techniques -- variants of which are
used in many planners today.

%As FF is still a competitive
%planner, this suggests that NLG might be interesting as a future
%challenge domain. 
%
% In particular, we show that by modifying the
%linguistic grammar from which we generate, we can vary the planning
%problems quite widely and emphasize different aspects that make our
%domain hard.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: