### 2010-03-01 trunk/Papers/crisp-ff-10/experiments.tex
Joerg final

\begin{figure*}[t]
  \centering
(a)
\vtop{
\vskip-2ex
\hbox{\includegraphics[width=0.8\columnwidth]{pics/xtag-k2-dist0}}
}
$\qquad\qquad$
(b)
\vtop{
\vskip-2ex
\hbox{\includegraphics[width=0.8\columnwidth]{pics/xtag-k2-dist2}}
}
\vspace{-0.5cm}
\caption{Runtimes for $d=0$ (a) and $d=2$ (b).}
  \label{fig:runtimes}
\vspace{-0.5cm}
\end{figure*}

\section{Experiments} 
\label{sec:experiments}

To evaluate the effect of our changes on the efficiency of the planner
on practical inputs, we ran a series of experiments using the XTAG
Grammar \cite{xtag01:_tr}, a large-scale tree-adjoining grammar of
English, from which we selected all lexicon entries for seven words we
used in the experiments. We set up inputs to generate sentences of the
form ``$S_1$ and $S_2$ and \ldots and $S_n$'', i.e.\ a conjunction of
$n$ sentences. Each sentence $S_i$ is of the form ``X admires Y''. X
and Y are unique referring expressions, such as ``the rich
businessman''. For each experiment, we have a parameter $d$ that
specifies how many adjectives are needed to distinguish the referent
uniquely -- in particular, $d=0$ means we can say ``the businessman'',
and $d=2$ means we must say ``the rich sad businessman''.
% JOERG: ``new encoding'' gibt's ja gar nicht mehr als unterscheidung
%
%We translate
%these generation problems into planning problems with the new encoding
%as shown in Fig.~\ref{fig:white-rabbit-as-planning}.

%\begin{figure}[t]
%  \centering
%  \includegraphics[width=\columnwidth]{pics/xtag-k2-dist2}
%  \caption{Runtimes for $d=2$.}
%  \label{fig:runtimes-k2-dist2}
%\end{figure}

Consider first the results for $d=0$, shown in Fig.~\ref{fig:runtimes}
(a); the lines end where the planner exceeded its five-minute
timeout. Notice that the vertical axis is logarithmic. Clearly, ``Old
FF'' (the original FF) is much too slow for practical use. The same is
true -- although to a less dramatic extent -- for LAMA, which already
takes three seconds to generate the relatively short sentence at $n=3$
and times out for larger instances. ``New preprocessor'' is FF with
only our fixes to FF's preprocess, ``New FF'' applies all fixes.
% The other two lines represent FF with the improved preprocessor,
%using the standard enforced hillclimbing strategy and the best-first
%search with helpful actions (``New FF'') respectively.  
Even the preprocessor fixes alone lead to a speed-up of at least $3$
orders of magnitude, leading to practically useful runtime -- for
instance, the 24-word sentence at $n=4$ is generated in under two
seconds. Note that, by contrast, the search configuration does not
make much of a difference here. This changes in our next experiment.

Results for $d=2$ are shown in Fig.~\ref{fig:runtimes} (b).  Here the
old FF preprocessor crashed with a segmentation fault even at
$n=1$. LAMA performs better, but times out beyond $n>2$.  Best-first
search with helpful actions is clearly the best-performing search
strategy here.  It takes about 14 seconds for $n=3$, but this sentence
is already quite long at 29 words, and EHC takes 3.5 minutes on the
same input.  Best-first search without helpful actions timed out at
$n=3$, illustrating the importance of that pruning technique.

We remark that the relative behavior of the different search methods
is not uniform in our domain, which exhibits a huge amount of
structural variance due to the many different forms that the input
grammar may take. We are far from an exhaustive exploration of all the
possible parameter settings. For instance, with particular inputs (no
distractors), EHC can be more effective than best-first search.

%Suffice it to say that, in particular encoding variants and
%with particular grammars (e.g.\ grammars not featuring any
%distractors), in stark contrast to Fig.~\ref{fig:runtimes} (b) EHC is
%much more effective than best-first search.

%While this brings the overall planner runtimes into a range where FF
%can now be useful in practical NLG applications, it is important to
%note that it still has serious limitations.  The best-first search
%timed out for any $n>4$ in either experiment.  While many NLG
%applications can get by with shorter sentences, improving the search
%on our domain is still an open question for the future.  Even within
%our domain, the relative quality of EHC and BFS+H depended on the
%exact generation problem instance: While BFS+H outperformed EHC at
%$d=2$, EHC was actually slightly faster at $d=0$ because the absence
%of distractors meant there was not much room for the characteristic
%underestimation of the relaxed plan evaluation discussed above.  This
%means that sentence generation remains as a varied and challenging
%domain for planning.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: