### 2010-05-19 trunk/Papers/pcrisp-10/conclusion.tex
Added draft for TAG+ paper about pcrisp

\section{Conclusion}
\label{sec:conclusion}

We describe an approach to sentence generation based on probabilistic tree adjoining
grammar and AI planning. Following previous work by \newcite{stonedoran1997,kollerstone2007}, we treat sentence 
generation as an integrated task comprising sentence planning and surface realization using declarative descriptions of the syntactic
and semantic contribution of each word and contextual constraints on its use. Our system uses AI planning technology to find the best {\sc tag} 
derivation that fulfills a given communicative goal, while satisfying these constraints. 
Our experiments show that our generation system based on {\sc pltag} and planning makes it possible to use an integrated approach to sentence generation with treebank induced grammars for the first time. The results also highlight the trade-off between output quality, efficiency and robustness. Future work work should optimize the smoothing model to boost output quality. Improved and/or tailored heuristics could help to improve runtime behavior. Finally it would be interesting to evaluate the system in an extrinsic way that also addresses sentence planning.