### 2010-05-19 trunk/Papers/pcrisp-10/introduction.tex

\section{Introduction}
Traditional sentence generation systems adopt a pipeline architecture that includes sentence planning and surface realization as two independent modules \cite{reiterdale2000}.  A different view is advocated by \newcite{stonedoran1997}. Their sentence generation system {\sc spud} allows a combined treatment of sentence planning and surface realization.
{\sc spud} exploits {\sc ltag}'s localization of argument structure and uses a top-down construction algorithm to incrementally builds both syntactic and flat semantic structure that satisfies the communicative goal and is consistent with semantic and pragmatic constraints. 
This unified approach to sentence generation is attractive because it tackles different phenomena in generation in an elegant way using a unified, linguistically well motivated framework.

While {\sc spud} uses an incomplete heuristic greedy search, to solve the {\sc NP-}complete problem of finding such a derivation \cite{kollerstriegnitz2002}, \newcite{kollerstone2007} present an encoding of this problem into an AI planning problem.

We expect that using larger grammars with this approach leads to problems in both performance and output quality, due to over-generation. 
Our work approaches this problem by combining integrated sentence generation as planning (following \newcite{kollerstone2007}), numeric AI planning and a probability model ({\sc pltag}, \newcite{resnik1992}).

{\bf Structure of the paper:}
In section \ref{sec:related} we discuss related work. Section \ref{sec:crisp} reviews the original generation as planning approach ({\sc crisp}) suggested by \newcite{kollerstone2007}. We describe our statistical extension of this approach ({\sc pcrisp}) in section \ref{sec:pcrisp} and evaluate our new system in section \ref{sec:experiments}. Finally we conclude this paper in section \ref{sec:conclusion}.