### 2010-05-19 trunk/Papers/pcrisp-10/pcrisp-10.tex

%
% File naaclhlt2010.tex
%
% Contact: nasmith@cs.cmu.edu

\documentclass[11pt,letterpaper]{article}
\usepackage{naaclhlt2010}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

\newlength{\rhwidth}
\newcommand{\planaction}[3]{
\begin{minipage}{.5\textwidth}
\footnotesize
\setlength{\rhwidth}{.5\textwidth}
\addtolength{\rhwidth}{2.0cm}
{#1:}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Precond:}\end{minipage}\begin{minipage}[t]{\rhwidth} #2 \end{minipage}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Effect:}\end{minipage} \begin{minipage}[t]{\rhwidth}\ #3 \end{minipage}
\end{minipage}
}

\newcommand{\cplanaction}[4]{
\begin{minipage}{.5\textwidth}
\setlength{\rhwidth}{.5\textwidth}
\addtolength{\rhwidth}{2.0cm}
{#1:}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Precond:}\end{minipage}\begin{minipage}[t]{\rhwidth} #2 \end{minipage}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Effect:}\end{minipage} \begin{minipage}[t]{\rhwidth} #3 \end{minipage}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Cost:}\end{minipage} \begin{minipage}[t]{\rhwidth} \ensuremath{#4} \end{minipage}
\end{minipage}
}

\title{Sentence Generation as Planning with Probabilistic LTAG\\
%\Thanks{}
}

\author{Daniel Bauer\\
  Columbia University\\
  New York, NY, USA\\
  {\tt bauer@cs.columbia.edu} 
  \And
  Alexander Koller\\
  Saarland University\\
  Saarbr\"{u}cken, Germany\\ 
  {\tt koller@mmci.uni-saarland.de}}

\date{}

\begin{document}
\maketitle
\begin{abstract}
Our work merges combined sentence planning and surface realization using 
{\sc ltag} and statistical generation. This makes it possible to address
integrated sentence generation with large, treebank induced grammars for
the first time.
\end{abstract}

\input{introduction}
\input{related} % Related work
\input{crisp} % Sentence Generation as Planning
\input{pcrisp} % Statistical Generation as Planning
\input{experiments} % Experiments
\input{conclusion}

\bibliographystyle{naaclhlt2010}
\bibliography{literature}

\end{document}