### 2010-05-19 trunk/Papers/pcrisp-10/pcrisp.tex

\section{Statistical Generation as Planning}
\label{sec:pcrisp}

We now extend {\sc crisp} to statistical generation ({\sc pcrisp}). The basic idea is to add a statistical grammar model while leaving the sentence generation mechanism untouched. This way we can select the highest scoring derivation, that satisfies all constraints (grammaticality, expresses the communicative goal, use unambiguous refering expressions etc.). We first propose three probability models in the next section and adapt {\sc crisp} planning problems to acomodate probabilities in section \ref{ssec:pcrisp-domains}.

\subsection{Probability Models}
\label{ssec:probmodels}

{\bf Probabilistic TAG:}
As a basic probability model over {\sc ltag} derivations we use probabilistic {\sc tag}\footnote{we refer to this formalism as \sc{pltag}} \cite{resnik1992}, 
which provides a probability for each substitution and adjunction event in a derivation. In addition, it assigns a probability to the event of starting a derivation with a specific initial tree and to the event of not adjoining anything to an open adjunction node at all. Assuming that the probabilities of these events within a derivation are independent of each other, the probability of a derivation is then simply defined as the product of all elementary event probabilities.
Formally,

\newcommand{\adj}[0]{\textit{ adj}}
\newcommand{\init}[0]{\textit{ init}}
\newcommand{\subst}[0]{\textit{ subst}}
$$ \sum\limits_{\alpha \in I}\sum\limits_{lex \textit{ of } \alpha} P_{init}(\init(\alpha, lex)) = 1 $$

$$\sum\limits_{\alpha \in I}\sum\limits_{lex \textit{ of }\alpha} P_{subst}(\subst(\tau,lex_\tau, \alpha, lex_\alpha, n)) = 1,$$
for each $\tau \in I \cup A$, each lexicalization $lex_\tau$  of $\tau$ and for each substitution node $n$ in $\tau$.

$$ P_{adj}(\adj(\tau, lex_\tau, \textit{None},n)) +$$
$$ \sum\limits_{\beta \in A}\sum\limits_{lex \textit{ of }\beta} P_{adj}(\adj(\tau,lex_\tau, \alpha, lex_\alpha, n)) = 1,$$ for each $\tau \in I \cup A$, each lexicalization $lex_\tau$  of $\tau$ and for each node $n$ that is open for adjunction in $\tau$.

Our choice of {\sc pltag} for sentence generation is motivated by a number of attractive properties.
 First, as {\sc pltag} is lexicalized, it does not only assign probabilities to operations in the grammar (as for example plain {\sc pcfg}), but also accounts for binary dependencies between words. Unlike n-gram models however, these co-occurences are structured according to local syntactic context as a result of {\sc tag}'s extended domain of locality. The probability model describes how the syntactic arguments of a word are typically filled. 
Furthermore, as {\sc tag} factors recursion from the domain of dependencies, the probability for core constructions remains the same independent of additional adjunctions. 
Second, {\sc pltag} is a generative model. This is crucial because it allows us to assign probabilities to partial derivations. In addition we use the probability of partial derivations as a cost function, to help guide search through the space of possible PLTAG derivations. In contrast, other statistical generation systems use discriminative models that select the best possible generation output from a set of candidate sentences.
Finally, the independence assumption makes it easy to estimate {\sc pltag}s from a treebank.

To alleviate data sparseness, caused by the specific event definition of {\sc pltag}, we propose two alternative probability models. The first model is an unlexicalized version of {\sc pltag}, the second model uses a simple smoothing approach.\\ 
{\bf Unlexicalized Probabilities:}
  An easy way to deal with data sparseness is to drop all lexicalization from event definitions, as illustrated in figure \ref{fig:modelillustration}, A. Unfortunately the model does no longer account for bilexical dependencies between words. Since our system has to add a lexicalized tree in each step, lexicalization for this child tree should always be taken into account by the probability model, if available. Despite these drawbacks, we perform experiments with the unlexicalized model as a baseline. This allows us to investigate if purely syntactic information is sufficient to achieve high quality generation output. \\  
{\bf Linear Interpolation:}
 The third model computes linear interpolation between three back-off levels. The first level is just standard {\sc pltag} (figure \ref{fig:modelillustration}, B.1), for the second level the lexicalization of the parent tree is dropped (figure \ref{fig:modelillustration}, B.2), for the third level the model describes only the distribution of lexicalized child trees over each category (figure \ref{fig:modelillustration}, B.3). 
The third level is similar in spirit to a probabilistic version of the original generation as planning formulation in \cite{kollerstone2007}.
\begin{figure}[t]
\begin{center}
\includegraphics[width=.6\textwidth]{figures/modelillustration}
\caption{\label{fig:modelillustration} Illustration of the unlexicalized probability model (A) and the three back-off levels of the linear interpolation model (B). B.1 corresponds to the original {\sc pltag} definition.}

\end{center}
\end{figure}

%It is obvious that the number of total planning operators grows quadratically with the number of possible input words in the grammar. In practice this causes a problem for most heuristic search planners, which will instantiate planning operators to all possible actions before attempting to solve the problem. Our generation system therefore only selects operators that are compatible with the input semantics before running the planner. 
%On the other hand,

\subsection{PCRISP Planning Domains}
\label{ssec:pcrisp-domains}
In this section we reformulate the {\sc crisp} planning operators described in section \ref{ssec:crispdomain}.
The independence assumption in {\sc pltag} allows us to maintain that planning operator add a single tree, but are now assigned a probability score. However, while {\sc crisp} planning operators can add an elementary tree to any site of the right category, {\sc pltag} events are binary events between lexicalized trees at a specific node. We therefore adapt the literals that record open substitution and adjunction sites in partial derivations accordingly and create one operator for each node in each possible combination of lexicalized trees.

Different approaches to planning with numeric values exist. We can encode probabilities either as timespans required by each operator and use a temporal planner, or as as generic numeric variable that is increased by each operator. In both cases this value (which can be seen as a cost) is summed over actions and a resulting plan should seek to minimize it. We therefore set the cost of an operator to be its negative log probability. 
Figure \ref{example-action} shows an example planning operator. We adopt this scheme for all planning operators regardless of the probability model.
\begin{figure}[t]
\begin{center}
\cplanaction{\bf subst-t3-cat-t28-eats-n2(u,~x1)}{step(step1),referent(u,~x1),\\ subst(t28-eats,n2,~u), cat(x1)}{
$\lnot$needtoexpress(pred-cat,~x1),\\ $\lnot$subst(t-28-eats, n2,~u),\\
adj(t3-cat, n0, u), adj(t3-cat, n1 u)\\
$\lnot$ step(step1), step(step2)}{4.3012}\\\smallskip
\caption{\label{example-action} {\sc pcrisp} Operator to substitute $t3$ lexicalized with 'cat' for node 2 of $t28$ lexicalized with `eats'.}
\end{center}
\end{figure}
%\newcommand{\init}[0]{\textit{ init}}
%\newcommand{\subst}[0]{\textit{ subst}}

%\begin{figure}[p]
%\caption{\label{modelillustration} The three back-off levels.}
%\begin{center}
%\includegraphics[width=.5\textwidth]{modelillustration}
%\end{center}
%\end{figure}