### 2010-05-19 trunk/Papers/pcrisp-10/related.tex

\section{Related Work}
\label{sec:related}

Previous statistical NLG systems (e.g. \newcite{knight1995}, {\sc nitrogen} \cite{langkildeknight1998}, {\sc openCCG} \cite{whitebaldridge2003}
and the {\sc ltag} based realizer {\sc fergus} \cite{bangalorerambow2000}), follow a generate-and-select strategy using n-gram models over output structures
for discrimination. But since this approach focuses on discriminating realization output for the same sentence plan only, it is incompatible with the
integrated treatment of sentence planning and realization.

One of the few systems uses generative models for sentence generation is {\sc pCru} \newcite{belz2008}. However, her model is defined on possible states of the generation system itself, not on the underlying grammar formalism.

$\cdots$