### 2010-05-20 trunk/Papers/pcrisp-10/abstract.tex

\begin{abstract}
We present PCRISP, a sentence generation system for probabilistic TAG
grammars which performs sentence planning and surface realization in
an integrated fashion, in the style of the SPUD system.  PCRISP
operates by converting the generation problem into a metric planning
problem and solving it using an off-the-shelf planner.  We evaluate
PCRISP on the WSJ corpus and identify trade-offs between coverage,
efficiency, and accuracy.
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "pcrisp-10"
%%% End: