### 2010-05-21 trunk/Papers/pcrisp-10/pcrisp-10.tex

%
% File naaclhlt2010.tex
%
% Contact: nasmith@cs.cmu.edu

\documentclass[11pt,letterpaper]{article}
\usepackage{naaclhlt2010}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{color}

\usepackage{url}

\setlength\titlebox{6.5cm}    % Expanding the titlebox

\newlength{\rhwidth}

\newcommand{\f}[1]{\mbox{#1}}
\newcommand{\adj}[0]{\textit{adj}}
\newcommand{\noadj}[0]{\textit{noadj}}
\newcommand{\init}[0]{\textit{init}}
\newcommand{\subst}[0]{\textit{subst}}

\newcommand{\pcrisp}{PCRISP}
\newcommand{\crisp}{CRISP}
\newcommand{\ltag}{LTAG}
\newcommand{\spud}{SPUD}

\newcommand{\todo}[1]{\textcolor{red}{(#1)}}

\newcommand{\planaction}[3]{
\begin{minipage}{.5\textwidth}
\footnotesize
\setlength{\rhwidth}{.5\textwidth}
\addtolength{\rhwidth}{2.0cm}
{#1:}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Precond:}\end{minipage}\begin{minipage}[t]{\rhwidth} #2 \end{minipage}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Effect:}\end{minipage} \begin{minipage}[t]{\rhwidth}\ #3 \end{minipage}
\end{minipage}
}

\newcommand{\cplanaction}[4]{
\begin{minipage}{.5\textwidth}
\footnotesize
\setlength{\rhwidth}{.5\textwidth}
\addtolength{\rhwidth}{2.0cm}
{#1:}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Precond:}\end{minipage}\begin{minipage}[t]{\rhwidth} #2 \end{minipage}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Effect:}\end{minipage} \begin{minipage}[t]{\rhwidth} #3 \end{minipage}

\noindent \hspace{.2cm} \begin{minipage}{1.5cm}{Cost:}\end{minipage} \begin{minipage}[t]{\rhwidth} \ensuremath{#4} \end{minipage}
\end{minipage}
}

\title{Sentence Generation as Planning with Probabilistic LTAG\\
%\Thanks{}
}

\author{Daniel Bauer\\
  Columbia University\\
  New York, NY, USA\\
  {\tt bauer@cs.columbia.edu} 
  \And
  Alexander Koller\\
  Saarland University\\
  Saarbr\"{u}cken, Germany\\ 
  {\tt koller@mmci.uni-saarland.de}}

\date{}

\begin{document}
\maketitle

\input abstract

\input{introduction}
\input{related} % Related work
\input{crisp} % Sentence Generation as Planning
\input{pcrisp} % Statistical Generation as Planning
\input{experiments} % Experiments
\input{conclusion}

\bibliographystyle{naaclhlt2010}
\bibliography{literature}

\end{document}