### 2010-05-21 trunk/Papers/pcrisp-10/related.tex
Minor changes. Camera ready. :)

\section{Related Work}
\label{sec:related}

Statistical methods are popular for surface realization, but have not been used in systems that integrate sentence planning. Most statistical generation approaches follow a generate-and-select strategy, first proposed by \newcite{knight1995} in their {\sc nitrogen} system. Such systems generate a set of candidate sentences using a (possibly overgenerating) grammar and then select the best output sentence by applying a statistical language model. This family includes systems such as {\sc HAL}ogen \cite{langkildeknight1998,langkilde2000} and OpenCCG \cite{whitebaldridge2003}.  The {\sc Fergus} system \cite{bangalorerambow2000} is a variant of this approach which, like {\sc pcrisp}, employs {\sc ltag}. It first assigns elementary trees to each entity in the input sentence plan using a statistical tree model and then computes the most likely derivation using only these trees with an $n$-gram model on the output sentence. An alternative to the $n$-gram based generate and select approach is to use a probabilistic grammar model, like {\sc   ptag}, trained on automatic parses \cite{zhongstent2005}. A related approach uses a model over local decisions of the generation system itself \cite{belz2008}. Both models can either be used to discriminate a set of output candidates, or more directly to choose the next best decision locally. Our approach is similar in that it uses {\sc ptag} to find the most likely output structure. However, the previous work discussed so far addresses surface realization only. We extend this to a statistical NLG algorithm which does surface realization and sentence planning at the same time.

Our treatment of integrated sentence planning and surface realization as planning is inherited directly from {\sc crisp} \cite{kollerstone2007}.  Planning has long played a role in generation, but has focused on discourse planning instead of specifically addressing sentence generation \cite{hovy1988,appelt1992}. The applicability of these ideas was limited at that time because efficient planning technology was not available. Recently the development of more efficient planning algorithms \cite{hoffmannnebel2001} spawned a renewed interest in planning for NLG.  {\sc Crisp} uses such algorithms to efficiently solve the sentence generation problem defined by {\sc spud} \cite{Stone2003a}. {\sc Spud}, which instead uses an incomplete greedy algorithm, is based on an {\sc ltag} whose trees are augmented with semantic and pragmatic constraints. Given a communicative goal, a solution to the {\sc spud} problem realizes this goal and simultaneously selects referring expressions. The next section explores {\sc crisp} in more detail.

%%% Local Variables:  
%%% mode: latex 
%%% TeX-master: "pcrisp-10" 
%%% End: