### 2010-10-27 trunk/Papers/operator-splitting/equiv-definition-carlos.tex
structure for "homework" defs

\subsection{Definition: Carlos}
\label{definition-carlos}