### 2010-10-27 trunk/Papers/operator-splitting/equiv-definition-gabi.tex

\subsection{Definition: Gabi}
\label{definition-gabi}