### 2010-11-18 trunk/Papers/operator-splitting/equiv-definition-carlos.tex

\newcommand{\tup}[1]{\langle #1 \rangle}

\subsection{Definition: Carlos}
\label{definition-carlos}

Let start by introducing some basic notation

\begin{definition}
An \emph{action} is a pair $a = \tup{P,E}$ where both $P$ and $E$ are finite sets of 
propositional literals. $P$ is the set of \emph{preconditions} of $a$. $E$ is the set of \emph{effects} of $a$.

A \emph{planning domain} $D$ is a finite set of actions. \fixme{If we want, we can define something like a `\emph{restricted planning domain}' as a pair $D=\tup{A,R}$ where $A$ is a finite set of actions, and $R$ is a finite set of 
propositional formulas$.  We would use $R$ to restrict planning tasks by requesting consistency with $R$.}

A \emph{planning task} is a tuple $\tup{D, P}$
where $D$ is a planning domain and  $P=\tup{I,F)$ 
is an \emph{instance description}, consisting of a finite set of literals $I$
(the \emph{initial state}) and the a finite set of literals $O$ (the \emph{final state}).

\end{definition}

\begin{definition}[Reduction]
A reduction of planning domain $D$ to planning domain $D'$ is a tuple
$\mathbf f=\left<f_P, f_{\pi'}\right>$ of functions
such that for all planning tasks $\Pi = \left<D, P\right>$:
\begin{enumerate}
    \item $\left<D', f_P(P)\right>$ is a planning task,
    \item for each plan $\pi'$ of $\left<D', f_P(P)\right>$, $f_{\pi'}(\pi')$ is
          a plan for $\Pi$,
    \item $f_P$ and $f_{\pi'}$ are polynomial-time computable,
    \item for each two plans $\pi'_1$ and $\pi'_2$ of $\left<D', f_P(P)\right>$
          it holds that $\textit{cost}(f_{\pi'}(\pi'_1)) >
          \textit{cost}(f_{\pi'}(\pi'_2))$ iff $\textit{cost}(\pi'_1) >
          \textit{cost}(\pi'_2)$, and
    \item for at least one optimal plan $\pi^*$ of $\Pi$ there is a
          plan $\pi'$ of $\left<D', f_P(P)\right>$
          whose size is polynomial in the size of $\pi^*$ and
          $\textit{cost}(\pi^*) = \textit{cost}(f_{\pi'}(\pi'))$.
\end{enumerate}
\end{definition}

\begin{definition}[Reducibility]
A planning domain $D$ is reducible to planning domain $D'$ (written $D\leq D'$)
if there is a reduction $\mathbf f$ from $D$ to $D'$.
\end{definition}

\begin{definition}[Equivalence of planning domains]
Two planning domains $D$ and $D'$ are equivalent (written $D\equiv D'$) iff
$D\leq D'$ and $D'\leq D$.
\end{definition}