### 2011-05-23 tags/crisp-1.1.6/trunk/Papers/operator-splitting/equiv-definition-alex.tex

\subsection{Definition: Alex}
\label{definition-alex}

My definition is inspired by existing work on comparing the generative
capacity of grammar formalisms \cite{miller99:_stron}.  The basic idea
is that in order to compare two parse trees (which may be in terms of
two different grammar formalisms, and therefore look quite different),
we interpret them both into the same algebra $A$.  We take the parse
trees to be equivalent if they map to the same object in $A$.  Two
grammars are equivalent if their parse tree languages map to the same
set of objects in $A$.  One example is classical weak equivalence, in
which $A$ is the string algebra and the interpretation function just
takes the yield of the parse trees; but clearly, more powerful
interpretation algebras can be used for a more fine-grained view on
equivalence.

\newcommand{\A}{{\mathcal A}}

The fundamental parallel I see is that plans are essentially like
parse trees, and planning problem instances are essentially like
grammars.  Let's say we have a planning domain $D$, and let $A =
\{a_1,\ldots,a_n\}$ be the set of all operator instances of $D$.  Then
a plan is a string in $A^*$. Note that $(A^*,\cdot,a_1,\ldots,a_n)$ is
an algebra where the $a_i$ are constants and $\cdot$ is binary
concatenation.  I write $\A_D$ for the string algebra I get from the
planning domain $D$.  I can now define plan equivalence as follows.

\begin{definition}
  Let $D$ and $D'$ be planning domains.  Let $I$ be some algebra, and
  let $h:\A_D \rightarrow I$ and $h':\A_{D'} \rightarrow I$ be algebra
  homomorphisms. Now if $\pi$ is a sequence of action instances of $D$
  and $\pi'$ is a sequence of action instances of $D'$, $\pi$ is
  equivalent to $\pi'$ with respect to $I,h,h'$ ($\pi \equiv \pi'$)
  iff $h(\pi) = h'(\pi')$.
\end{definition}

Note that equivalence of plans (i.e., sequences of action instances
that go from a specific initial state to a specific goal) is a special
case of this. It is straightforward to lift this definition to
planning tasks and domains:

\begin{definition}
  Let $t = \langle D,P \rangle$ and $t' = \langle D',P'\rangle$ be
  planning tasks, and let $plans(t)$ be the set of all plans for the task
  $t$. Then $t$ and $t'$ are equivalent if $h(plans(t)) =
  h'(plans(t'))$.

  Let $plans(D)$ be the union over all $plans(\langle D,P \rangle)$
  for all tasks $\langle D,P \rangle$ that can be formed with the
  signature defined by $D$.  Then we take two domains $D$ and $D'$ to
  be equivalent if $h(plans(D)) = h'(plans(D'))$.
\end{definition}

Let's look at some special cases to illustrate the definitions.

\newcommand{\states}{{\mathcal S}}
\newcommand{\partialfunctions}{{\mathcal PF}}

\begin{itemize}
\item The most trivial special case is to require plan equivalence to
  mean plan identity.  For this, we take $I = \A_D$ and $h$ and $h'$
  the identity function.
\item The other extreme case is a notion of plan equivalence in which
  all plans are equivalent.  Clearly, this can be achieved if we take
  $I$ to be an algebra that contains a single element $e$, and $h$ and
  $h'$ map all plans to $e$.
\item A more interesting case is Gabi's question about reordering of
  actions in a plan: We want to see two plans over the same domain as
  equivalent if they take us from the same initial states to the same
  goals, but we don't care how the actions are ordered inside the
  plans, as long as the ``meaning'' of the plan stays the same.  One
  way to approach this is to let $I$  be the algebra
  $\partialfunctions_D$ of partial functions $f:\states_D
  \rightsquigarrow \states_D$, where $\states_D$ is the set of all
  states specified by $D$.  If $a$ is an action instance, we say that
  $h(a)$ is defined on a given state $s$ if the preconditions of $a$
  are satisfied in $s$.  $h(a)(s)$ is then the state we get by
  executing $a$ in $s$.  Concatenation of action sequences is
  interpreted as composition of partial functions.  In this way, any
  action sequence is interpreted as the partial function that is
  defined if the action sequence can be executed in a state, and that
  modifies the state as the actions specify.
\item Finally, operator splitting.  This is basically like the
  previous point, except that one domain has a richer signature,
  containing predicates like ``processing-none'' and such.  Let's say
  that $D$ was the original domain, and $D'$ is the enriched domain
  for the split operators.  We can then interpret both the original
  plan $\pi$ and the operator-split plan $\pi'$ into
  $\partialfunctions_{D}$.  The homomorphism $h:\A_D
  \rightarrow \partialfunctions_D$ is as above.  The homomorphism
  $h':\A_{D'} \rightarrow \partialfunctions_D$ maps actions of $D'$
  into partial functions over $\states_D$ by ignoring all
  preconditions and effects that use predicates that exist in $D'$ but
  not in $D$.  In this way, the fact that plans over $D$ are shorter
  than the split-operator plans over $D'$ is irrelevant: Because $\pi$
  can be executed in the same states in which $\pi'$ can be executed,
  and the effects of the two plans are the same (as far as predicates
  of $D$ are concerned), the plans count as equivalent.
\end{itemize}

My definition proposal is quite different in character from those of
Carlos and Gabi. It is a model-theoretic rather than a computational
notion: Plan equivalence is defined in terms of an interpretation of
the plans into a third algebra, rather than in terms of a reduction.
However, I feel that this is actually an advantage because the
definition is more flexible and, as far as I can tell, simpler, and
relies on very standard concepts such as algebras and homomorphisms.

It is clear that we will typically be interested in algorithms that
transform one class of plans into another, such as the
operator-splitting algorithm in Section~\ref{initial-definition}.  But
this does not mean that equivalence itself must be defined
procedurally.  I would bet that any procedural definition of
equivalence that we can come up with can be cast in terms of my
definition, by selecting the appropriate interpretation algebra and
homomorphisms for which the reduction is sound and complete.  One
interesting twist is that Gabi's definition suggests that we may care
about runtime and space bounds for the reduction.  This is one point
for which it is not entirely clear to me how to capture it in my
world.  Perhaps we can talk about why we care about them.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "total"
%%% End: